### 2020-09-10 manuscript.Rmd
First rough draft of manuscript
Note - does not contain full citations

---
title             : "School Socioeconomic Status Context and Social Skills in Children"
shorttitle        : "School SES and Social Skills"
author: 
  - name          : "Philip Parker"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "33 Berry St, North Sydney, NSW, 2060, Australia"
    email         : "philip.parker@acu.edu.au"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing
affiliation:
  - id            : "1"
    institution   : "Institute for Positive Psychology and Education, Australian Catholic University"
  - id            : "2"
    institution   : "Konstanz Business School"
authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
abstract: |

  As researchers and policymakers try to understand the consequences of SES for educational and occupational attainment, they increasingly consider social skills to be critical. Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development. In a representative longitudinal sample of 5361 Australian children, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers. All models controlled for prior social skills at age 4, and a wide range of covariates. We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems. We found a consistent interaction between individual and school average socioeconomic status that suggested assimilation effects were only present for children from low socioeconomic backgrounds. We found that children from low socioeconomic backgrounds enter school with lower average social skills. They are more likely to be enrolled in more disadvantaged schools. Finally, the social skills of children from low socioeconomic backgrounds worsen in low SES contexts. Taking the evidence together, social stratified schools may harm children from low SES backgrounds while not benefiting children from high SES backgrounds.
keywords          : "social skills; assimilation effects; socioeconomic status; school context"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : 
  pdf: 
    papaja::apa6_pdf
  html_document: 
    fig_caption: yes
    toc: yes
  html_notebook: 
    fig_caption: yes
    toc: yes
    highlight: haddock
    theme: spacelab
  word:
    papaja::apa6_doc
---

```{r setup, include = FALSE}
# Create Reference List
library("papaja")
r_refs("r-references.bib")

# Set default Knitr Options
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, dpi = 300)

# Set tidyverse to quiet
library("tidyverse")
options(tidyverse.quiet = TRUE)
```

After decades of research and intervention aimed at reducing educational and occupational inequality, inequalities persist [@reardon2011]. In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman. [@heckman2006] argues that research and intervention on reducing educational inequality have focused too narrowly on "cognitive" abilities. Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment. For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment <!-- REFERENCE -->. Children from low socioeconomic backgrounds (here after low SES children[^1]) appear to enter school with poorer social skills (e.g., <!-- REFERENCE -->Jerrim & Sims, 2019). This is concerning as childhood social skills predict a wide variety of later life outcomes (e.g., <!-- REFERENCE -->Jones, Greenberg, & Crowley, 2015). Inspired by identity economics, our research asks whether the school context in the stratified school system of Australia may exacerbate this issue. In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior. We consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates in a representative longitudinal sample of Australian children.

[^1]: We acknowledge that children themselves do not have a socioeconomic status. Rather they are raised in an environment which is shaped by their parents' socioeconomic status (SES). However, for brevity we refer to children as low SES or high SES hereafter. Likewise, in place of low school average SES and high school average SES we refer to disadvantaged and advantaged schools respectively.

## Social Skills

Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable (Rios, Ling, Pugh, Becker, & Bacall, 2020). Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior (Durlak, Weissberg, Dymnicki, Taylor, & Schellinger, 2011). Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement (Corcoran, Cheung, Kim, & Xie, 2018) and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health (Jones et al., 2015). Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire (see Datta Gupta & Simonsen, 2010; Gutman & Schoon, 2013; Jerrim & Sims, 2019). We use this questionnaire to measure peer problems, conduct problems, and prosocial behavior. We model social skills using both parent and teacher reports.

## Contrast Processes and Identity Economics

Children's social skills depend in part on the context in which they develop (Jerrim & Sims, 2019). There are two major forces in social psychology---assimilation and contrast (Mussweiler, Rüter, & Epstude, 2004)---that may account for how school context affects children's social skill development. Contrast processes are in operation when children's perceptions, opinions, or behavior depend on their perceived rank order within their group. Assimilation processes are in operation when people's perceptions, opinions, or behaviour depend on reference group norms (Kelly, 1952).

The theory of *Identity Economics* implies an assimilative effect (Akerlof & Kranton, 2010). Identity Economics argues that the social context of the school possesses a gravity that attracts students' identity, values, and behavior toward the prototypical identity, values, and behavior of the school (Akerlof & Kranton, 2002). School ethos and norms around behavior incentivize children to act in a manner consistent with these norms. School average socioeconomic status could, thus, affect social skills. A now large body of research shows that social skills are correlated with socioeconomic status (SES; de Laat, Essink-Bot, van Wassenaer-Leemhuis, & Vrijkotte, 2015; Garratt, Chandola, Purdam, & Wood, 2016; McMunn, Nazroo, Marmot, Boreham, & Goodman, 2001; Rajmil et al., 2013) and, this relationship is present before children enter school (Washbrook & Waldfogel, 2011). In socially stratified school systems, children of similar SES tend to be schooled together. Thus, if children assimilate to the school context they are in, as Identity Economics would suggest, we would expect school average socioeconomic status to influence the development of social skills.

Jerrim and Sims (2019) provide one test of the potential assimilative effect of schools on social skills. They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school. They also found some assimilative effects of school average SES predicting social skills using the total score of the strengths and difficulties questionnaire (SDQ). The assimilation effects they found, however, were weak. This may be because the authors used the SDQ total score---which includes a range of social and emotional behaviors---rather than focusing on specific social skill components of the SDQ. Further, assimilative effects may be non-linear. Put another way, the assimilative power of school contexts may not be evenly distributed across the socioeconomic distribution. This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized (Gradstein & Justman, 2005).

# Current Study

Australia is the context of our research. Australia is a useful country to focus on because Australia's Programme for International Student Assessment (PISA) Index of Social Inclusion score is relatively low (OECD, 2015). This means that Australian schools are socially stratified (Parker et al., 2018b, 2019). Our major hypothesis is that school average socioeconomic status will have an assimilation effect on association with social skills at Year 3 (age 8) controlling for incoming social skills (age 4), individual socioeconomic status (age 4) and a range of demographic and achievement covariates[^2]. We also anticipate that there will be differential associations of school average socioeconomic status on social skills across the socioeconomic status gradient.

[^2]: This assimilation effect will be indicated by a negative coefficient for peer problems and conduct problems regressed on school average socioeconomic status and a positive coefficient for prosocial behavior regressed on school average socioeconomic status. This is because of the way in which these factors are valanced (high scores indicate high levels of prosocial behavior, but also high levels of peer and conduct problems).

We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills (de Laat, Essink-Bot, van Wassenaer-Leemhuis, & Vrijkotte, 2015; Garratt, Chandola, Purdam, & Wood, 2016; McMunn, Nazroo, Marmot, Boreham, & Goodman, 2001; Rajmil et al., 2013). Given Australian schools are socially stratified (OECD, 2015; Parker et al., 2018), we also expect low SES children to attend more disadvantaged schools on average.

# Method

## Participants and Study Design

We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC). LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004. Both cohorts of children have been followed every two years since (AIFS, 2015). LSAC is a complex survey with a stratified sampling design (that we account for in all analyses) and includes linked administration records for student performance in a standardized national numeracy test. We also use government collected data on school average socioeconomic status. Many children in the sample were the only child surveyed (27%) or were one of two children surveyed (36%) in their school. By using administrative records, we gain school average socioeconomic status based on the child's $\text{\underline{whole}}$ school grade from a high-quality source, thus avoiding sampling bias present in most school context research. We excluded participants who were not in school (e.g., were home schooled), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability). Together our total sample was 5,361 (52.1% male) children in Year 3 at school (\~age 8). In the vast majority of cases (98%), parent data came from mothers. Weights were provided by the survey organizers to ensure the sample represents the Australian population. This includes weights provided that aim to account for attrition.

## Measures

### Social skills at age 4 and 8

Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (Goodman, 1997). We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child’s *parent* and the child’s *teacher*. The SDQ asks respondents to rate a child’s behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true). Questions ask about the child’s peer problems (“Rather solitary, tends to play alone”), conduct problems (“Often fights with other children or bullies them”), and prosociality (“Is kind to younger children”). We used the total scores for these scales as developed by the LSAC administrators. Scores ranged from 0 to 10. These scores were heavily left censored (see Figures 2-3) for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher. prosocial behaviors were heavily right censored. Parents and teachers had moderate agreement for children at age 8 with correlations ranging from r = .40 for peer and r = .39 for conduct problems to r = .28 for prosocial behavior.

The SDQ items measure a collection of associated behaviours and as such we treat them as index or a formative variable. In this case common reliability metrics do not make much sense; nor do reflective models like confirmatory factor analysis. Instead, we would want items that are not too redundant with each other. For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index (Diamantopoulos, Riefler, & Roth, 2008; Thongrattana, 2010). In our case no VIF for any item for any social skills construct was higher than 2. Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct (Thongrattana, 2010). This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report. Parallel analysis suggested that all social skills constructs could be represented by a single principal component (but in all cases more than one reflective factor). The only exception was age 4 peer problems were parallel analysis suggested two components but the eigenvalue on the second component was barely greater than the simulated data suggesting a single index was likely still appropriate.

### Socioeconomic Status

*Individual SES* SES was measured using the Socioeconomic Position (SEP) index constructed by the LSAC survey organizers (Baker, Sipthorp, & Edwards, 2017). The SEP index is constructed from parent reported standardized weekly income, years of education, and ANU4 occupational prestige derived from the Australian Standard Classification of Occupations. The SEP has a mean of zero and a standard deviation of one.

*School average SES* was measured using the Index of Community Socioeconomic Advantage (ICSEA) that the Australian government uses to assess the relative advantage of schools for the purpose of funding allocation and policy. This measure was taken from government administration records. ICSEA has a mean of 1000 and a standard deviation of 100. We z-scored this variable for analysis.

### Covariates

Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program – Literacy and Numeracy (NAPLAN) test results. The relevant NAPLAN tests for this study are given to all eligible children in the country in Year 3 (age 8). The tests are scaled so they are comparable across age cohort and across year grade. These high-stakes tests have a mean of 500 and a standard deviation of 100. For verbal ability we used the Peabody Picture Vocabulary test (Dunn & Dunn, 2007) given to participants at age 4 (i.e., at or just before school commencement). Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender—all measured at age 4. Both numeracy achievement and verbal ability were z-scored in the analysis.

## Analysis

There was a relatively small amount of missing data for most variables. However, given the data were longitudinal and included both linked administrative records and data collected from both the child’s teacher and parent, missing data was inevitable. Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample. Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%. All other variables had 6% missing data or less. To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, Blackwell, 2011). These 10 imputations were used in all analyses. All models used cluster robust standard errors to adjust standard errors for school clustering. LSAC strata was included as a fixed effect to account for the complex survey structure of the data. Analysis was done in the Zelig package in R (Imia, King, & Lau, 2008).

All models in the paper that predicted social skills were estimated using Tobit models. Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see Kleiber & Zeileis, 2008). As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models. Johnson-Neyman regions of significance were used to investigate significant interactions. We also attempted to run random-intercept models using for school id both maximum likelihood estimation and MCMC sampling. In both cases models did not converge likely due to the unique nature of the data (i.e., many participants were the single representative of their school).

Contextual effects like those reported here can be confusing to interpret. This is because positive effects of contextual variables can be incorrectly interpreted as positive in implication (or vice-versa for negative effects). This issue is compounded when interactions involving contextual effects are included in the model. Thus, we also provide counterfactual estimates of expected score differences for different levels of student SES and school average SES when the interaction between these variables is significant. Because of the number of covariates (35 in total because many of the covariates, such as strata, were categorical and thus were included as a series of dummy variables), we only report results for the main analysis variables.

# Results

## Low SES Children Start School with Poorer Social Skills and Enroll in More Disadvantaged Schools

In order to provide a context for the main analyses we first aimed to show a) that richer children enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools. Our findings show that high SES children do enter school with better social skills. Student SES is correlated with less peer (B = -.284 [-.345, -.222]) and conduct (B = -.371 [-.438, -.304]) problems. They also had slightly higher prosocial behavior (B = .074 [.009, .139  ]). Further, student SES was strongly related to the school average SES of the school they attended (r = .480 [.458, .499]; see Figure 1).

## School Average SES Predicts Social Skills Controlling for Age 4 SES and Social Skills

We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4. Results are presented under the Model 1 column in Tables 1-3. Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3). School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills. School average SES also negatively predicted peer problems but only for parent reports.

## School Average SES Predicts Social Skills Mainly in Low SES Children

The predictive association of school average SES on social skills was not consistent across the SES gradient (see Model 2 results reported in Tables 1-3). Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior. Regions of significance for these interactions can be found in Figures 2-4. Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4. These results show that low SES children appear to be most influenced by school average SES. Middle to high SES children appeared to have largely similar scores regardless of the school they were in (see Figures 2-4). For example, both parents and teachers report that low SES children had approximately half a point more peer problems in more SES disadvantaged schools than in more SES advantaged schools (see Table 4). High SES children had almost no difference in peer problems regardless of the school they were in.

# Discussion

Research on social skills has repeatedly shown that there is a socioeconomic status gradient to social skills (see Datta Gupta & Simonsen, 2010; Gutman & Schoon, 2013; Jerrim & Sims, 2019). Yet little research in this area has considered the potential influence of school context on social skill development (cf. Jerrim & Sims, 2019). This is a lacuna that our research sought to fill. Our research considered the association of school socioeconomic context with social skills in early elementary school. We used both parent and teacher reported social skills, finding surprisingly consistent effect sizes regardless of reporting source despite their relatively modest agreement. Importantly, our research used a number of critical controls as well as prior social skills that helped us better identify the effect of school average socioeconomic status.

For both teacher and parent reports we found that the association of school context with social skills depended on the individual child’s own socioeconomic background. This was the case for all outcomes except for parent reported prosocial behavior. It is worth emphasizing that the nature of these interactions was consistent for all outcomes. Namely, that middle to high SES children were largely unaffected by school socioeconomic context while the children with the lowest socioeconomic status experienced the largest effect.

## School Context Theory

A greater focus on social skills as explanations for socioeconomic gaps in educational attainment has been an important step forward in inequality research (see Heckman, 2006). Now that research and theory has illuminated the importance of such skills, research needs to consider the conditions under which they develop. Previous economic theory has emphasized the role that schools play as a context for the development of non-academic factors like social skills and claimed this as one of the ways in which intergenerational inequality is transmitted (Bowles & Gintis, 1976). Such theory has tended to emphasize the role of teachers and systems, relegating  students’ fellow classmates to a more secondary role. In contrast, psychology research has tended to emphasize the role of frames-of-reference with particular emphasize given to the role of a child’s peers as providing a standard against which a child might assimilate to or contrast against (Mussweiler, Rüter, & Epstude, 2004). It seems likely that both economic and psychology theory is right to some extent but the relative contribution to school context effects of system, teacher, and peers is unclear. Our research shows that school context effects are significant and mainly seem to effect low SES children. Future research is needed to determine what the most important mechanisms are that explain this effect.

## School Context and Assimilation Effects

For children with SES status below the mean, school socioeconomic context had statistically significant associations with social skills in Year 3 (controlling for social skills at age 4). The implications of this are mixed. Our results do suggest that a low SES child who is enrolled in an advantaged school should expect to have similar levels of social skills as their high SES peers. The problem is that in countries like Australia low SES children are relatively rarely enrolled in advantaged schools. As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA (OECD, 2015). This means that Australian schools tend to be homogenous in terms of their student composition (see Parker, Sanders, & Guo, 2019). As our results show low SES children tend to attend disadvantaged schools and their social skills appear to suffer as a result. What our findings suggest is that children’s social skills, particularly low SES children, acclimatize toward the average of the school they are in. Consistent with the literature we found that low SES children start school with lower social skills. In a country like Australia, whose schools are socially stratified, low SES children will tend to be schooled in more disadvantaged schools (see Figure 1). The net effect in such a system is that low SES children will tend to have their already lower social skills depressed by the school climate they are most likely to find themselves in.

This counterfactual may suggest that school choice, in which low SES children receive vouchers or similar, may be beneficial (see Friedman, 1962). Strategies like school vouchers to attend magnet schools could be a powerful policy lever to overcome socioeconomic gaps in social skills. This approach tackles the problem of assimilative contextual effects via market-based systems. Yet this policy requires there to be few barriers, whether psychological or otherwise, to parents using vouchers to select the best school that matches the needs of their child. But this does not seem to be the case (Gradstein & Justman, 2005). Indeed, empirical evidence suggests that school choice tends to exacerbate inequality (e.g., Saporito, 2003).

Parker et al. (2016, 2018a, 2018b, 2019) have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children. They argue that empirical evidence suggests policy should encourage school selection policies that maximize within school heterogeneity. This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents’ rights to choose. However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a disadvantaged school.

There are strong arguments and good empirical support on both sides of this debate, suggesting that we are far from a settled position on the matter. At least for the current context in Australia where social stratification is moderately high and where the school system seems to ensure school choice is more clearly an option for the rich than the poor (Parker et al., 2019), our results are troubling. In particular, they suggest low SES children face a triple bind. First, they enter school with lower social skills. Second, they enter schools where assimilation effects are likely to further dampen social skill development. Third, they appear particularly vulnerable to their school context in this respect.

## Limitations

There are many strengths to this study. Most notably, the use of longitudinal data that allowed for us to control for incoming social skills and government administrative data that provided access to complete and high-quality data for school average socioeconomic status at the school level. Further the use of LSAC data meant that we were able to control for a number of potential confounding variables drawn from a longitudinal representative sample of Australian children. Nevertheless, there are limitations. Our aim was to try to build a model from high-quality data that could assist us in making as close to an *all else being equal* comparisons as possible (Angrist & Pischke, 2009). It is for this reason we have cautiously used causal language in relation to our claims. But our claim to causality would have been stronger had they been backed by an experimental design. While the Move to Opportunity program in the US suggest experiments where low SES children are randomly assigned to richer schools (or at least randomly assigned to areas with richer schools) are possible (see de Souza Briggs, Popkin & Goering, 2010), it is hard to imagine a situation in which richer children could be randomly assigned to more disadvantaged schools. Given the non-linearity in effects we observed this would be a serious limitation of an experimental design.

Finally, we were not able to identify and compare the relative impact of different mechanisms that may explain the influence of school average socioeconomic status on social skills. As educational psychologists our study is framed in relation to assimilation effects in response to children’s peer frames-of-reference. Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see Bowls & Gintis, 2001). Identifying and comparing these mechanisms is an important future direction for research.

# Conclusion

The influence of school average socioeconomic status on social skills represent the triple disadvantage that low SES children can face in social stratified school systems. First, low SES children are more likely to start school with lower social skills than their high SES peers. Second, because the school system is stratified by socioeconomic status, low SES children are likely to enroll in more disadvantaged schools. Third, assimilative associations suggest low SES children, already disadvantaged by prior social skill gaps, are more affected by their school context than are middle to high SES children. Thus, while low SES children assimilate to lower social skill environments, richer children appear to gain little advantage from their more conducive environments and appear to function equally well in advantaged and disadvantaged schools. Taken together our results support the call for a policy focus that aims to  a) decrease country level variance in social stratification, b) decrease between school heterogeneity is social status, and c) in combination with (a), encourage school selection policies that maximize within school heterogeneity in social status.

\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```
::: {#refs custom-style="Bibliography"}
:::

\endgroup

### 2020-09-10 manuscript.Rmd
Initial Rmd

@@
 ---
-title             : "School Socioeconomic Status Context and Social Skills in Children"
-shorttitle        : "School SES and Social Skills"
+title: "INSERT_TITLE_HERE"
 author: 
-  - name          : "Philip Parker"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "33 Berry St, North Sydney, NSW, 2060, Australia"
-    email         : "philip.parker@acu.edu.au"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Writing - Original Draft Preparation
-      - Writing - Review & Editing
-  - name          : "Ernst-August Doelle"
-    affiliation   : "1,2"
-    role:
-      - Writing - Review & Editing
-affiliation:
-  - id            : "1"
-    institution   : "Institute for Positive Psychology and Education, Australian Catholic University"
-  - id            : "2"
-    institution   : "Konstanz Business School"
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-  Enter author note here.
-abstract: |
+  - name: Philip Parker
+    affiliation: Australian Catholic University
 
-  As researchers and policymakers try to understand the consequences of SES for educational and occupational attainment, they increasingly consider social skills to be critical.
-Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development.
-In a representative longitudinal sample of 5361 Australian children, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers.
-All models controlled for prior social skills at age 4, and a wide range of covariates.
-We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems.
-We found a consistent interaction between individual and school average socioeconomic status that suggested assimilation effects were only present for children from low socioeconomic backgrounds.
-We found that children from low socioeconomic backgrounds enter school with lower average social skills.
-They are more likely to be enrolled in more disadvantaged schools.
-Finally, the social skills of children from low socioeconomic backgrounds worsen in low SES contexts.
-Taking the evidence together, social stratified schools may harm children from low SES backgrounds while not benefiting children from high SES backgrounds.
-keywords          : "social skills; assimilation effects; socioeconomic status; school context"
-wordcount         : "X"
-bibliography      : ["r-references.bib"]
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-documentclass     : "apa6"
-classoption       : "man"
-output            : 
-  pdf: 
-    papaja::apa6_pdf
-  html_document: 
-    fig_caption: yes
-    toc: yes
-  html_notebook: 
-    fig_caption: yes
-    toc: yes
-    highlight: haddock
-    theme: spacelab
-  word:
-    papaja::apa6_doc
+date: September 8, 2020
+output: 
+  hrbrthemes::ipsum:
+    toc: true
 ---
-
-```{r setup, include = FALSE}
-# Create Reference List
-library("papaja")
-r_refs("r-references.bib")
-
-# Set default Knitr Options
-knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, dpi = 300)
-
-# Set tidyverse to quiet
-library("tidyverse")
-options(tidyverse.quiet = TRUE)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300)
 ```
 
-After decades of research and intervention aimed at reducing educational and occupational inequality, inequalities persist [@reardon2011].
-In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman.
-[@heckman2006] argues that research and intervention on reducing educational inequality have focused too narrowly on "cognitive" abilities.
-Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment.
-For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment <!-- REFERENCE -->.
-Children from low socioeconomic backgrounds (here after low SES children[^1]) appear to enter school with poorer social skills (e.g., <!-- REFERENCE -->Jerrim & Sims, 2019).
-This is concerning as childhood social skills predict a wide variety of later life outcomes (e.g., <!-- REFERENCE -->Jones, Greenberg, & Crowley, 2015).
-Inspired by identity economics, our research asks whether the school context in the stratified school system of Australia may exacerbate this issue.
-In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior.
-We consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates in a representative longitudinal sample of Australian children.
-
-[^1]: We acknowledge that children themselves do not have a socioeconomic status.
-Rather they are raised in an environment which is shaped by their parents' socioeconomic status (SES).
-However, for brevity we refer to children as low SES or high SES hereafter.
-Likewise, in place of low school average SES and high school average SES we refer to disadvantaged and advantaged schools respectively.
-
-## Social Skills
-
-Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable (Rios, Ling, Pugh, Becker, & Bacall, 2020).
-Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior (Durlak, Weissberg, Dymnicki, Taylor, & Schellinger, 2011).
-Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement (Corcoran, Cheung, Kim, & Xie, 2018) and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health (Jones et al., 2015).
-Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire (see Datta Gupta & Simonsen, 2010; Gutman & Schoon, 2013; Jerrim & Sims, 2019).
-We use this questionnaire to measure peer problems, conduct problems, and prosocial behavior.
-We model social skills using both parent and teacher reports.
-
-## Contrast Processes and Identity Economics
-
-Children's social skills depend in part on the context in which they develop (Jerrim & Sims, 2019).
-There are two major forces in social psychology---assimilation and contrast (Mussweiler, Rüter, & Epstude, 2004)---that may account for how school context affects children's social skill development.
-Contrast processes are in operation when children's perceptions, opinions, or behavior depend on their perceived rank order within their group.
-Assimilation processes are in operation when people's perceptions, opinions, or behaviour depend on reference group norms (Kelly, 1952).
-
-The theory of *Identity Economics* implies an assimilative effect (Akerlof & Kranton, 2010).
-Identity Economics argues that the social context of the school possesses a gravity that attracts students' identity, values, and behavior toward the prototypical identity, values, and behavior of the school (Akerlof & Kranton, 2002).
-School ethos and norms around behavior incentivize children to act in a manner consistent with these norms.
-School average socioeconomic status could, thus, affect social skills.
-A now large body of research shows that social skills are correlated with socioeconomic status (SES; de Laat, Essink-Bot, van Wassenaer-Leemhuis, & Vrijkotte, 2015; Garratt, Chandola, Purdam, & Wood, 2016; McMunn, Nazroo, Marmot, Boreham, & Goodman, 2001; Rajmil et al., 2013) and, this relationship is present before children enter school (Washbrook & Waldfogel, 2011).
-In socially stratified school systems, children of similar SES tend to be schooled together.
-Thus, if children assimilate to the school context they are in, as Identity Economics would suggest, we would expect school average socioeconomic status to influence the development of social skills.
-
-Jerrim and Sims (2019) provide one test of the potential assimilative effect of schools on social skills.
-They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school.
-They also found some assimilative effects of school average SES predicting social skills using the total score of the strengths and difficulties questionnaire (SDQ).
-The assimilation effects they found, however, were weak.
-This may be because the authors used the SDQ total score---which includes a range of social and emotional behaviors---rather than focusing on specific social skill components of the SDQ.
-Further, assimilative effects may be non-linear.
-Put another way, the assimilative power of school contexts may not be evenly distributed across the socioeconomic distribution.
-This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized (Gradstein & Justman, 2005).
-
-# Current Study
-
-Australia is the context of our research.
-Australia is a useful country to focus on because Australia's Programme for International Student Assessment (PISA) Index of Social Inclusion score is relatively low (OECD, 2015).
-This means that Australian schools are socially stratified (Parker et al., 2018b, 2019).
-Our major hypothesis is that school average socioeconomic status will have an assimilation effect on association with social skills at Year 3 (age 8) controlling for incoming social skills (age 4), individual socioeconomic status (age 4) and a range of demographic and achievement covariates[^2].
-We also anticipate that there will be differential associations of school average socioeconomic status on social skills across the socioeconomic status gradient.
-
-[^2]: This assimilation effect will be indicated by a negative coefficient for peer problems and conduct problems regressed on school average socioeconomic status and a positive coefficient for prosocial behavior regressed on school average socioeconomic status.
-This is because of the way in which these factors are valanced (high scores indicate high levels of prosocial behavior, but also high levels of peer and conduct problems).
-
-We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills (de Laat, Essink-Bot, van Wassenaer-Leemhuis, & Vrijkotte, 2015; Garratt, Chandola, Purdam, & Wood, 2016; McMunn, Nazroo, Marmot, Boreham, & Goodman, 2001; Rajmil et al., 2013).
-Given Australian schools are socially stratified (OECD, 2015; Parker et al., 2018), we also expect low SES children to attend more disadvantaged schools on average.
-
-# Method
-
-## Participants and Study Design
-
-We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC).
-LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004.
-Both cohorts of children have been followed every two years since (AIFS, 2015).
-LSAC is a complex survey with a stratified sampling design (that we account for in all analyses) and includes linked administration records for student performance in a standardized national numeracy test.
-We also use government collected data on school average socioeconomic status.
-Many children in the sample were the only child surveyed (27%) or were one of two children surveyed (36%) in their school.
-By using administrative records, we gain school average socioeconomic status based on the child's $\text{\underline{whole}}$ school grade from a high-quality source, thus avoiding sampling bias present in most school context research.
-We excluded participants who were not in school (e.g., were home schooled), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
-Together our total sample was 5,361 (52.1% male) children in Year 3 at school (\~age 8).
-In the vast majority of cases (98%), parent data came from mothers.
-Weights were provided by the survey organizers to ensure the sample represents the Australian population.
-This includes weights provided that aim to account for attrition.
-
-## Measures
-
-### Social skills at age 4 and 8
-
-Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (Goodman, 1997).
-We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child’s *parent* and the child’s *teacher*.
-The SDQ asks respondents to rate a child’s behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
-Questions ask about the child’s peer problems (“Rather solitary, tends to play alone”), conduct problems (“Often fights with other children or bullies them”), and prosociality (“Is kind to younger children”).
-We used the total scores for these scales as developed by the LSAC administrators.
-Scores ranged from 0 to 10.
-These scores were heavily left censored (see Figures 2-3) for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher.
-prosocial behaviors were heavily right censored.
-Parents and teachers had moderate agreement for children at age 8 with correlations ranging from r = .40 for peer and r = .39 for conduct problems to r = .28 for prosocial behavior.
-
-The SDQ items measure a collection of associated behaviours and as such we treat them as index or a formative variable.
-In this case common reliability metrics do not make much sense; nor do reflective models like confirmatory factor analysis.
-Instead, we would want items that are not too redundant with each other.
-For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index (Diamantopoulos, Riefler, & Roth, 2008; Thongrattana, 2010).
-In our case no VIF for any item for any social skills construct was higher than 2.
-Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct (Thongrattana, 2010).
-This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report.
-Parallel analysis suggested that all social skills constructs could be represented by a single principal component (but in all cases more than one reflective factor).
-The only exception was age 4 peer problems were parallel analysis suggested two components but the eigenvalue on the second component was barely greater than the simulated data suggesting a single index was likely still appropriate.
-
-### Socioeconomic Status
-
-*Individual SES* SES was measured using the Socioeconomic Position (SEP) index constructed by the LSAC survey organizers (Baker, Sipthorp, & Edwards, 2017).
-The SEP index is constructed from parent reported standardized weekly income, years of education, and ANU4 occupational prestige derived from the Australian Standard Classification of Occupations.
-The SEP has a mean of zero and a standard deviation of one.
-
-*School average SES* was measured using the Index of Community Socioeconomic Advantage (ICSEA) that the Australian government uses to assess the relative advantage of schools for the purpose of funding allocation and policy.
-This measure was taken from government administration records.
-ICSEA has a mean of 1000 and a standard deviation of 100.
-We z-scored this variable for analysis.
-
-### Covariates
-
-Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program – Literacy and Numeracy (NAPLAN) test results.
-The relevant NAPLAN tests for this study are given to all eligible children in the country in Year 3 (age 8).
-The tests are scaled so they are comparable across age cohort and across year grade.
-These high-stakes tests have a mean of 500 and a standard deviation of 100.
-For verbal ability we used the Peabody Picture Vocabulary test (Dunn & Dunn, 2007) given to participants at age 4 (i.e., at or just before school commencement).
-Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender—all measured at age 4.
-Both numeracy achievement and verbal ability were z-scored in the analysis.
-
-## Analysis
-
-There was a relatively small amount of missing data for most variables.
-However, given the data were longitudinal and included both linked administrative records and data collected from both the child’s teacher and parent, missing data was inevitable.
-Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample.
-Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
-All other variables had 6% missing data or less.
-To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, Blackwell, 2011).
-These 10 imputations were used in all analyses.
-All models used cluster robust standard errors to adjust standard errors for school clustering.
-LSAC strata was included as a fixed effect to account for the complex survey structure of the data.
-Analysis was done in the Zelig package in R (Imia, King, & Lau, 2008).
-
-All models in the paper that predicted social skills were estimated using Tobit models.
-Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see Kleiber & Zeileis, 2008).
-As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
-Johnson-Neyman regions of significance were used to investigate significant interactions.
-We also attempted to run random-intercept models using for school id both maximum likelihood estimation and MCMC sampling.
-In both cases models did not converge likely due to the unique nature of the data (i.e., many participants were the single representative of their school).
-
-Contextual effects like those reported here can be confusing to interpret.
-This is because positive effects of contextual variables can be incorrectly interpreted as positive in implication (or vice-versa for negative effects).
-This issue is compounded when interactions involving contextual effects are included in the model.
-Thus, we also provide counterfactual estimates of expected score differences for different levels of student SES and school average SES when the interaction between these variables is significant.
-Because of the number of covariates (35 in total because many of the covariates, such as strata, were categorical and thus were included as a series of dummy variables), we only report results for the main analysis variables.
-
-# Results
-
-## Low SES Children Start School with Poorer Social Skills and Enroll in More Disadvantaged Schools
-
-In order to provide a context for the main analyses we first aimed to show a) that richer children enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
-Our findings show that high SES children do enter school with better social skills.
-Student SES is correlated with less peer (B = -.284 [-.345, -.222]) and conduct (B = -.371 [-.438, -.304]) problems.
-They also had slightly higher prosocial behavior (B = .074 [.009, .139  ]).
-Further, student SES was strongly related to the school average SES of the school they attended (r = .480 [.458, .499]; see Figure 1).
-
-## School Average SES Predicts Social Skills Controlling for Age 4 SES and Social Skills
-
-We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under the Model 1 column in Tables 1-3.
-Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).
-School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills.
-School average SES also negatively predicted peer problems but only for parent reports.
-
-## School Average SES Predicts Social Skills Mainly in Low SES Children
-
-The predictive association of school average SES on social skills was not consistent across the SES gradient (see Model 2 results reported in Tables 1-3).
-Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior.
-Regions of significance for these interactions can be found in Figures 2-4.
-Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4.
-These results show that low SES children appear to be most influenced by school average SES.
-Middle to high SES children appeared to have largely similar scores regardless of the school they were in (see Figures 2-4).
-For example, both parents and teachers report that low SES children had approximately half a point more peer problems in more SES disadvantaged schools than in more SES advantaged schools (see Table 4).
-High SES children had almost no difference in peer problems regardless of the school they were in.
-
-# Discussion
-
-Research on social skills has repeatedly shown that there is a socioeconomic status gradient to social skills (see Datta Gupta & Simonsen, 2010; Gutman & Schoon, 2013; Jerrim & Sims, 2019).
-Yet little research in this area has considered the potential influence of school context on social skill development (cf.
-Jerrim & Sims, 2019).
-This is a lacuna that our research sought to fill.
-Our research considered the association of school socioeconomic context with social skills in early elementary school.
-We used both parent and teacher reported social skills, finding surprisingly consistent effect sizes regardless of reporting source despite their relatively modest agreement.
-Importantly, our research used a number of critical controls as well as prior social skills that helped us better identify the effect of school average socioeconomic status.
-
-For both teacher and parent reports we found that the association of school context with social skills depended on the individual child’s own socioeconomic background.
-This was the case for all outcomes except for parent reported prosocial behavior.
-It is worth emphasizing that the nature of these interactions was consistent for all outcomes.
-Namely, that middle to high SES children were largely unaffected by school socioeconomic context while the children with the lowest socioeconomic status experienced the largest effect.
-
-## School Context Theory
-
-A greater focus on social skills as explanations for socioeconomic gaps in educational attainment has been an important step forward in inequality research (see Heckman, 2006).
-Now that research and theory has illuminated the importance of such skills, research needs to consider the conditions under which they develop.
-Previous economic theory has emphasized the role that schools play as a context for the development of non-academic factors like social skills and claimed this as one of the ways in which intergenerational inequality is transmitted (Bowles & Gintis, 1976).
-Such theory has tended to emphasize the role of teachers and systems, relegating  students’ fellow classmates to a more secondary role.
-In contrast, psychology research has tended to emphasize the role of frames-of-reference with particular emphasize given to the role of a child’s peers as providing a standard against which a child might assimilate to or contrast against (Mussweiler, Rüter, & Epstude, 2004).
-It seems likely that both economic and psychology theory is right to some extent but the relative contribution to school context effects of system, teacher, and peers is unclear.
-Our research shows that school context effects are significant and mainly seem to effect low SES children.
-Future research is needed to determine what the most important mechanisms are that explain this effect.
-
-## School Context and Assimilation Effects
-
-For children with SES status below the mean, school socioeconomic context had statistically significant associations with social skills in Year 3 (controlling for social skills at age 4).
-The implications of this are mixed.
-Our results do suggest that a low SES child who is enrolled in an advantaged school should expect to have similar levels of social skills as their high SES peers.
-The problem is that in countries like Australia low SES children are relatively rarely enrolled in advantaged schools.
-As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA (OECD, 2015).
-This means that Australian schools tend to be homogenous in terms of their student composition (see Parker, Sanders, & Guo, 2019).
-As our results show low SES children tend to attend disadvantaged schools and their social skills appear to suffer as a result.
-What our findings suggest is that children’s social skills, particularly low SES children, acclimatize toward the average of the school they are in.
-Consistent with the literature we found that low SES children start school with lower social skills.
-In a country like Australia, whose schools are socially stratified, low SES children will tend to be schooled in more disadvantaged schools (see Figure 1).
-The net effect in such a system is that low SES children will tend to have their already lower social skills depressed by the school climate they are most likely to find themselves in.
-
-This counterfactual may suggest that school choice, in which low SES children receive vouchers or similar, may be beneficial (see Friedman, 1962).
-Strategies like school vouchers to attend magnet schools could be a powerful policy lever to overcome socioeconomic gaps in social skills.
-This approach tackles the problem of assimilative contextual effects via market-based systems.
-Yet this policy requires there to be few barriers, whether psychological or otherwise, to parents using vouchers to select the best school that matches the needs of their child.
-But this does not seem to be the case (Gradstein & Justman, 2005).
-Indeed, empirical evidence suggests that school choice tends to exacerbate inequality (e.g., Saporito, 2003).
-
-Parker et al.
-(2016, 2018a, 2018b, 2019) have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
-They argue that empirical evidence suggests policy should encourage school selection policies that maximize within school heterogeneity.
-This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents’ rights to choose.
-However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a disadvantaged school.
-
-There are strong arguments and good empirical support on both sides of this debate, suggesting that we are far from a settled position on the matter.
-At least for the current context in Australia where social stratification is moderately high and where the school system seems to ensure school choice is more clearly an option for the rich than the poor (Parker et al., 2019), our results are troubling.
-In particular, they suggest low SES children face a triple bind.
-First, they enter school with lower social skills.
-Second, they enter schools where assimilation effects are likely to further dampen social skill development.
-Third, they appear particularly vulnerable to their school context in this respect.
-
-## Limitations
-
-There are many strengths to this study.
-Most notably, the use of longitudinal data that allowed for us to control for incoming social skills and government administrative data that provided access to complete and high-quality data for school average socioeconomic status at the school level.
-Further the use of LSAC data meant that we were able to control for a number of potential confounding variables drawn from a longitudinal representative sample of Australian children.
-Nevertheless, there are limitations.
-Our aim was to try to build a model from high-quality data that could assist us in making as close to an *all else being equal* comparisons as possible (Angrist & Pischke, 2009).
-It is for this reason we have cautiously used causal language in relation to our claims.
-But our claim to causality would have been stronger had they been backed by an experimental design.
-While the Move to Opportunity program in the US suggest experiments where low SES children are randomly assigned to richer schools (or at least randomly assigned to areas with richer schools) are possible (see de Souza Briggs, Popkin & Goering, 2010), it is hard to imagine a situation in which richer children could be randomly assigned to more disadvantaged schools.
-Given the non-linearity in effects we observed this would be a serious limitation of an experimental design.
-
-Finally, we were not able to identify and compare the relative impact of different mechanisms that may explain the influence of school average socioeconomic status on social skills.
-As educational psychologists our study is framed in relation to assimilation effects in response to children’s peer frames-of-reference.
-Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see Bowls & Gintis, 2001).
-Identifying and comparing these mechanisms is an important future direction for research.
-
-# Conclusion
-
-The influence of school average socioeconomic status on social skills represent the triple disadvantage that low SES children can face in social stratified school systems.
-First, low SES children are more likely to start school with lower social skills than their high SES peers.
-Second, because the school system is stratified by socioeconomic status, low SES children are likely to enroll in more disadvantaged schools.
-Third, assimilative associations suggest low SES children, already disadvantaged by prior social skill gaps, are more affected by their school context than are middle to high SES children.
-Thus, while low SES children assimilate to lower social skill environments, richer children appear to gain little advantage from their more conducive environments and appear to function equally well in advantaged and disadvantaged schools.
-Taken together our results support the call for a policy focus that aims to  a) decrease country level variance in social stratification, b) decrease between school heterogeneity is social status, and c) in combination with (a), encourage school selection policies that maximize within school heterogeneity in social status.
-
-\newpage
-
-# References
-
-```{=tex}
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-```
-::: {#refs custom-style="Bibliography"}
-:::
-
-\endgroup
+## References

### 2020-09-10 manuscript.Rmd

--- 2020-09-10 manuscript.Rmd
@@
 Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment.
-For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment <!-- REFERENCE -->.
-Children from low socioeconomic backgrounds (here after low SES children[^1]) appear to enter school with poorer social skills (e.g., <!-- REFERENCE -->Jerrim & Sims, 2019).
-This is concerning as childhood social skills predict a wide variety of later life outcomes (e.g., <!-- REFERENCE -->Jones, Greenberg, & Crowley, 2015).
+For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment [@gutman2013].
+Children from low socioeconomic backgrounds (here after low SES children[^1]) appear to enter school with poorer social skills [e.g., @jerrim2019].
+This is concerning as childhood social skills predict a wide variety of later life outcomes [e.g., @jones2015].
 Inspired by identity economics, our research asks whether the school context in the stratified school system of Australia may exacerbate this issue.
@@
 
-Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable (Rios, Ling, Pugh, Becker, & Bacall, 2020).
-Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior (Durlak, Weissberg, Dymnicki, Taylor, & Schellinger, 2011).
-Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement (Corcoran, Cheung, Kim, & Xie, 2018) and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health (Jones et al., 2015).
-Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire (see Datta Gupta & Simonsen, 2010; Gutman & Schoon, 2013; Jerrim & Sims, 2019).
+Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable [@rios2020].
+Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior [@durlak2011].
+Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement [@corcoran2018] and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health [@jones2015].
+Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire [see @dattagupta2010; @gutman2013; @jerrim2019].
 We use this questionnaire to measure peer problems, conduct problems, and prosocial behavior.
@@
 
-Children's social skills depend in part on the context in which they develop (Jerrim & Sims, 2019).
-There are two major forces in social psychology---assimilation and contrast (Mussweiler, Rüter, & Epstude, 2004)---that may account for how school context affects children's social skill development.
+Children's social skills depend in part on the context in which they develop [@jerrim2019].
+There are two major forces in social psychology---assimilation and contrast [@mussweiler2004]---that may account for how school context affects children's social skill development.
 Contrast processes are in operation when children's perceptions, opinions, or behavior depend on their perceived rank order within their group.
-Assimilation processes are in operation when people's perceptions, opinions, or behaviour depend on reference group norms (Kelly, 1952).
+Assimilation processes are in operation when people's perceptions, opinions, or behaviour depend on reference group norms [@kelley1952].
 
-The theory of *Identity Economics* implies an assimilative effect (Akerlof & Kranton, 2010).
-Identity Economics argues that the social context of the school possesses a gravity that attracts students' identity, values, and behavior toward the prototypical identity, values, and behavior of the school (Akerlof & Kranton, 2002).
+The theory of *Identity Economics* implies an assimilative effect [@akerlof2010].
+Identity Economics argues that the social context of the school possesses a gravity that attracts students' identity, values, and behavior toward the prototypical identity, values, and behavior of the school [@akerlof2002].
 School ethos and norms around behavior incentivize children to act in a manner consistent with these norms.
 School average socioeconomic status could, thus, affect social skills.
-A now large body of research shows that social skills are correlated with socioeconomic status (SES; de Laat, Essink-Bot, van Wassenaer-Leemhuis, & Vrijkotte, 2015; Garratt, Chandola, Purdam, & Wood, 2016; McMunn, Nazroo, Marmot, Boreham, & Goodman, 2001; Rajmil et al., 2013) and, this relationship is present before children enter school (Washbrook & Waldfogel, 2011).
+A now large body of research shows that social skills are correlated with socioeconomic status [SES; @delaat2016; @garratt2017; @mcmunn2001; @rajmil2014] and, this relationship is present before children enter school [@washbrook2011].
 In socially stratified school systems, children of similar SES tend to be schooled together.
@@
 
-Jerrim and Sims (2019) provide one test of the potential assimilative effect of schools on social skills.
+[@jerrim2019] provide one test of the potential assimilative effect of schools on social skills.
 They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school.
@@
 Put another way, the assimilative power of school contexts may not be evenly distributed across the socioeconomic distribution.
-This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized (Gradstein & Justman, 2005).
+This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized [@gradstein2005].
 
@@
 Australia is the context of our research.
-Australia is a useful country to focus on because Australia's Programme for International Student Assessment (PISA) Index of Social Inclusion score is relatively low (OECD, 2015).
-This means that Australian schools are socially stratified (Parker et al., 2018b, 2019).
+Australia is a useful country to focus on because Australia's Programme for International Student Assessment (PISA) Index of Social Inclusion score is relatively low [@oecd2015].
+This means that Australian schools are socially stratified [@parker2018; @parker2019].
 Our major hypothesis is that school average socioeconomic status will have an assimilation effect on association with social skills at Year 3 (age 8) controlling for incoming social skills (age 4), individual socioeconomic status (age 4) and a range of demographic and achievement covariates[^2].
@@
 
-We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills (de Laat, Essink-Bot, van Wassenaer-Leemhuis, & Vrijkotte, 2015; Garratt, Chandola, Purdam, & Wood, 2016; McMunn, Nazroo, Marmot, Boreham, & Goodman, 2001; Rajmil et al., 2013).
-Given Australian schools are socially stratified (OECD, 2015; Parker et al., 2018), we also expect low SES children to attend more disadvantaged schools on average.
+We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills [@delaat2016; @garratt2017; @mcmunn2001; @rajmil2014].
+Given Australian schools are socially stratified [@oecd2015; @parker2018], we also expect low SES children to attend more disadvantaged schools on average.
 
@@
 LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004.
-Both cohorts of children have been followed every two years since (AIFS, 2015).
+Both cohorts of children have been followed every two years since [@aifs2015].
 LSAC is a complex survey with a stratified sampling design (that we account for in all analyses) and includes linked administration records for student performance in a standardized national numeracy test.
@@
 
-Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (Goodman, 1997).
-We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child’s *parent* and the child’s *teacher*.
-The SDQ asks respondents to rate a child’s behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
-Questions ask about the child’s peer problems (“Rather solitary, tends to play alone”), conduct problems (“Often fights with other children or bullies them”), and prosociality (“Is kind to younger children”).
+Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ [@goodman1997].
+We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child's *parent* and the child's *teacher*.
+The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
+Questions ask about the child's peer problems ("Rather solitary, tends to play alone"), conduct problems ("Often fights with other children or bullies them"), and prosociality ("Is kind to younger children").
 We used the total scores for these scales as developed by the LSAC administrators.
@@
 Instead, we would want items that are not too redundant with each other.
-For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index (Diamantopoulos, Riefler, & Roth, 2008; Thongrattana, 2010).
+For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index [@diamantopoulos2008; @thongrattana2010].
 In our case no VIF for any item for any social skills construct was higher than 2.
-Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct (Thongrattana, 2010).
+Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct [@thongrattana2010].
 This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report.
@@
 
-*Individual SES* SES was measured using the Socioeconomic Position (SEP) index constructed by the LSAC survey organizers (Baker, Sipthorp, & Edwards, 2017).
+*Individual SES* was measured using the Socioeconomic Position (SEP) index constructed by the LSAC survey organizers [@baker2017].
 The SEP index is constructed from parent reported standardized weekly income, years of education, and ANU4 occupational prestige derived from the Australian Standard Classification of Occupations.
@@
 
-Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program – Literacy and Numeracy (NAPLAN) test results.
+Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program -- Literacy and Numeracy (NAPLAN) test results.
 The relevant NAPLAN tests for this study are given to all eligible children in the country in Year 3 (age 8).
@@
 These high-stakes tests have a mean of 500 and a standard deviation of 100.
-For verbal ability we used the Peabody Picture Vocabulary test (Dunn & Dunn, 2007) given to participants at age 4 (i.e., at or just before school commencement).
-Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender—all measured at age 4.
+For verbal ability we used the Peabody Picture Vocabulary test [@dunn2007] given to participants at age 4 (i.e., at or just before school commencement).
+Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4.
 Both numeracy achievement and verbal ability were z-scored in the analysis.
@@
 There was a relatively small amount of missing data for most variables.
-However, given the data were longitudinal and included both linked administrative records and data collected from both the child’s teacher and parent, missing data was inevitable.
+However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable.
 Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample.
@@
 All other variables had 6% missing data or less.
-To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, Blackwell, 2011).
+To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
 These 10 imputations were used in all analyses.
@@
 LSAC strata was included as a fixed effect to account for the complex survey structure of the data.
-Analysis was done in the Zelig package in R (Imia, King, & Lau, 2008).
+Analysis was done in the Zelig package in R [@imai2008].
 
 All models in the paper that predicted social skills were estimated using Tobit models.
-Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see Kleiber & Zeileis, 2008).
+Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see [see @kleiber2008].
 As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
@@
 
-Research on social skills has repeatedly shown that there is a socioeconomic status gradient to social skills (see Datta Gupta & Simonsen, 2010; Gutman & Schoon, 2013; Jerrim & Sims, 2019).
-Yet little research in this area has considered the potential influence of school context on social skill development (cf.
-Jerrim & Sims, 2019).
+Research on social skills has repeatedly shown that there is a socioeconomic status gradient to social skills [see @dattagupta2010; @gutman2013; @jerrim2019].
+Yet little research in this area has considered the potential influence of school context on social skill development [*cf.* @jerrim2019].
 This is a lacuna that our research sought to fill.
@@
 
-For both teacher and parent reports we found that the association of school context with social skills depended on the individual child’s own socioeconomic background.
+For both teacher and parent reports we found that the association of school context with social skills depended on the individual child's own socioeconomic background.
 This was the case for all outcomes except for parent reported prosocial behavior.
@@
 
-A greater focus on social skills as explanations for socioeconomic gaps in educational attainment has been an important step forward in inequality research (see Heckman, 2006).
+A greater focus on social skills as explanations for socioeconomic gaps in educational attainment has been an important step forward in inequality research [see @heckman2006].
 Now that research and theory has illuminated the importance of such skills, research needs to consider the conditions under which they develop.
-Previous economic theory has emphasized the role that schools play as a context for the development of non-academic factors like social skills and claimed this as one of the ways in which intergenerational inequality is transmitted (Bowles & Gintis, 1976).
-Such theory has tended to emphasize the role of teachers and systems, relegating  students’ fellow classmates to a more secondary role.
-In contrast, psychology research has tended to emphasize the role of frames-of-reference with particular emphasize given to the role of a child’s peers as providing a standard against which a child might assimilate to or contrast against (Mussweiler, Rüter, & Epstude, 2004).
+Previous economic theory has emphasized the role that schools play as a context for the development of non-academic factors like social skills and claimed this as one of the ways in which intergenerational inequality is transmitted [@bowles1976].
+Such theory has tended to emphasize the role of teachers and systems, relegating students' fellow classmates to a more secondary role.
+In contrast, psychology research has tended to emphasize the role of frames-of-reference with particular emphasize given to the role of a child's peers as providing a standard against which a child might assimilate to or contrast against [@mussweiler2004].
 It seems likely that both economic and psychology theory is right to some extent but the relative contribution to school context effects of system, teacher, and peers is unclear.
@@
 The problem is that in countries like Australia low SES children are relatively rarely enrolled in advantaged schools.
-As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA (OECD, 2015).
-This means that Australian schools tend to be homogenous in terms of their student composition (see Parker, Sanders, & Guo, 2019).
+As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA [@oecd2015].
+This means that Australian schools tend to be homogenous in terms of their student composition [see @parker2019].
 As our results show low SES children tend to attend disadvantaged schools and their social skills appear to suffer as a result.
-What our findings suggest is that children’s social skills, particularly low SES children, acclimatize toward the average of the school they are in.
+What our findings suggest is that children's social skills, particularly low SES children, acclimatize toward the average of the school they are in.
 Consistent with the literature we found that low SES children start school with lower social skills.
@@
 
-This counterfactual may suggest that school choice, in which low SES children receive vouchers or similar, may be beneficial (see Friedman, 1962).
+This counterfactual may suggest that school choice, in which low SES children receive vouchers or similar, may be beneficial [see @friedman1962].
 Strategies like school vouchers to attend magnet schools could be a powerful policy lever to overcome socioeconomic gaps in social skills.
@@
 Yet this policy requires there to be few barriers, whether psychological or otherwise, to parents using vouchers to select the best school that matches the needs of their child.
-But this does not seem to be the case (Gradstein & Justman, 2005).
-Indeed, empirical evidence suggests that school choice tends to exacerbate inequality (e.g., Saporito, 2003).
+But this does not seem to be the case [@gradstein2005].
+Indeed, empirical evidence suggests that school choice tends to exacerbate inequality [e.g., @saporito2003].
 
 Parker et al.
-(2016, 2018a, 2018b, 2019) have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
+[-@parker2016; -@parker2018 ; -@parker_information_2018; -@parker2019] have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
 They argue that empirical evidence suggests policy should encourage school selection policies that maximize within school heterogeneity.
-This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents’ rights to choose.
+This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents' rights to choose.
 However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a disadvantaged school.
@@
 There are strong arguments and good empirical support on both sides of this debate, suggesting that we are far from a settled position on the matter.
-At least for the current context in Australia where social stratification is moderately high and where the school system seems to ensure school choice is more clearly an option for the rich than the poor (Parker et al., 2019), our results are troubling.
+At least for the current context in Australia where social stratification is moderately high and where the school system seems to ensure school choice is more clearly an option for the rich than the poor [@parker2019], our results are troubling.
 In particular, they suggest low SES children face a triple bind.
@@
 Nevertheless, there are limitations.
-Our aim was to try to build a model from high-quality data that could assist us in making as close to an *all else being equal* comparisons as possible (Angrist & Pischke, 2009).
+Our aim was to try to build a model from high-quality data that could assist us in making as close to an *all else being equal* comparisons as possible [@angrist2008].
 It is for this reason we have cautiously used causal language in relation to our claims.
 But our claim to causality would have been stronger had they been backed by an experimental design.
-While the Move to Opportunity program in the US suggest experiments where low SES children are randomly assigned to richer schools (or at least randomly assigned to areas with richer schools) are possible (see de Souza Briggs, Popkin & Goering, 2010), it is hard to imagine a situation in which richer children could be randomly assigned to more disadvantaged schools.
+While the Move to Opportunity program in the US suggest experiments where low SES children are randomly assigned to richer schools (or at least randomly assigned to areas with richer schools) are possible [see @desouzabriggs2010], it is hard to imagine a situation in which richer children could be randomly assigned to more disadvantaged schools.
 Given the non-linearity in effects we observed this would be a serious limitation of an experimental design.
@@
 Finally, we were not able to identify and compare the relative impact of different mechanisms that may explain the influence of school average socioeconomic status on social skills.
-As educational psychologists our study is framed in relation to assimilation effects in response to children’s peer frames-of-reference.
-Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see Bowls & Gintis, 2001).
+As educational psychologists our study is framed in relation to assimilation effects in response to children's peer frames-of-reference.
+Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see <!--# @PHIL this citation wasn't in reference list -->Bowls & Gintis, 2001).
 Identifying and comparing these mechanisms is an important future direction for research.

### 2020-09-10 manuscript.tex
References not working in html.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={School Socioeconomic Status Context and Social Skills in Children},
  pdfauthor={Philip Parker1 \& Ernst-August Doelle1,2},
  pdflang={en-EN},
  pdfkeywords={social skills; assimilation effects; socioeconomic status; school context},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{School SES and Social Skills}
\keywords{social skills; assimilation effects; socioeconomic status; school context\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{School Socioeconomic Status Context and Social Skills in Children}
\author{Philip Parker\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
\date{}

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
Enter author note here.

The authors made the following contributions. Philip Parker: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Philip Parker, 33 Berry St, North Sydney, NSW, 2060, Australia. E-mail: \href{mailto:philip.parker@acu.edu.au}{\nolinkurl{philip.parker@acu.edu.au}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute for Positive Psychology and Education, Australian Catholic University\\\textsuperscript{2} Konstanz Business School}

\abstract{
As researchers and policymakers try to understand the consequences of SES for educational and occupational attainment, they increasingly consider social skills to be critical. Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development. In a representative longitudinal sample of 5361 Australian children, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers. All models controlled for prior social skills at age 4, and a wide range of covariates. We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems. We found a consistent interaction between individual and school average socioeconomic status that suggested assimilation effects were only present for children from low socioeconomic backgrounds. We found that children from low socioeconomic backgrounds enter school with lower average social skills. They are more likely to be enrolled in more disadvantaged schools. Finally, the social skills of children from low socioeconomic backgrounds worsen in low SES contexts. Taking the evidence together, social stratified schools may harm children from low SES backgrounds while not benefiting children from high SES backgrounds.
}

\begin{document}
\maketitle

After decades of research and intervention aimed at reducing educational and occupational inequality, inequalities persist (Reardon, 2011). In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman. (Heckman, 2006) argues that research and intervention on reducing educational inequality have focused too narrowly on ``cognitive'' abilities. Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment. For example, social skills are a potentially powerful explanation for socioeconomic status gaps in educational and occupational attainment (Gutman \& Schoon, 2013). Children from low socioeconomic backgrounds (here after low SES children\footnote{We acknowledge that children themselves do not have a socioeconomic status. Rather they are raised in an environment which is shaped by their parents' socioeconomic status (SES). However, for brevity we refer to children as low SES or high SES hereafter. Likewise, in place of low school average SES and high school average SES we refer to disadvantaged and advantaged schools respectively.}) appear to enter school with poorer social skills (e.g., Jerrim \& Sims, 2019). This is concerning as childhood social skills predict a wide variety of later life outcomes (e.g., Jones, Greenberg, \& Crowley, 2015). Inspired by identity economics, our research asks whether the school context in the stratified school system of Australia may exacerbate this issue. In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior. We consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates in a representative longitudinal sample of Australian children.

\hypertarget{social-skills}{%
\subsection{Social Skills}\label{social-skills}}

Of the skills that employers are looking for when hiring candidates, social skills are some of the most desirable (Rios, Ling, Pugh, Becker, \& Bacall, 2020). Social skills also seem to be a viable target for intervention with meta-analysis showing social and emotional learning programs have moderate effects on improving key social skills like reducing conduct problems and improving social-emotional learning skills and prosocial behavior (Durlak, Weissberg, Dymnicki, Taylor, \& Schellinger, 2011). Social skills appear to be worthwhile targets from a societal perspective given their relationship to academic achievement (Corcoran, Cheung, Kim, \& Xie, 2018) and their role in predicting adult employment, crime, public assistance, substance abuse, and mental health (Jones et al., 2015). Among young children, it is increasingly common to focus on social skills as measured by the Strengths and Difficulties Questionnaire (see Datta Gupta \& Simonsen, 2010; Gutman \& Schoon, 2013; Jerrim \& Sims, 2019). We use this questionnaire to measure peer problems, conduct problems, and prosocial behavior. We model social skills using both parent and teacher reports.

\hypertarget{contrast-processes-and-identity-economics}{%
\subsection{Contrast Processes and Identity Economics}\label{contrast-processes-and-identity-economics}}

Children's social skills depend in part on the context in which they develop (Jerrim \& Sims, 2019). There are two major forces in social psychology---assimilation and contrast (Mussweiler, Rüter, \& Epstude, 2004)---that may account for how school context affects children's social skill development. Contrast processes are in operation when children's perceptions, opinions, or behavior depend on their perceived rank order within their group. Assimilation processes are in operation when people's perceptions, opinions, or behaviour depend on reference group norms (Kelley, 1952).

The theory of \emph{Identity Economics} implies an assimilative effect (Akerlof \& Kranton, 2010). Identity Economics argues that the social context of the school possesses a gravity that attracts students' identity, values, and behavior toward the prototypical identity, values, and behavior of the school (Akerlof \& Kranton, 2002). School ethos and norms around behavior incentivize children to act in a manner consistent with these norms. School average socioeconomic status could, thus, affect social skills. A now large body of research shows that social skills are correlated with socioeconomic status (SES; de Laat, Essink-Bot, van Wassenaer-Leemhuis, \& Vrijkotte, 2016; Garratt, Chandola, Purdam, \& Wood, 2017; McMunn, Nazroo, Marmot, Boreham, \& Goodman, 2001; Rajmil et al., 2014) and, this relationship is present before children enter school (Washbrook \& Waldfogel, 2011). In socially stratified school systems, children of similar SES tend to be schooled together. Thus, if children assimilate to the school context they are in, as Identity Economics would suggest, we would expect school average socioeconomic status to influence the development of social skills.

(Jerrim \& Sims, 2019) provide one test of the potential assimilative effect of schools on social skills. They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school. They also found some assimilative effects of school average SES predicting social skills using the total score of the strengths and difficulties questionnaire (SDQ). The assimilation effects they found, however, were weak. This may be because the authors used the SDQ total score---which includes a range of social and emotional behaviors---rather than focusing on specific social skill components of the SDQ. Further, assimilative effects may be non-linear. Put another way, the assimilative power of school contexts may not be evenly distributed across the socioeconomic distribution. This is an important consideration that research has not hitherto considered, though differential assimilation across the socioeconomic status gradient has been theorized (Gradstein \& Justman, 2005).

\hypertarget{current-study}{%
\section{Current Study}\label{current-study}}

Australia is the context of our research. Australia is a useful country to focus on because Australia's Programme for International Student Assessment (PISA) Index of Social Inclusion score is relatively low (OECD, 2015). This means that Australian schools are socially stratified (Parker, Guo, \& Sanders, 2019; Philip D. Parker et al., 2018). Our major hypothesis is that school average socioeconomic status will have an assimilation effect on association with social skills at Year 3 (age 8) controlling for incoming social skills (age 4), individual socioeconomic status (age 4) and a range of demographic and achievement covariates\footnote{This assimilation effect will be indicated by a negative coefficient for peer problems and conduct problems regressed on school average socioeconomic status and a positive coefficient for prosocial behavior regressed on school average socioeconomic status. This is because of the way in which these factors are valanced (high scores indicate high levels of prosocial behavior, but also high levels of peer and conduct problems).}. We also anticipate that there will be differential associations of school average socioeconomic status on social skills across the socioeconomic status gradient.

We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills (de Laat et al., 2016; Garratt et al., 2017; McMunn et al., 2001; Rajmil et al., 2014). Given Australian schools are socially stratified (OECD, 2015; Parker et al., 2018), we also expect low SES children to attend more disadvantaged schools on average.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants-and-study-design}{%
\subsection{Participants and Study Design}\label{participants-and-study-design}}

We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC). LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004. Both cohorts of children have been followed every two years since (AIFS, 2015). LSAC is a complex survey with a stratified sampling design (that we account for in all analyses) and includes linked administration records for student performance in a standardized national numeracy test. We also use government collected data on school average socioeconomic status. Many children in the sample were the only child surveyed (27\%) or were one of two children surveyed (36\%) in their school. By using administrative records, we gain school average socioeconomic status based on the child's \(\text{\underline{whole}}\) school grade from a high-quality source, thus avoiding sampling bias present in most school context research. We excluded participants who were not in school (e.g., were home schooled), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability). Together our total sample was 5,361 (52.1\% male) children in Year 3 at school (\textasciitilde age 8). In the vast majority of cases (98\%), parent data came from mothers. Weights were provided by the survey organizers to ensure the sample represents the Australian population. This includes weights provided that aim to account for attrition.

\hypertarget{measures}{%
\subsection{Measures}\label{measures}}

\hypertarget{social-skills-at-age-4-and-8}{%
\subsubsection{Social skills at age 4 and 8}\label{social-skills-at-age-4-and-8}}

Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (\emph{Strengths and difficulties questionnaire}, 1997). We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}. The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true). Questions ask about the child's peer problems (``Rather solitary, tends to play alone''), conduct problems (``Often fights with other children or bullies them''), and prosociality (``Is kind to younger children''). We used the total scores for these scales as developed by the LSAC administrators. Scores ranged from 0 to 10. These scores were heavily left censored (see Figures 2-3) for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher. prosocial behaviors were heavily right censored. Parents and teachers had moderate agreement for children at age 8 with correlations ranging from r = .40 for peer and r = .39 for conduct problems to r = .28 for prosocial behavior.

The SDQ items measure a collection of associated behaviours and as such we treat them as index or a formative variable. In this case common reliability metrics do not make much sense; nor do reflective models like confirmatory factor analysis. Instead, we would want items that are not too redundant with each other. For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index (Diamantopoulos, Riefler, \& Roth, 2008; Thongrattana, 2010). In our case no VIF for any item for any social skills construct was higher than 2. Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct (Thongrattana, 2010). This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report. Parallel analysis suggested that all social skills constructs could be represented by a single principal component (but in all cases more than one reflective factor). The only exception was age 4 peer problems were parallel analysis suggested two components but the eigenvalue on the second component was barely greater than the simulated data suggesting a single index was likely still appropriate.

\hypertarget{socioeconomic-status}{%
\subsubsection{Socioeconomic Status}\label{socioeconomic-status}}

\emph{Individual SES} was measured using the Socioeconomic Position (SEP) index constructed by the LSAC survey organizers (Baker, Sipthorp, \& Edwards, 2017). The SEP index is constructed from parent reported standardized weekly income, years of education, and ANU4 occupational prestige derived from the Australian Standard Classification of Occupations. The SEP has a mean of zero and a standard deviation of one.

\emph{School average SES} was measured using the Index of Community Socioeconomic Advantage (ICSEA) that the Australian government uses to assess the relative advantage of schools for the purpose of funding allocation and policy. This measure was taken from government administration records. ICSEA has a mean of 1000 and a standard deviation of 100. We z-scored this variable for analysis.

\hypertarget{covariates}{%
\subsubsection{Covariates}\label{covariates}}

Numeracy for the child and for their complete school grade was taken from administration records of National Assessment Program -- Literacy and Numeracy (NAPLAN) test results. The relevant NAPLAN tests for this study are given to all eligible children in the country in Year 3 (age 8). The tests are scaled so they are comparable across age cohort and across year grade. These high-stakes tests have a mean of 500 and a standard deviation of 100. For verbal ability we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement). Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4. Both numeracy achievement and verbal ability were z-scored in the analysis.

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

There was a relatively small amount of missing data for most variables. However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable. Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample. Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14\%. All other variables had 6\% missing data or less. To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, \& Blackwell, 2011). These 10 imputations were used in all analyses. All models used cluster robust standard errors to adjust standard errors for school clustering. LSAC strata was included as a fixed effect to account for the complex survey structure of the data. Analysis was done in the Zelig package in R (Imai, King, \& Lau, 2008).

All models in the paper that predicted social skills were estimated using Tobit models. Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see (see Kleiber \& Zeileis, 2008). As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models. Johnson-Neyman regions of significance were used to investigate significant interactions. We also attempted to run random-intercept models using for school id both maximum likelihood estimation and MCMC sampling. In both cases models did not converge likely due to the unique nature of the data (i.e., many participants were the single representative of their school).

Contextual effects like those reported here can be confusing to interpret. This is because positive effects of contextual variables can be incorrectly interpreted as positive in implication (or vice-versa for negative effects). This issue is compounded when interactions involving contextual effects are included in the model. Thus, we also provide counterfactual estimates of expected score differences for different levels of student SES and school average SES when the interaction between these variables is significant. Because of the number of covariates (35 in total because many of the covariates, such as strata, were categorical and thus were included as a series of dummy variables), we only report results for the main analysis variables.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{low-ses-children-start-school-with-poorer-social-skills-and-enroll-in-more-disadvantaged-schools}{%
\subsection{Low SES Children Start School with Poorer Social Skills and Enroll in More Disadvantaged Schools}\label{low-ses-children-start-school-with-poorer-social-skills-and-enroll-in-more-disadvantaged-schools}}

In order to provide a context for the main analyses we first aimed to show a) that richer children enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools. Our findings show that high SES children do enter school with better social skills. Student SES is correlated with less peer (B = -.284 {[}-.345, -.222{]}) and conduct (B = -.371 {[}-.438, -.304{]}) problems. They also had slightly higher prosocial behavior (B = .074 {[}.009, .139 {]}). Further, student SES was strongly related to the school average SES of the school they attended (r = .480 {[}.458, .499{]}; see Figure 1).

\hypertarget{school-average-ses-predicts-social-skills-controlling-for-age-4-ses-and-social-skills}{%
\subsection{School Average SES Predicts Social Skills Controlling for Age 4 SES and Social Skills}\label{school-average-ses-predicts-social-skills-controlling-for-age-4-ses-and-social-skills}}

We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4. Results are presented under the Model 1 column in Tables 1-3. Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3). School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills. School average SES also negatively predicted peer problems but only for parent reports.

\hypertarget{school-average-ses-predicts-social-skills-mainly-in-low-ses-children}{%
\subsection{School Average SES Predicts Social Skills Mainly in Low SES Children}\label{school-average-ses-predicts-social-skills-mainly-in-low-ses-children}}

The predictive association of school average SES on social skills was not consistent across the SES gradient (see Model 2 results reported in Tables 1-3). Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior. Regions of significance for these interactions can be found in Figures 2-4. Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4. These results show that low SES children appear to be most influenced by school average SES. Middle to high SES children appeared to have largely similar scores regardless of the school they were in (see Figures 2-4). For example, both parents and teachers report that low SES children had approximately half a point more peer problems in more SES disadvantaged schools than in more SES advantaged schools (see Table 4). High SES children had almost no difference in peer problems regardless of the school they were in.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Research on social skills has repeatedly shown that there is a socioeconomic status gradient to social skills (see Datta Gupta \& Simonsen, 2010; Gutman \& Schoon, 2013; Jerrim \& Sims, 2019). Yet little research in this area has considered the potential influence of school context on social skill development (\emph{cf.} Jerrim \& Sims, 2019). This is a lacuna that our research sought to fill. Our research considered the association of school socioeconomic context with social skills in early elementary school. We used both parent and teacher reported social skills, finding surprisingly consistent effect sizes regardless of reporting source despite their relatively modest agreement. Importantly, our research used a number of critical controls as well as prior social skills that helped us better identify the effect of school average socioeconomic status.

For both teacher and parent reports we found that the association of school context with social skills depended on the individual child's own socioeconomic background. This was the case for all outcomes except for parent reported prosocial behavior. It is worth emphasizing that the nature of these interactions was consistent for all outcomes. Namely, that middle to high SES children were largely unaffected by school socioeconomic context while the children with the lowest socioeconomic status experienced the largest effect.

\hypertarget{school-context-theory}{%
\subsection{School Context Theory}\label{school-context-theory}}

A greater focus on social skills as explanations for socioeconomic gaps in educational attainment has been an important step forward in inequality research (see Heckman, 2006). Now that research and theory has illuminated the importance of such skills, research needs to consider the conditions under which they develop. Previous economic theory has emphasized the role that schools play as a context for the development of non-academic factors like social skills and claimed this as one of the ways in which intergenerational inequality is transmitted (Bowles \& Gintis, 1976). Such theory has tended to emphasize the role of teachers and systems, relegating students' fellow classmates to a more secondary role. In contrast, psychology research has tended to emphasize the role of frames-of-reference with particular emphasize given to the role of a child's peers as providing a standard against which a child might assimilate to or contrast against (Mussweiler et al., 2004). It seems likely that both economic and psychology theory is right to some extent but the relative contribution to school context effects of system, teacher, and peers is unclear. Our research shows that school context effects are significant and mainly seem to effect low SES children. Future research is needed to determine what the most important mechanisms are that explain this effect.

\hypertarget{school-context-and-assimilation-effects}{%
\subsection{School Context and Assimilation Effects}\label{school-context-and-assimilation-effects}}

For children with SES status below the mean, school socioeconomic context had statistically significant associations with social skills in Year 3 (controlling for social skills at age 4). The implications of this are mixed. Our results do suggest that a low SES child who is enrolled in an advantaged school should expect to have similar levels of social skills as their high SES peers. The problem is that in countries like Australia low SES children are relatively rarely enrolled in advantaged schools. As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA (OECD, 2015). This means that Australian schools tend to be homogenous in terms of their student composition (see Parker et al., 2019). As our results show low SES children tend to attend disadvantaged schools and their social skills appear to suffer as a result. What our findings suggest is that children's social skills, particularly low SES children, acclimatize toward the average of the school they are in. Consistent with the literature we found that low SES children start school with lower social skills. In a country like Australia, whose schools are socially stratified, low SES children will tend to be schooled in more disadvantaged schools (see Figure 1). The net effect in such a system is that low SES children will tend to have their already lower social skills depressed by the school climate they are most likely to find themselves in.

This counterfactual may suggest that school choice, in which low SES children receive vouchers or similar, may be beneficial (see Friedman, 1962). Strategies like school vouchers to attend magnet schools could be a powerful policy lever to overcome socioeconomic gaps in social skills. This approach tackles the problem of assimilative contextual effects via market-based systems. Yet this policy requires there to be few barriers, whether psychological or otherwise, to parents using vouchers to select the best school that matches the needs of their child. But this does not seem to be the case (Gradstein \& Justman, 2005). Indeed, empirical evidence suggests that school choice tends to exacerbate inequality (e.g., Saporito, 2003).

Parker et al.~(2019, 2018 ; 2016; 2018) have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children. They argue that empirical evidence suggests policy should encourage school selection policies that maximize within school heterogeneity. This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents' rights to choose. However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a disadvantaged school.

There are strong arguments and good empirical support on both sides of this debate, suggesting that we are far from a settled position on the matter. At least for the current context in Australia where social stratification is moderately high and where the school system seems to ensure school choice is more clearly an option for the rich than the poor (Parker et al., 2019), our results are troubling. In particular, they suggest low SES children face a triple bind. First, they enter school with lower social skills. Second, they enter schools where assimilation effects are likely to further dampen social skill development. Third, they appear particularly vulnerable to their school context in this respect.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

There are many strengths to this study. Most notably, the use of longitudinal data that allowed for us to control for incoming social skills and government administrative data that provided access to complete and high-quality data for school average socioeconomic status at the school level. Further the use of LSAC data meant that we were able to control for a number of potential confounding variables drawn from a longitudinal representative sample of Australian children. Nevertheless, there are limitations. Our aim was to try to build a model from high-quality data that could assist us in making as close to an \emph{all else being equal} comparisons as possible (Angrist \& Pischke, 2008). It is for this reason we have cautiously used causal language in relation to our claims. But our claim to causality would have been stronger had they been backed by an experimental design. While the Move to Opportunity program in the US suggest experiments where low SES children are randomly assigned to richer schools (or at least randomly assigned to areas with richer schools) are possible (see de Souza Briggs, Popkin, \& Goering, 2010), it is hard to imagine a situation in which richer children could be randomly assigned to more disadvantaged schools. Given the non-linearity in effects we observed this would be a serious limitation of an experimental design.

Finally, we were not able to identify and compare the relative impact of different mechanisms that may explain the influence of school average socioeconomic status on social skills. As educational psychologists our study is framed in relation to assimilation effects in response to children's peer frames-of-reference. Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see Bowls \& Gintis, 2001). Identifying and comparing these mechanisms is an important future direction for research.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The influence of school average socioeconomic status on social skills represent the triple disadvantage that low SES children can face in social stratified school systems. First, low SES children are more likely to start school with lower social skills than their high SES peers. Second, because the school system is stratified by socioeconomic status, low SES children are likely to enroll in more disadvantaged schools. Third, assimilative associations suggest low SES children, already disadvantaged by prior social skill gaps, are more affected by their school context than are middle to high SES children. Thus, while low SES children assimilate to lower social skill environments, richer children appear to gain little advantage from their more conducive environments and appear to function equally well in advantaged and disadvantaged schools. Taken together our results support the call for a policy focus that aims to a) decrease country level variance in social stratification, b) decrease between school heterogeneity is social status, and c) in combination with (a), encourage school selection policies that maximize within school heterogeneity in social status.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-aifs2015}{}%
AIFS, undefined. (2015). \emph{Longitudinal study of australian children data user guide}.

\leavevmode\hypertarget{ref-akerlof2002}{}%
Akerlof, G., \& Kranton, R. (2002). Identity and schooling: Some lessons for the economics of education. \emph{Journal of Economic Literature}, \emph{40}(4), 1167--1201. \url{https://doi.org/10.1257/.40.4.1167}

\leavevmode\hypertarget{ref-akerlof2010}{}%
Akerlof, G., \& Kranton, R. (2010). \emph{Identity economics: How our identities shape our work, wages, and well-being}. Princeton: Princeton University Press. \url{https://doi.org/10.1515/9781400834181}

\leavevmode\hypertarget{ref-angrist2008}{}%
Angrist, J., \& Pischke, J. (2008). \emph{Mostly harmless econometrics: An empiricist's companion}. Princeton university press.

\leavevmode\hypertarget{ref-baker2017}{}%
Baker, K., Sipthorp, M., \& Edwards, B. (2017). \emph{A longitudinal measure of socioeconomic position in lsac}. Retrieved from \url{https://growingupinaustralia.gov.au/sites/default/files/tp18.pdf}

\leavevmode\hypertarget{ref-bowles1976}{}%
Bowles, S., \& Gintis, H. (1976). \emph{Schooling in capitalist america: Educational reform and the contradictions of economic life}. Routledge \& Kegan Paul.

\leavevmode\hypertarget{ref-corcoran2018}{}%
Corcoran, R., Cheung, A., Kim, E., \& Xie, C. (2018). Effective universal school-based social and emotional learning programs for improving academic achievement: A systematic review and meta-analysis of 50 years of research. \emph{Educational Research Review}, \emph{25}, 56--72. \url{https://doi.org/10.1016/j.edurev.2017.12.001}

\leavevmode\hypertarget{ref-dattagupta2010}{}%
Datta Gupta, N., \& Simonsen, M. (2010). Non-cognitive child outcomes and universal high quality child care. \emph{Journal of Public Economics}, \emph{94}(1-2), 30--43. \url{https://doi.org/10.1016/j.jpubeco.2009.10.001}

\leavevmode\hypertarget{ref-delaat2016}{}%
de Laat, S., Essink-Bot, M., van Wassenaer-Leemhuis, A., \& Vrijkotte, T. (2016). Effect of socioeconomic status on psychosocial problems in 5- to 6-year-old preterm- and term-born children: The abcd study. \emph{European Child \& Adolescent Psychiatry}, \emph{25}(7), 757--767. \url{https://doi.org/10.1007/s00787-015-0791-4}

\leavevmode\hypertarget{ref-desouzabriggs2010}{}%
de Souza Briggs, X., Popkin, S., \& Goering, J. (2010). \emph{Moving to opportunity: The story of an american experiment to fight ghetto poverty}. Oxford University Press.

\leavevmode\hypertarget{ref-diamantopoulos2008}{}%
Diamantopoulos, A., Riefler, P., \& Roth, K. (2008). Advancing formative measurement models. \emph{Journal of Business Research}, \emph{61}(12), 1203--1218. \url{https://doi.org/10.1016/j.jbusres.2008.01.009}

\leavevmode\hypertarget{ref-durlak2011}{}%
Durlak, J., Weissberg, R., Dymnicki, A., Taylor, R., \& Schellinger, K. (2011). The impact of enhancing students' social and emotional learning: A meta-analysis of school-based universal interventions: Social and emotional learning. \emph{Child Development}, \emph{82}(1), 405--432. \url{https://doi.org/10.1111/j.1467-8624.2010.01564.x}

\leavevmode\hypertarget{ref-friedman1962}{}%
Friedman, M. (1962). \emph{Capitalism and freedom}. University of Chicago press.

\leavevmode\hypertarget{ref-garratt2017}{}%
Garratt, E., Chandola, T., Purdam, K., \& Wood, A. (2017). Income and social rank influence uk children's behavioral problems: A longitudinal analysis. \emph{Child Development}, \emph{88}(4), 1302--1320. \url{https://doi.org/10.1111/cdev.12649}

\leavevmode\hypertarget{ref-gradstein2005}{}%
Gradstein, M., \& Justman, M. (2005). The melting pot and school choice. \emph{Journal of Public Economics}, \emph{89}(5-6), 871--896. \url{https://doi.org/10.1016/j.jpubeco.2004.05.007}

\leavevmode\hypertarget{ref-gutman2013}{}%
Gutman, L., \& Schoon, I. (2013). \emph{The impact of non-cognitive skills on outcomes for young people} (p. 2019). Retrieved from \url{https://educationendowmentfoundation.org.uk/public/files/Publications/EEF_Lit_Review_Non-CognitiveSkills.pdf}

\leavevmode\hypertarget{ref-heckman2006}{}%
Heckman, J. (2006). Skill formation and the economics of investing in disadvantaged children. \emph{Science}, \emph{312}(5782), 1900--1902. \url{https://doi.org/10.1126/science.1128898}

\leavevmode\hypertarget{ref-honaker2011}{}%
Honaker, J., King, G., \& Blackwell, M. (2011). Amelia ii: A program for missing data. \emph{Journal of Statistical Software}, \emph{45}(7), 1--47.

\leavevmode\hypertarget{ref-imai2008}{}%
Imai, K., King, G., \& Lau, O. (2008). Toward a common framework for statistical analysis and development. \emph{Journal of Computational and Graphical Statistics}, \emph{17}(4), 892--913.

\leavevmode\hypertarget{ref-jerrim2019}{}%
Jerrim, J., \& Sims, S. (2019). How do academically selective school systems affect pupils' social-emotional competencies? New evidence from the millennium cohort study. \emph{American Educational Research Journal}, \emph{56}(5), 1769--1799. \url{https://doi.org/10.3102/0002831219830965}

\leavevmode\hypertarget{ref-jones2015}{}%
Jones, D., Greenberg, M., \& Crowley, M. (2015). Early social-emotional functioning and public health: The relationship between kindergarten social competence and future wellness. \emph{American Journal of Public Health}, \emph{105}(11), 2283--2290. \url{https://doi.org/10.2105/AJPH.2015.302630}

\leavevmode\hypertarget{ref-kelley1952}{}%
Kelley, H. (1952). Two functions of reference groups. In G. Swanson, T. Newcomb, \& E. Hartley (Eds.), \emph{Readings in social psychology} (Vol. 2, pp. 410--414). New York, NY: Henry Holt and Company.

\leavevmode\hypertarget{ref-kleiber2008}{}%
Kleiber, C., \& Zeileis, A. (2008). \emph{Applied econometrics with r}. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-mcmunn2001}{}%
McMunn, A., Nazroo, J., Marmot, M., Boreham, R., \& Goodman, R. (2001). Children's emotional and behavioural well-being and the family environment: Findings from the health survey for england. \emph{Social Science \& Medicine}, \emph{53}(4), 423--440. \url{https://doi.org/10.1016/S0277-9536(00)00346-4}

\leavevmode\hypertarget{ref-mussweiler2004}{}%
Mussweiler, T., Rüter, K., \& Epstude, K. (2004). The ups and downs of social comparison: Mechanisms of assimilation and contrast. \emph{Journal of Personality and Social Psychology}, \emph{87}(6), 832--844. \url{https://doi.org/10.1037/0022-3514.87.6.832}

\leavevmode\hypertarget{ref-oecd2015}{}%
OECD, undefined. (2015). \emph{How schools have changed over the past decade}. Retrieved from \url{http://www.oecd.org/pisa/pisaproducts/pisainfocus/pisa-in-focus-n52-(eng)-final.pdf}

\leavevmode\hypertarget{ref-parker2019}{}%
Parker, P., Guo, J., \& Sanders, T. (2019). Socioeconomic inequality and student outcomes in australia. In L. Volante, S. Schnepf, J. Jerrim, \& D. Klinger (Eds.), \emph{Socioeconomic inequality and student outcomes: Cross-national trends, policies, and practices} (pp. 189--204). Singapore: Springer Singapore. \url{https://doi.org/10.1007/978-981-13-9863-6_11}

\leavevmode\hypertarget{ref-parker2016}{}%
Parker, P., Jerrim, J., Schoon, I., \& Marsh, H. (2016). A multination study of socioeconomic inequality in expectations for progression to higher education: The role of between-school tracking and ability stratification. \emph{American Educational Research Journal}, \emph{53}(1), 6--32. \url{https://doi.org/10.3102/0002831215621786}

\leavevmode\hypertarget{ref-parker_information_2018}{}%
Parker, P., Marsh, H., Guo, J., Anders, J., Shure, N., \& Dicke, T. (2018). An information distortion model of social class differences in math self-concept, intrinsic value, and utility value. \emph{Journal of Educational Psychology}, \emph{110}(3), 445--463. \url{https://doi.org/10.1037/edu0000215}

\leavevmode\hypertarget{ref-parker2018}{}%
Parker, P., Marsh, H., Jerrim, J., Guo, J., \& Dicke, T. (2018). Inequity and excellence in academic performance: Evidence from 27 countries. \emph{American Educational Research Journal}, \emph{55}(4), 836--858. \url{https://doi.org/10.3102/0002831218760213}

\leavevmode\hypertarget{ref-dunn2007}{}%
\emph{Peabody picture vocabulary test--fourth edition}. (2007). American Psychological Association. \url{https://doi.org/10.1037/t15144-000}

\leavevmode\hypertarget{ref-rajmil2014}{}%
Rajmil, L., Herdman, M., Ravens-Sieberer, U., Erhart, M., Alonso, J., \& The European KIDSCREEN group, undefined. (2014). Socioeconomic inequalities in mental health and health-related quality of life (hrqol) in children and adolescents from 11 european countries. \emph{International Journal of Public Health}, \emph{59}(1), 95--105. \url{https://doi.org/10.1007/s00038-013-0479-9}

\leavevmode\hypertarget{ref-reardon2011}{}%
Reardon, S. (2011). The widening academic achievement gap between the rich and the poor: New evidence and possible explanations. In G. Duncan \& R. Murnane (Eds.), \emph{Whither opportunity} (pp. 91--116). Sage.

\leavevmode\hypertarget{ref-rios2020}{}%
Rios, J., Ling, G., Pugh, R., Becker, D., \& Bacall, A. (2020). Identifying critical 21st-century skills for workplace success: A content analysis of job advertisements. \emph{Educational Researcher}, \emph{49}(2), 80--89. \url{https://doi.org/10.3102/0013189X19890600}

\leavevmode\hypertarget{ref-saporito2003}{}%
Saporito, S. (2003). Private choices, public consequences: Magnet school choice and segregation by race and poverty. \emph{Social Problems}, \emph{50}(2), 181--203. \url{https://doi.org/10.1525/sp.2003.50.2.181}

\leavevmode\hypertarget{ref-goodman1997}{}%
\emph{Strengths and difficulties questionnaire}. (1997). American Psychological Association. \url{https://doi.org/10.1037/t00540-000}

\leavevmode\hypertarget{ref-thongrattana2010}{}%
Thongrattana, P. (2010). Assessing reliability and validity of a measurement instrument for studying uncertain factors in thai rice supply chain. In. Presented at the SBS hdr student conference. Retrieved from \url{https://ro.uow.edu.au/sbshdr/2010/papers/4/}

\leavevmode\hypertarget{ref-washbrook2011}{}%
Washbrook, E., \& Waldfogel, J. (2011). \emph{On your marks: Measuring the school readiness of children in low-to-middle income families}. Retrieved from \url{https://www.resolutionfoundation.org/app/uploads/2014/08/On-your-marks.pdf}
\end{cslreferences}

\endgroup

\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
\begin{quote}
76a9308631048ea51918fe017cdcc2e8d7af4a52
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}
\end{quote}

\end{document}

### 2020-10-15 R/functions.R
full models

@@ -1,3 +1,7 @@
+# None Target Functions ####
+z <- function(x) (x-mean(x,na.rm=TRUE))/sd(x, na.rm = TRUE)
+
+#Taget functions ####
 #' Extract and Merge Age 4 data
 #'
 #' @return data
@@ -46,7 +50,8 @@ age_4_data <- function(){
     )
   #Merge Data
   age_4_b <- left_join(age_4_b, age_0_b)
-  age_4_data <- bind_rows(age_4_k, age_4_b)
+  age_4_data <- bind_rows(age_4_k, age_4_b) %>%
+    select(-indig1, -indig2)
   # Return data
   return(age_4_data)
 }
@@ -63,16 +68,14 @@ age_8_data <- function(){
            stratum = stratum, par_sdq = easdqtb, par_social = eapsoc, par_peer = eapeer,
            par_conduct = eacondb,teach_sdq = etsdqtb,teach_social = etpsoc, 
            teach_peer = etpeer, teach_conduct = etcondb
-           ) %>%
-    mutate(cohort = 'K')
+           ) 
   # Age 8 data birth cohort
   age_8_b <- readit::readit(here::here("data","lsacgrb8.sas7bdat")) %>%
     select(cid = hicid, grade = epc06a1,state = estate,weight = eweight,
            stratum = stratum, par_sdq = easdqtb, par_social = eapsoc, 
            par_peer = eapeer, par_conduct = eacondb,
            teach_sdq = etsdqtb,teach_social = etpsoc, teach_peer = etpeer, teach_conduct = etcondb
-           ) %>%
-    mutate(cohort = 'B')
+           )
   # Merge data and return
   age_8_data <- bind_rows(age_8_k, age_8_b)
   return(age_8_data)
@@ -86,13 +89,13 @@ age_8_data <- function(){
 #'
 child_achievement_data <- function(){
   child_achievement_data <- readit::readit(here::here("data","lsacnaplan.sas7bdat")) %>%
-    select(cid = hicid, numeracy = y3num, read = y3read, writing = y3write, spelling = y3spel,
-           grammar = y3gram,status = y3status
+    select(cid = hicid, ach_numeracy = y3num, ach_read = y3read, ach_writing = y3write, ach_spelling = y3spel,
+           ach_grammar = y3gram,status = y3status
            ) %>%
-    mutate(across(numeracy:grammar,
+    mutate(across(ach_numeracy:ach_grammar,
                   ~replace(., . < 0, NA)
[…]

### 2020-10-15 _targets.R

@@ -1,0 +1,59 @@
+library(targets)
+source(here::here("R","functions.R"))
+options(tidyverse.quiet = TRUE)
+tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue"))
+tar_pipeline(
+  tar_target(
+    age_4_data,
+    age_4_data()
+  ) ,
+  tar_target(
+    age_8_data,
+    age_8_data()
+  ),
+  tar_target(
+    child_achievement_data,
+    child_achievement_data()
+  ),
+  tar_target(
+    school_achievement_data,
+    school_achievement_data(data = age_4_data)
+  ),
+  tar_target(
+    data,
+    data(data_list = list(age_4_data, age_8_data, child_achievement_data, school_achievement_data))
+  ),
+  tar_target(
+    codebook,
+    make_codebook(data),
+    format = "file"
+  ),
+  tar_target(
+    data_imp,
+    data_imp(data)
+  ),
+  tar_target(
+    peer_par,
+    models(data_imp, outcome = "peer", source = "par")
+  ),
+  tar_target(
+    conduct_par,
+    models(data_imp, outcome = "conduct", source = "par")
+  ),
+  tar_target(
+    social_par,
+    models(data_imp, outcome = "social", source = "par")
+  ),
+  tar_target(
+    peer_teach,
+    models(data_imp, outcome = "peer", source = "teach")
+  ),
+  tar_target(
+    conduct_teach,
+    models(data_imp, outcome = "conduct", source = "teach")
+  ),
+  tar_target(
+    social_teach,
+    models(data_imp, outcome = "social", source = "teach")
+  )
+)

### 2020-10-16 manuscript.Rmd
References not working in html.
Fixed some issues

@@
       - Conceptualization
+      - Formal Analysis
       - Writing - Original Draft Preparation
       - Writing - Review & Editing
-  - name          : "Ernst-August Doelle"
-    affiliation   : "1,2"
-    role:
+  - name          : "Taren Sanders"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "33 Berry St, North Sydney, NSW, 2060, Australia"
+    email         : "taren.sanders@acu.edu.au"
+    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - Conceptualization
+      - Writing - Original Draft Preparation
       - Writing - Review & Editing
@@
   - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "UCL Institute of Education, UCL"
 authornote: |
@@
 In the vast majority of cases (98%), parent data came from mothers.
-Weights were provided by the survey organizers to ensure the sample represents the Australian population.
-This includes weights provided that aim to account for attrition.
+A participant flow diagram is present end in <!--# include figure 1 --> including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 These high-stakes tests have a mean of 500 and a standard deviation of 100.
+Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN[^3].
 For verbal ability we used the Peabody Picture Vocabulary test [@dunn2007] given to participants at age 4 (i.e., at or just before school commencement).
 Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4.
-Both numeracy achievement and verbal ability were z-scored in the analysis.
+The sector (government or non-government) of the school the child attended was also included as a covariate.
+
+[^3]: Parallel analysis suggested a single principal component was sufficient to account for the variation in the NAPLAN scores.
 
@@
 All other variables had 6% missing data or less.
-To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
-These 10 imputations were used in all analyses.
-All models used cluster robust standard errors to adjust standard errors for school clustering.
-LSAC strata was included as a fixed effect to account for the complex survey structure of the data.
-Analysis was done in the Zelig package in R [@imai2008].
+To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
+These imputations were used in all analyses.
+All analyses were done using Bayes via the BRMS package in R [@bürkner2017].
 
-All models in the paper that predicted social skills were estimated using Tobit models.
+All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobi regression.
 Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see [see @kleiber2008].
@@
 Johnson-Neyman regions of significance were used to investigate significant interactions.
-We also attempted to run random-intercept models using for school id both maximum likelihood estimation and MCMC sampling.
-In both cases models did not converge likely due to the unique nature of the data (i.e., many participants were the single representative of their school).
+The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
 
@@
 \endgroup
+
+> > > > > > > 76a9308631048ea51918fe017cdcc2e8d7af4a52

### 2020-10-16 R/functions.R

@@ -154,6 +154,79 @@ make_codebook <- function(data){
   return(here::here("data",glue::glue("{Sys.Date()}_codebook.pdf")))
 }
 
+consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievement_data, school_achievement_data)) {
+  child_data = reduce(data_list,left_join, by = "cid")
+  
+  total = nrow(child_data)
+  excluded_grade = nrow(child_data %>% filter(grade != 19) )
+  excluded_eligable = nrow(child_data %>% filter(status > 3) )
+  excluded_home = nrow(child_data %>% filter(!is.na(sid)) )
+  include = nrow(reduce(data_list,left_join, by = "cid") %>% filter(status < 4,grade == 19,!is.na(sid)))
+  
+  a1 <- glue::glue('Total available participants at age 4\n(n = {total})')
+  b1 <- ''
+  c1 <- ''
+  d1 <- ''
+  e1 <- glue::glue('Included for analysis\n(n = {include})')
+  f1 <- 'Data linked with\nadministrative data (NAPLAN; MySchool)'
+  
+  a2 <- ''
+  b2 <- glue::glue('Excluded because of\nwrong school grade (n = {excluded_grade})')
+  c2 <- glue::glue('Excluded because of\nnot eligable for NAPLAN test (n = {excluded_eligable})')
+  d2 <- glue::glue('Excluded because of\nout of school (n = {excluded_home})')
+  e2 <- ''
+  f2 <- ''
+  
+  #Create a node dataframe
+  ndf <- create_node_df(
+    n = 12,
+    label = c(a1, b1, c1, d1, e1, f1, #Column 1
+              a2, b2, c2, d2, e2, f2), #Column 2
+    style = rep('solid',12),
+    shape = rep('box',12),
+    # style = c('solid', 'invis', 'invis','invis', 'solid', 'solid', #Column 1
+    #           'invis', 'solid', 'solid', 'solid','invis', 'invis'), #Column 2
+    # shape = c('box', 'point', 'point', 'point','box','box', #Column 1 
+    #           'plaintext', 'box', 'box','box', 'point', 'point'), #Column 2
+    # width = c(3, 0.001, 0.001, 0.001,3, 3, #Column 1
+    #           2.5, 2.5, 2.5, 2.5, 0.001, 0.001), #Column 2
+    # height = c(1, 0.001, 0.001, 0.001, 1, 1, #Column 1
[…]

### 2020-10-16 R/make.R

@@ -1,0 +1,9 @@
+# Make file for analysis ####
+library(targets)
+# Run all analysis
+tar_make()
+# Check for outdated targets
+#tar_outdated()
+# Visualize status of the project
+#tar_glimpse()
+

### 2020-10-16 _targets.R

@@ -1,7 +1,7 @@
 library(targets)
 source(here::here("R","functions.R"))
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue"))
+tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue", "brms"))
 tar_pipeline(
   tar_target(
     age_4_data,

### 2020-10-18 R/functions.R
Flow diagramme added

@@ -160,71 +160,78 @@ consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievem
   total = nrow(child_data)
   excluded_grade = nrow(child_data %>% filter(grade != 19) )
   excluded_eligable = nrow(child_data %>% filter(status > 3) )
-  excluded_home = nrow(child_data %>% filter(!is.na(sid)) )
+  excluded_home = nrow(child_data %>% filter(is.na(sid)) )
   include = nrow(reduce(data_list,left_join, by = "cid") %>% filter(status < 4,grade == 19,!is.na(sid)))
   
-  a1 <- glue::glue('Total available participants at age 4\n(n = {total})')
+  a1 <- glue::glue("\'Total available participants at age 4 \\n (n = {total})\'")
   b1 <- ''
   c1 <- ''
   d1 <- ''
-  e1 <- glue::glue('Included for analysis\n(n = {include})')
-  f1 <- 'Data linked with\nadministrative data (NAPLAN; MySchool)'
+  e1 <- glue::glue("\'Included for analysis\\n(n = {include})\'")
+  f1 <- glue::glue("'Data linked with administrative data \\n (NAPLAN; MySchool)'")
   
   a2 <- ''
-  b2 <- glue::glue('Excluded because of\nwrong school grade (n = {excluded_grade})')
-  c2 <- glue::glue('Excluded because of\nnot eligable for NAPLAN test (n = {excluded_eligable})')
-  d2 <- glue::glue('Excluded because of\nout of school (n = {excluded_home})')
+  b2 <- glue::glue("\'Excluded: \\nWrong school grade\\n (n = {excluded_grade})\'")
+  c2 <- glue::glue("\'Excluded: \\nNot eligable for NAPLAN test\\n (n = {excluded_eligable})\'")
+  d2 <- glue::glue("\'Excluded: \\nSchool Uncertain\\n (n = {excluded_home})\'")
   e2 <- ''
   f2 <- ''
   
-  #Create a node dataframe
-  ndf <- create_node_df(
-    n = 12,
-    label = c(a1, b1, c1, d1, e1, f1, #Column 1
-              a2, b2, c2, d2, e2, f2), #Column 2
-    style = rep('solid',12),
-    shape = rep('box',12),
-    # style = c('solid', 'invis', 'invis','invis', 'solid', 'solid', #Column 1
-    #           'invis', 'solid', 'solid', 'solid','invis', 'invis'), #Column 2
[…]

### 2020-10-18 R/functions.R
flow added to targets 2

@@ -225,12 +225,11 @@ consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievem
       [11]: ''
       [12]: ''
       ")
-  #library(DiagrammeRsvg)#
-  #library(rsvg)#
+
   grViz(mod) %>%
     export_svg %>% charToRaw %>% rsvg_png("graph.png")
   
-  return(here::here("data",glue::glue("{Sys.Date()}_codebook.pdf")))
+  return(here::here("fig","flow.png"))
   
 }
 

### 2020-10-18 R/make.R

@@ -3,7 +3,6 @@ library(targets)
 # Run all analysis
 tar_make()
 # Check for outdated targets
-#tar_outdated()
+# tar_outdated()
 # Visualize status of the project
 #tar_glimpse()
-

### 2020-10-18 _targets.R

@@ -1,7 +1,10 @@
 library(targets)
+
 source(here::here("R","functions.R"))
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue", "brms"))
+tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue", "brms",
+                            "DiagrammeR","DiagrammeR", "rsvg")
+               )
 tar_pipeline(
   tar_target(
     age_4_data,
@@ -26,6 +29,11 @@ tar_pipeline(
   tar_target(
     codebook,
     make_codebook(data),
+    format = "file"
+  ),
+  tar_target(
+    flow,
+    consort_flow(data_list = list(age_4_data, age_8_data, child_achievement_data, school_achievement_data)),
     format = "file"
   ),
   tar_target(

### 2020-10-18 R/make.R
moved flow to fig folder

@@ -6,3 +6,29 @@ tar_make()
 # tar_outdated()
 # Visualize status of the project
 #tar_glimpse()
+
+
+tar_load(ends_with("teach"))
+tar_load(ends_with("par"))
+
+list(conduct_teach, social_teach, peer_teach,
+     conduct_par, social_par, peer_par) %>%
+  map(`[[`,1) %>%
+  map_dfr(broom.mixed::tidy, .id = "model") %>%
+  filter(term == 'ses_sch')
+
+
+list(conduct_teach, social_teach, peer_teach,
+     conduct_par, social_par, peer_par) %>%
+  map(`[[`,2) %>%
+  map_dfr(broom.mixed::tidy, .id = "model") %>%
+  filter(term %in% c('ses_sch', 'ses_sch:ses') ) %>%
+  mutate(sig = ifelse(
+    sign(conf.low) == sign(conf.high), '*', ''
+  ))
+
+plot(ggeffects::ggpredict(conduct_teach$interaction,c('ses_sch', 'ses [-2,0,2]')))
+
+plot(brms::marginal_effects(conduct_teach$interaction,'ses_sch:ses',
+                            spaghetti = T, nsamples = 150),
+     theme = theme_light() )

### 2020-10-19 R/functions.R
October 19, 2020

@@ -227,7 +227,7 @@ consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievem
       ")
 
   grViz(mod) %>%
-    export_svg %>% charToRaw %>% rsvg_png("graph.png")
+    export_svg %>% charToRaw %>% rsvg_png(here::here("fig","flow.png"))
   
   return(here::here("fig","flow.png"))
   

### 2020-10-20 manuscript.Rmd
Plots Added

@@
 
+```{r participants}
+library(targets)
+library(scales)
+tar_load(data)
+school_no = table(data$sid) %>%
+  table()
+
+school_one = as.vector(school_no[1]*1/nrow(data)) %>% scales::percent(accuracy = .01)
+school_two = as.vector(school_no[2]*2/nrow(data)) %>% scales::percent(accuracy = .01)
+boys = as.vector(table(data$gender)[1]/nrow(data)) %>% scales::percent(accuracy = .01)
+```
+
+```{r consort,fig.height=4,fig.width=2, fig.cap = 'Participant flow diagramme'}
+knitr::include_graphics(here::here('fig','flow.pdf'))
+```
+
 We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC).
@@
 Both cohorts of children have been followed every two years since [@aifs2015].
-LSAC is a complex survey with a stratified sampling design (that we account for in all analyses) and includes linked administration records for student performance in a standardized national numeracy test.
+LSAC includes linked administration records for student performance in a standardized national numeracy test.
 We also use government collected data on school average socioeconomic status.
-Many children in the sample were the only child surveyed (27%) or were one of two children surveyed (36%) in their school.
+Many children in the sample were the only child surveyed (`r school_one`) or were one of two children surveyed (`r school_two`) in their school.
 By using administrative records, we gain school average socioeconomic status based on the child's $\text{\underline{whole}}$ school grade from a high-quality source, thus avoiding sampling bias present in most school context research.
-We excluded participants who were not in school (e.g., were home schooled), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
-Together our total sample was 5,361 (52.1% male) children in Year 3 at school (\~age 8).
+We excluded participants who were not in school (e.g., were home schooled, or were otherwise missing an identifier of which school they attended), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
+Together our total sample was `r nrow(data)` (`r boys` boys) children in Year 3 at school (\~age 8).
 In the vast majority of cases (98%), parent data came from mothers.
-A participant flow diagram is present end in <!--# include figure 1 --> including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
+A participant flow diagram is present end in Figure \@ref(fig:consort) including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 
-All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobi regression.
-Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see [see @kleiber2008].
+All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression.
+Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails ([see @kleiber2008].
 As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
-Johnson-Neyman regions of significance were used to investigate significant interactions.
 The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
-
-Contextual effects like those reported here can be confusing to interpret.
-This is because positive effects of contextual variables can be incorrectly interpreted as positive in implication (or vice-versa for negative effects).
-This issue is compounded when interactions involving contextual effects are included in the model.
-Thus, we also provide counterfactual estimates of expected score differences for different levels of student SES and school average SES when the interaction between these variables is significant.
-Because of the number of covariates (35 in total because many of the covariates, such as strata, were categorical and thus were included as a series of dummy variables), we only report results for the main analysis variables.
 
@@
 
+```{r table-1, caption='Effect of School Average SES on Social Skills',results='asis'}
+tar_load(linear_table)
+library(gt)
+library(tidyMB)
+
+linear_table %>%
+  gt() %>%
+  fmt_number(columns = 3:5, decimals = 3) %>%
+  apa_style(title = 'Table 1', subtitle = 'Effect of School Average SES on Social Skills') %>%
+  as_latex()
+```
+
 ## School Average SES Predicts Social Skills Controlling for Age 4 SES and Social Skills
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under the Model 1 column in Tables 1-3.
+Results are presented under in Table\\ \@ref(tab:table-1).
 Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).
@@
 School average SES also negatively predicted peer problems but only for parent reports.
+
+```{r table-2, caption='Effect of School Average SES by Student SES on Social Skills',results='asis'}
+tar_load(interaction_table)
+library(gt)
+library(tidyMB)
+
+interaction_table %>%
+  gt() %>%
+  fmt_number(columns = 3:5, decimals = 3) %>%
+  apa_style(title = 'Table 2', subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
+  as_latex()
+```
 
@@
 \endgroup
-
-> > > > > > > 76a9308631048ea51918fe017cdcc2e8d7af4a52

### 2020-10-20 R/functions.R

@@ -155,6 +155,7 @@ make_codebook <- function(data){
 }
 
 consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievement_data, school_achievement_data)) {
+  
   child_data = reduce(data_list,left_join, by = "cid")
   
   total = nrow(child_data)
@@ -212,24 +213,26 @@ consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievem
       {rank=same; constraint = FALSE; tab4 -> tab10}
 }
       
-      [1]: ",a1,"
+      [1]: ", a1,"
       [2]: ''
       [3]: ''
       [4]: ''
       [5]: ", e1,"
-      [6]: ",f1,"
+      [6]: ", f1,"
       [7]: ''
-      [8]: ",b2,"
-      [9]: ",c2,"
+      [8]: ", b2,"
+      [9]: ", c2,"
       [10]: ",d2,"
       [11]: ''
       [12]: ''
       ")
 
-  grViz(mod) %>%
-    export_svg %>% charToRaw %>% rsvg_png(here::here("fig","flow.png"))
+  DiagrammeR::grViz(mod) %>%
+    DiagrammeRsvg::export_svg() %>%
+    charToRaw %>%
+    rsvg::rsvg_pdf(here::here("fig","flow.pdf"))
   
-  return(here::here("fig","flow.png"))
+  return(here::here("fig","flow.pdf"))
   
 }
 
@@ -279,5 +282,102 @@ models <- function(d = data_imp, outcome = c("peer","social","conduct"), source 
   }
   
   
-  
+
 }
+
+
+linear_output <- function(model_list = list(conduct_teach, social_teach, peer_teach,
+                          conduct_par, social_par, peer_par)) {
+  out = model_list %>%
+    map(`[[`,'linear') %>%
+    map_dfr(broom.mixed::tidy, .id = "model") %>%
+    filter(term == 'ses_sch') %>%
+    mutate(model = case_when(
+      model == 1 ~ "conduct_teacher",
+      model == 2 ~ "social_teacher",
+      model == 3 ~ "peer_teacher",
+      model == 4 ~ "conduct_parent",
+      model == 5 ~ "social_parent",
+      model == 6 ~ "peer_parent"
+    )) %>%
+    tidyr::separate(model, into = c('Outcome', 'Report Source')) %>%
+    select(Outcome,`Report Source`, Estimate = estimate,
+           `-95% CI` = conf.low,`+95% CI` = conf.high)
+  
+  return(out)
+}
+
[…]

### 2020-10-20 R/make.R

@@ -8,27 +8,9 @@ tar_make()
 #tar_glimpse()
 
 
-tar_load(ends_with("teach"))
-tar_load(ends_with("par"))
-
-list(conduct_teach, social_teach, peer_teach,
-     conduct_par, social_par, peer_par) %>%
-  map(`[[`,1) %>%
-  map_dfr(broom.mixed::tidy, .id = "model") %>%
-  filter(term == 'ses_sch')
 
 
-list(conduct_teach, social_teach, peer_teach,
-     conduct_par, social_par, peer_par) %>%
-  map(`[[`,2) %>%
-  map_dfr(broom.mixed::tidy, .id = "model") %>%
-  filter(term %in% c('ses_sch', 'ses_sch:ses') ) %>%
-  mutate(sig = ifelse(
-    sign(conf.low) == sign(conf.high), '*', ''
-  ))
 
-plot(ggeffects::ggpredict(conduct_teach$interaction,c('ses_sch', 'ses [-2,0,2]')))
+           
+           
 
-plot(brms::marginal_effects(conduct_teach$interaction,'ses_sch:ses',
-                            spaghetti = T, nsamples = 150),
-     theme = theme_light() )

### 2020-10-20 _targets.R

@@ -3,7 +3,7 @@ library(targets)
 source(here::here("R","functions.R"))
 options(tidyverse.quiet = TRUE)
 tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue", "brms",
-                            "DiagrammeR","DiagrammeR", "rsvg")
+                            "DiagrammeR","DiagrammeR", "rsvg", "ggtext", "tarchetypes")
                )
 tar_pipeline(
   tar_target(
@@ -63,5 +63,45 @@ tar_pipeline(
   tar_target(
     social_teach,
     models(data_imp, outcome = "social", source = "teach")
+  ),
+  tar_target(
+    linear_table,
+    linear_output(list(conduct_teach, social_teach, peer_teach,
+                       conduct_par, social_par, peer_par))
+  ),
+  tar_target(
+    interaction_table,
+    interaction_output(list(conduct_teach, social_teach, peer_teach,
+                       conduct_par, social_par, peer_par))
+  ),
+  tar_target(
+    peer_par_plot,
+    plots(peer_par),
+    format = "file"
+  ),
+  tar_target(
+    conduct_par_plot,
+    plots(conduct_par),
+    format = 'file'
+  ),
+  tar_target(
+    social_par_plot,
+    plots(social_par),
+    format = 'file'
+  ),
+  tar_target(
+    peer_teach_plot,
+    plots(peer_teach),
+    format = "file"
+  ),
+  tar_target(
+    conduct_teach_plot,
+    plots(conduct_teach),
+    format = "file"
+  ),
+  tar_target(
+    social_teach_plot,
+    plots(social_teach),
+    format = "file"
   )
 )

### 2020-10-20 manuscript.Rmd
Bad files

@@
   apa_style(title = 'Table 1', subtitle = 'Effect of School Average SES on Social Skills') %>%
+  tab_header(label = "tab:table-1")
   as_latex()
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under in Table\\ \@ref(tab:table-1).
+Results are presented under in Table \@ref(tab:table-1).
 Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).

### 2020-10-20 manuscript.tex
Bad files

@@
   pdftitle={School Socioeconomic Status Context and Social Skills in Children},
-  pdfauthor={Philip Parker1 \& Ernst-August Doelle1,2},
+  pdfauthor={Philip Parker1 \& Taren Sanders1},
   pdflang={en-EN},
@@
 \title{School Socioeconomic Status Context and Social Skills in Children}
-\author{Philip Parker\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
+\author{Philip Parker\textsuperscript{1} \& Taren Sanders\textsuperscript{1}}
 \date{}
@@
 
-The authors made the following contributions. Philip Parker: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.
+The authors made the following contributions.
+Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
 
-Correspondence concerning this article should be addressed to Philip Parker, 33 Berry St, North Sydney, NSW, 2060, Australia.
-E-mail: \href{mailto:philip.parker@acu.edu.au}{\nolinkurl{philip.parker@acu.edu.au}}
+Correspondence concerning this article should be addressed to Taren Sanders, 33 Berry St, North Sydney, NSW, 2060, Australia.
+E-mail: \href{mailto:taren.sanders@acu.edu.au}{\nolinkurl{taren.sanders@acu.edu.au}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute for Positive Psychology and Education, Australian Catholic University\\\textsuperscript{2} Konstanz Business School}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute for Positive Psychology and Education, Australian Catholic University\\\textsuperscript{2} UCL Institute of Education, UCL}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{/Users/phparker/Dropbox/Projects_Research/CURRENT_WORKING/social_skills_2020/fig/flow.pdf}
+\caption{\label{fig:consort}Participant flow diagramme}
+\end{figure}
+
 We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC).
@@
 Both cohorts of children have been followed every two years since (AIFS, 2015).
-LSAC is a complex survey with a stratified sampling design (that we account for in all analyses) and includes linked administration records for student performance in a standardized national numeracy test.
+LSAC includes linked administration records for student performance in a standardized national numeracy test.
 We also use government collected data on school average socioeconomic status.
-Many children in the sample were the only child surveyed (27\%) or were one of two children surveyed (36\%) in their school.
+Many children in the sample were the only child surveyed (25.75\%) or were one of two children surveyed (20.96\%) in their school.
 By using administrative records, we gain school average socioeconomic status based on the child's \(\text{\underline{whole}}\) school grade from a high-quality source, thus avoiding sampling bias present in most school context research.
-We excluded participants who were not in school (e.g., were home schooled), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
-Together our total sample was 5,361 (52.1\% male) children in Year 3 at school (\textasciitilde age 8).
+We excluded participants who were not in school (e.g., were home schooled, or were otherwise missing an identifier of which school they attended), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
+Together our total sample was 5440 (52.08\% boys) children in Year 3 at school (\textasciitilde age 8).
 In the vast majority of cases (98\%), parent data came from mothers.
-Weights were provided by the survey organizers to ensure the sample represents the Australian population.
-This includes weights provided that aim to account for attrition.
+A participant flow diagram is present end in Figure~/\ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 These high-stakes tests have a mean of 500 and a standard deviation of 100.
+Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN\footnote{Parallel analysis suggested a single principal component was sufficient to account for the variation in the NAPLAN scores.}.
 For verbal ability we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement).
 Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4.
-Both numeracy achievement and verbal ability were z-scored in the analysis.
+The sector (government or non-government) of the school the child attended was also included as a covariate.
 
@@
 All other variables had 6\% missing data or less.
-To account for this, we constructed 10 imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, \& Blackwell, 2011).
-These 10 imputations were used in all analyses.
-All models used cluster robust standard errors to adjust standard errors for school clustering.
-LSAC strata was included as a fixed effect to account for the complex survey structure of the data.
-Analysis was done in the Zelig package in R (Imai, King, \& Lau, 2008).
+To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, \& Blackwell, 2011).
+These imputations were used in all analyses.
+All analyses were done using Bayes via the BRMS package in R (Bürkner, 2017).
 
-All models in the paper that predicted social skills were estimated using Tobit models.
-Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see (see Kleiber \& Zeileis, 2008).
+All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression.
+Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails ((see Kleiber \& Zeileis, 2008).
 As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
-Johnson-Neyman regions of significance were used to investigate significant interactions.
-We also attempted to run random-intercept models using for school id both maximum likelihood estimation and MCMC sampling.
-In both cases models did not converge likely due to the unique nature of the data (i.e., many participants were the single representative of their school).
-
-Contextual effects like those reported here can be confusing to interpret.
-This is because positive effects of contextual variables can be incorrectly interpreted as positive in implication (or vice-versa for negative effects).
-This issue is compounded when interactions involving contextual effects are included in the model.
-Thus, we also provide counterfactual estimates of expected score differences for different levels of student SES and school average SES when the interaction between these variables is significant.
-Because of the number of covariates (35 in total because many of the covariates, such as strata, were categorical and thus were included as a series of dummy variables), we only report results for the main analysis variables.
+The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
 
@@
 
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{llrrr}
+\caption*{
+\large Table 1\\ 
+\small Effect of School Average SES on Social Skills\\ 
+} \\ 
+\toprule
+Outcome & Report Source & Estimate & -95\% CI & +95\% CI \\ 
+\midrule
+conduct & teacher & $-0.153$ & $-0.259$ & $-0.043$ \\ 
+social & teacher & $0.148$ & $0.053$ & $0.243$ \\ 
+peer & teacher & $-0.116$ & $-0.204$ & $-0.027$ \\ 
+conduct & parent & $-0.108$ & $-0.172$ & $-0.045$ \\ 
+social & parent & $0.087$ & $0.015$ & $0.157$ \\ 
+peer & parent & $-0.134$ & $-0.206$ & $-0.063$ \\ 
+\bottomrule
+\end{longtable}
+
 \hypertarget{school-average-ses-predicts-social-skills-controlling-for-age-4-ses-and-social-skills}{%
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under the Model 1 column in Tables 1-3.
+Results are presented under in Table\textbackslash{} \ref{tab:table-1}.
 Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).
@@
 School average SES also negatively predicted peer problems but only for parent reports.
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{llrrr}
+\caption*{
+\large Table 2\\ 
+\small Effect of School Average SES by Student SES on Social Skills\\ 
+} \\ 
+\toprule
+Outcome & Report Source & Estimate & -95\% CI & +95\% CI \\ 
+\midrule
+conduct & teacher & $-0.160$ & $-0.269$ & $-0.049$ \\ 
+conduct & teacher & $0.151$ & $0.063$ & $0.239$ \\ 
+social & teacher & $0.156$ & $0.060$ & $0.250$ \\ 
+social & teacher & $-0.094$ & $-0.170$ & $-0.018$ \\ 
+peer & teacher & $-0.125$ & $-0.214$ & $-0.036$ \\ 
+peer & teacher & $0.116$ & $0.041$ & $0.189$ \\ 
+conduct & parent & $-0.112$ & $-0.174$ & $-0.049$ \\ 
+conduct & parent & $0.059$ & $0.006$ & $0.112$ \\ 
+social & parent & $0.090$ & $0.020$ & $0.159$ \\ 
+social & parent & $-0.031$ & $-0.088$ & $0.026$ \\ 
+peer & parent & $-0.143$ & $-0.214$ & $-0.071$ \\ 
+peer & parent & $0.118$ & $0.058$ & $0.177$ \\ 
+\bottomrule
+\end{longtable}
 
@@
 
+\leavevmode\hypertarget{ref-buxfcrkner2017}{}%
+Bürkner, P. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1). \url{https://doi.org/10.18637/jss.v080.i01}
+
 \leavevmode\hypertarget{ref-corcoran2018}{}%
@@
 Honaker, J., King, G., \& Blackwell, M. (2011). Amelia ii: A program for missing data. \emph{Journal of Statistical Software}, \emph{45}(7), 1--47.
-
-\leavevmode\hypertarget{ref-imai2008}{}%
-Imai, K., King, G., \& Lau, O. (2008). Toward a common framework for statistical analysis and development. \emph{Journal of Computational and Graphical Statistics}, \emph{17}(4), 892--913.
 
@@
 
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-\begin{quote}
-76a9308631048ea51918fe017cdcc2e8d7af4a52
-\end{quote}
-\end{quote}
-\end{quote}
-\end{quote}
-\end{quote}
-\end{quote}
-\end{quote}
-
 \end{document}

### 2020-10-21 manuscript.Rmd
Figure patchwork

--- 2020-10-20 manuscript.Rmd
@@
   fmt_number(columns = 3:5, decimals = 3) %>%
-  apa_style(title = 'Table 2', subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
+  apa_style(title = 'Table 2',
+            subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
   as_latex()
@@
 
-The predictive association of school average SES on social skills was not consistent across the SES gradient (see Model 2 results reported in Tables 1-3).
+The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table 2).
 Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior.
-Regions of significance for these interactions can be found in Figures 2-4.
+Interactions are plotted in Figures 2-4.
 Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4.
@@
 High SES children had almost no difference in peer problems regardless of the school they were in.
+
+```{r conduct}
+library(patchwork)
+tar_load(social_par_plot)
+tar_load(social_teach_plot)
+social_par_plot + social_teach_plot + plot_layout(guides = "collect")  & theme(legend.position = 'bottom')
+```
 

### 2020-10-21 manuscript.tex
Figure patchwork

@@
 In the vast majority of cases (98\%), parent data came from mothers.
-A participant flow diagram is present end in Figure~/\ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
+A participant flow diagram is present end in Figure~\ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under in Table\textbackslash{} \ref{tab:table-1}.
+Results are presented under in Table \ref{tab:table-1}.
 Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).

### 2020-10-21 R/functions.R

@@ -329,6 +329,7 @@ interaction_output <- function(model_list = list(conduct_teach, social_teach, pe
 }
 
 plots <- function(model=conduct_teach){
+  
   nm <- deparse(substitute(model)) %>%
     str_split(., "_",simplify = TRUE) %>% 
     str_to_sentence()
@@ -342,41 +343,61 @@ plots <- function(model=conduct_teach){
   
   conditions <- data.frame(ses = c(-2, 0, 2))
   
-  plot(brms::conditional_effects(model$interaction,'ses_sch:ses',
+  p = plot(brms::conditional_effects(model$interaction,'ses_sch:ses',
                                  spaghetti = T, nsamples = 500, 
                                  int_conditions = conditions))[[1]] +
-    labs(title = glue::glue('{source} Reported {outcome}'),
-         subtitle = "<span style='color:red'>High (+2 SD)</span>, <span style='color:green'>Average</span>, and <span style='color:blue'>Low (-2 SD)</span> SES Children",
-         color = "SES (S.D. Units",
-         y = "Outcome",
-         x = "School Average SES") +
+    labs(#title = glue::glue('{source} Reported {outcome}'),
+         #subtitle = "<span style='color:red'>High (+2 SD)</span>, <span style='color:green'>Average</span>, and <span style='color:blue'>Low (-2 SD)</span> SES Children",
+         color = "SES (Units: SD)",
+         y = "Outcome (Units: censorded 0-10)",
+         x = "School Average SES (Units: SD)") +
     tidyMB::theme_mb() +
     theme(
-      plot.subtitle = element_markdown()
-    )
+      legend.key = element_blank(),
+      plot.subtitle = element_markdown(),
+      legend.position = "bottom",
+      legend.background = element_rect(fill="white", color = "white")
+    ) 
   
-  ggsave(here::here("fig", glue::glue("interaction_{nm[1]}_{nm[2]}.png")),
+  ggsave(plot = p, filename = here::here("fig", glue::glue("interaction_{nm[1]}_{nm[2]}.png")),
          dpi = 300, width = 6, height = 4)
-  
-  return(here::here("fig", glue::glue("interaction_{nm[1]}_{nm[2]}.png")))
+  # 
+  # return(here::here("fig", glue::glue("interaction_{nm[1]}_{nm[2]}.png")))
[…]

### 2020-10-21 R/make.R

@@ -5,7 +5,7 @@ tar_make()
 # Check for outdated targets
 # tar_outdated()
 # Visualize status of the project
-#tar_glimpse()
+# tar_glimpse()
 
 
 

### 2020-10-21 _targets.R

@@ -5,6 +5,7 @@ options(tidyverse.quiet = TRUE)
 tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue", "brms",
                             "DiagrammeR","DiagrammeR", "rsvg", "ggtext", "tarchetypes")
                )
+
 tar_pipeline(
   tar_target(
     age_4_data,
@@ -76,32 +77,27 @@ tar_pipeline(
   ),
   tar_target(
     peer_par_plot,
-    plots(peer_par),
-    format = "file"
+    plots(peer_par)
   ),
   tar_target(
     conduct_par_plot,
-    plots(conduct_par),
-    format = 'file'
+    plots(conduct_par)
   ),
   tar_target(
     social_par_plot,
-    plots(social_par),
-    format = 'file'
+    plots(social_par)
   ),
   tar_target(
     peer_teach_plot,
-    plots(peer_teach),
-    format = "file"
+    plots(peer_teach)
   ),
   tar_target(
     conduct_teach_plot,
-    plots(conduct_teach),
-    format = "file"
+    plots(conduct_teach)
   ),
   tar_target(
     social_teach_plot,
-    plots(social_teach),
-    format = "file"
-  )
+    plots(social_teach)
+  ),
+  tarchetypes::tar_render(manuscript,"manuscript.Rmd")
 )

### 2020-10-23 manuscript.Rmd
Full results section

--- 2020-10-20 manuscript.Rmd
@@
     affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
+    corresponding : no    # Define only one corresponding author
     address       : "33 Berry St, North Sydney, NSW, 2060, Australia"
@@
       - Writing - Review & Editing
+
 affiliation:
@@
 In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior.
-We consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates in a representative longitudinal sample of Australian children.
+Using prospective representative data, integrated with government administrative data, we consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates.
 
@@
 We model social skills using both parent and teacher reports.
+Measures of SES at the student level are taken from in-person interviews with parents <!--# @Taren is this true --> and school SES (as well as student achievement) is taken from government administrative records.
 
@@
 
-[@jerrim2019] provide one test of the potential assimilative effect of schools on social skills.
+Jerrim and colleagues [-@jerrim2019] provide one test of the potential assimilative effect of schools on social skills.
 They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school.
@@
 Many children in the sample were the only child surveyed (`r school_one`) or were one of two children surveyed (`r school_two`) in their school.
-By using administrative records, we gain school average socioeconomic status based on the child's $\text{\underline{whole}}$ school grade from a high-quality source, thus avoiding sampling bias present in most school context research.
+By using administrative records, we gain school average socioeconomic status based on the child's $\text{\underline{whole}}$ school grade from a high-quality source, thus avoiding sampling bias present in much school context research.
 We excluded participants who were not in school (e.g., were home schooled, or were otherwise missing an identifier of which school they attended), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
@@
 
+```{r descriptives}
+tar_load(des_peer)
+tar_load(des_social)
+tar_load(des_conduct)
+```
+
 ### Social skills at age 4 and 8
@@
 Questions ask about the child's peer problems ("Rather solitary, tends to play alone"), conduct problems ("Often fights with other children or bullies them"), and prosociality ("Is kind to younger children").
-We used the total scores for these scales as developed by the LSAC administrators.
+Greatest lower bound estimates of reliability were all greater than .95 and parallel analysis suggested that a single component was sufficient for each of the parent and teacher and age 4 and age 8 social skills factors.
+On the basis of this evidence, we used the total scores for these scales as developed by the LSAC administrators.
 Scores ranged from 0 to 10.
-These scores were heavily left censored (see Figures 2-3) for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher.
-prosocial behaviors were heavily right censored.
-Parents and teachers had moderate agreement for children at age 8 with correlations ranging from r = .40 for peer and r = .39 for conduct problems to r = .28 for prosocial behavior.
+These scores were heavily left censored for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher.
+Prosocial behaviors were heavily right censored.
+Censoring can be viewed as a special type of missing data where scores on $y*$ below or above the bounds are curtailed to fit within the bounds resulting in the observed scores $y$ [@gelman2020regression].
+Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable $y*$ believed to underlie the censored variable (hence negative predicted scores were possible).
 
-The SDQ items measure a collection of associated behaviours and as such we treat them as index or a formative variable.
-In this case common reliability metrics do not make much sense; nor do reflective models like confirmatory factor analysis.
-Instead, we would want items that are not too redundant with each other.
-For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index [@diamantopoulos2008; @thongrattana2010].
-In our case no VIF for any item for any social skills construct was higher than 2.
-Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct [@thongrattana2010].
-This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report.
-Parallel analysis suggested that all social skills constructs could be represented by a single principal component (but in all cases more than one reflective factor).
-The only exception was age 4 peer problems were parallel analysis suggested two components but the eigenvalue on the second component was barely greater than the simulated data suggesting a single index was likely still appropriate.
+Parents and teachers had moderate agreement for children at age 8 with correlations ranging from *r* = `r glue::glue("{format(des_peer$agreement$fixed[1], digits = 2)} 95% CI[{format(des_peer$agreement$fixed[3], digits = 2)}, {format(des_peer$agreement$fixed[4], digits = 2)}]")` for peer and *r* = `r glue::glue("{format(des_conduct$agreement$fixed[1], digits = 2)} 95% CI[{format(des_conduct$agreement$fixed[3], digits = 2)}, {format(des_conduct$agreement$fixed[4], digits = 2)}]")` for conduct problems to *r* = `r glue::glue("{format(des_social$agreement$fixed[1], digits = 2)} 95% CI[{format(des_social$agreement$fixed[3], digits = 2)}, {format(des_social$agreement$fixed[4], digits = 2)}]")` for prosocial behavior.
 
@@
 Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN[^3].
-For verbal ability we used the Peabody Picture Vocabulary test [@dunn2007] given to participants at age 4 (i.e., at or just before school commencement).
-Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4.
+For verbal ability (as a proxy for cognitive school readiness) we used the Peabody Picture Vocabulary test [@dunn2007] given to participants at age 4 (i.e., at or just before school commencement).
+Cohort was included as a covariate in all models as was measures of rural status and gender---all measured at age 4.
 The sector (government or non-government) of the school the child attended was also included as a covariate.
@@
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
-All other variables had 6% missing data or less.
+All other variables had 6% missing data or less (see code book in supplementary materials <!--# COnvert to OSF link when ready -->).
 To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
@@
 All analyses were done using Bayes via the BRMS package in R [@bürkner2017].
+The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before point and uncertainty estimates were extracted.
+Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantages (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students [@gelman2020regression].
 
@@
 The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
+We clearly demarcate the units of analysis in tables and figures.
+I general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
+We retain the SDQ scores on the original scale because of the presence of norms [\@youth2016scoring].
+Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the 'close to average' group and may thus be a potential concern [@youth2016scoring].
+<!--# We could use post-stratification weights to provide clinically relevant proportions of the population for different levels of SES and school SES -->
 
@@
 
-In order to provide a context for the main analyses we first aimed to show a) that richer children enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
-Our findings show that high SES children do enter school with better social skills.
-Student SES is correlated with less peer (B = -.284 [-.345, -.222]) and conduct (B = -.371 [-.438, -.304]) problems.
-They also had slightly higher prosocial behavior (B = .074 [.009, .139 ]).
-Further, student SES was strongly related to the school average SES of the school they attended (r = .480 [.458, .499]; see Figure 1).
+In order to provide a context for the main analyses we first aimed to show: a) that children from advantaged backgrounds enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
+Student SES is slightly correlated with prosocial behaviour (`r glue::glue("{format(des_social$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$prior_ses$fixed[3], digits = 2)}, {format(des_social$prior_ses$fixed[4], digits = 2)}]")`) but more strongly and negatively with conduct (`r glue::glue("{format(des_conduct$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_conduct$prior_ses$fixed[3], digits = 2)}, {format(des_conduct$prior_ses$fixed[4], digits = 2)}]")`) and peer (`r glue::glue("{format(des_peer$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_peer$prior_ses$fixed[3], digits = 2)}, {format(des_peer$prior_ses$fixed[4], digits = 2)}]")`) problems.
+Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (`r glue::glue("{format(des_social$sch_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$sch_ses$fixed[3], digits = 2)}, {format(des_social$sch_ses$fixed[4], digits = 2)}]")`)
 
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under in Table\\ \@ref(tab:table-1).
-Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).
+Results for school average SES are presented under in Table \@ref(tab:table-1) (the full model results are in supplementary materials).
+Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (at or just prior to school enrollment) to age 8 (Year 3).
 School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills.
@@
   gt() %>%
-  fmt_number(columns = 3:5, decimals = 3) %>%
-  apa_style(title = 'Table 2', subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
+  fmt_number(columns = 4:6, decimals = 3) %>%
+  apa_style(title = 'Table 2',
+            subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
   as_latex()
@@
 
-The predictive association of school average SES on social skills was not consistent across the SES gradient (see Model 2 results reported in Tables 1-3).
+The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \@ref(tab:table-2)).
 Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior.
-Regions of significance for these interactions can be found in Figures 2-4.
-Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4.
-These results show that low SES children appear to be most influenced by school average SES.
-Middle to high SES children appeared to have largely similar scores regardless of the school they were in (see Figures 2-4).
-For example, both parents and teachers report that low SES children had approximately half a point more peer problems in more SES disadvantaged schools than in more SES advantaged schools (see Table 4).
-High SES children had almost no difference in peer problems regardless of the school they were in.
+Interactions are plotted in Figures \@ref(fig:conduct)-\@ref(fig:social).
+Full model results can be found in supplementary materials.
+The interaction plots suggest that school context effects were particularly potent for disadvantaged children but school context had minimal impact on social skills for children from advantaged backgrounds.
+
+```{r peer, fig.cap = 'Interaction Plot for Peer Problems'}
+library(patchwork)
+tar_load(peer_par_plot)
+tar_load(peer_teach_plot)
+peer_par_plot + peer_teach_plot + plot_layout(guides = "collect")  & theme(legend.position = 'bottom')
+```
+
+```{r conduct, fig.cap = 'Interaction Plot for Conduct Problems'}
+library(patchwork)
+tar_load(conduct_par_plot)
+tar_load(conduct_teach_plot)
+conduct_par_plot + conduct_teach_plot + plot_layout(guides = "collect")  & theme(legend.position = 'bottom')
+```
+
+```{r social, fig.cap = 'Interaction Plot for Prosociality'}
+library(patchwork)
+tar_load(social_par_plot)
+tar_load(social_teach_plot)
+social_par_plot + social_teach_plot + plot_layout(guides = "collect")  & theme(legend.position = 'bottom')
+```
 

### 2020-10-23 manuscript.tex
Full results section

@@
 We model social skills using both parent and teacher reports.
+Measures of SES at the student level are taken from in-person interviews with parents and school SES (as well as student achievement) is taken from government administrative records.
 
@@
 
-(Jerrim \& Sims, 2019) provide one test of the potential assimilative effect of schools on social skills.
+Jerrim and colleagues (2019) provide one test of the potential assimilative effect of schools on social skills.
 They found that living in an area with selective schools is associated with better social skills when compared to children living in districts without a selective school.
@@
 \centering
-\includegraphics{/Users/phparker/Dropbox/Projects_Research/CURRENT_WORKING/social_skills_2020/fig/flow.pdf}
+\includegraphics{/Users/philipparker/Dropbox/Projects_Research/CURRENT_WORKING/social_skills_2020/fig/flow.pdf}
 \caption{\label{fig:consort}Participant flow diagramme}
@@
 Many children in the sample were the only child surveyed (25.75\%) or were one of two children surveyed (20.96\%) in their school.
-By using administrative records, we gain school average socioeconomic status based on the child's \(\text{\underline{whole}}\) school grade from a high-quality source, thus avoiding sampling bias present in most school context research.
+By using administrative records, we gain school average socioeconomic status based on the child's \(\text{\underline{whole}}\) school grade from a high-quality source, thus avoiding sampling bias present in much school context research.
 We excluded participants who were not in school (e.g., were home schooled, or were otherwise missing an identifier of which school they attended), were not in Year 3 at the time of testing, or were not eligible to complete national standardized testing (e.g., children with an intellectual disability).
@@
 Questions ask about the child's peer problems (``Rather solitary, tends to play alone''), conduct problems (``Often fights with other children or bullies them''), and prosociality (``Is kind to younger children'').
-We used the total scores for these scales as developed by the LSAC administrators.
+Greatest lower bound estimates of reliability were all greater than .95 and parallel analysis suggested that a single component was sufficient for each of the parent and teacher and age 4 and age 8 social skills factors.
+On the basis of this evidence, we used the total scores for these scales as developed by the LSAC administrators.
 Scores ranged from 0 to 10.
-These scores were heavily left censored (see Figures 2-3) for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher.
-prosocial behaviors were heavily right censored.
-Parents and teachers had moderate agreement for children at age 8 with correlations ranging from r = .40 for peer and r = .39 for conduct problems to r = .28 for prosocial behavior.
+These scores were heavily left censored for peer and conduct problems with a preponderance of students being scored as a zero by their parent or teacher.
+Prosocial behaviors were heavily right censored.
+Censoring can be viewed as a special type of missing data where scores on \(y*\) below or above the bounds are curtailed to fit within the bounds resulting in the observed scores \(y\) (Gelman, Hill, \& Vehtari, 2020).
+Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable \(y*\) believed to underlie the censored variable (hence negative predicted scores were possible).
 
-The SDQ items measure a collection of associated behaviours and as such we treat them as index or a formative variable.
-In this case common reliability metrics do not make much sense; nor do reflective models like confirmatory factor analysis.
-Instead, we would want items that are not too redundant with each other.
-For formative scales a variable inflation factor (VIF) of less than 10 has been suggested has a cut-off score; with VIF values of less than 3.3 showing an item makes an excellent contribution to the index (Diamantopoulos, Riefler, \& Roth, 2008; Thongrattana, 2010).
-In our case no VIF for any item for any social skills construct was higher than 2.
-Likewise, a limit of 30 on the condition index has been suggested as a good measure of the reliability of a formative construct (Thongrattana, 2010).
-This was the case for all our social skills constructs at both age 4 and age 8 and for both parent and teacher report.
-Parallel analysis suggested that all social skills constructs could be represented by a single principal component (but in all cases more than one reflective factor).
-The only exception was age 4 peer problems were parallel analysis suggested two components but the eigenvalue on the second component was barely greater than the simulated data suggesting a single index was likely still appropriate.
+Parents and teachers had moderate agreement for children at age 8 with correlations ranging from \emph{r} = 0.41 95\% CI{[}0.38, 0.43{]} for peer and \emph{r} = 0.4 95\% CI{[}0.37, 0.42{]} for conduct problems to \emph{r} = 0.29 95\% CI{[}0.26, 0.31{]} for prosocial behavior.
 
@@
 Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN\footnote{Parallel analysis suggested a single principal component was sufficient to account for the variation in the NAPLAN scores.}.
-For verbal ability we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement).
-Cohort was included as a covariate in all models as was measures of rural status, Indigenous status, language other than English status, state of residence, and gender---all measured at age 4.
+For verbal ability (as a proxy for cognitive school readiness) we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement).
+Cohort was included as a covariate in all models as was measures of rural status and gender---all measured at age 4.
 The sector (government or non-government) of the school the child attended was also included as a covariate.
@@
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14\%.
-All other variables had 6\% missing data or less.
+All other variables had 6\% missing data or less (see code book in supplementary materials ).
 To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package (Honaker, King, \& Blackwell, 2011).
@@
 All analyses were done using Bayes via the BRMS package in R (Bürkner, 2017).
+The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before point and uncertainty estimates were extracted.
+Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantages (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students (Gelman et al., 2020).
 
@@
 The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
+We clearly demarcate the units of analysis in tables and figures.
+I general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
+We retain the SDQ scores on the original scale because of the presence of norms {[}@youth2016scoring{]}.
+Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the `close to average' group and may thus be a potential concern (Mind, 2016).
 
@@
 
-In order to provide a context for the main analyses we first aimed to show a) that richer children enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
-Our findings show that high SES children do enter school with better social skills.
-Student SES is correlated with less peer (B = -.284 {[}-.345, -.222{]}) and conduct (B = -.371 {[}-.438, -.304{]}) problems.
-They also had slightly higher prosocial behavior (B = .074 {[}.009, .139 {]}).
-Further, student SES was strongly related to the school average SES of the school they attended (r = .480 {[}.458, .499{]}; see Figure 1).
+In order to provide a context for the main analyses we first aimed to show: a) that children from advantaged backgrounds enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
+Student SES is slightly correlated with prosocial behaviour (0.05 95\% CI{[}0.023, 0.078{]}) but more strongly and negatively with conduct (-0.18 95\% CI{[}-0.2, -0.15{]}) and peer (-0.16 95\% CI{[}-0.19, -0.13{]}) problems.
+Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (0.48 95\% CI{[}0.46, 0.51{]})
 
@@
 \midrule
-conduct & teacher & $-0.153$ & $-0.259$ & $-0.043$ \\ 
-social & teacher & $0.148$ & $0.053$ & $0.243$ \\ 
-peer & teacher & $-0.116$ & $-0.204$ & $-0.027$ \\ 
-conduct & parent & $-0.108$ & $-0.172$ & $-0.045$ \\ 
-social & parent & $0.087$ & $0.015$ & $0.157$ \\ 
-peer & parent & $-0.134$ & $-0.206$ & $-0.063$ \\ 
+conduct & teacher & $-0.152$ & $-0.261$ & $-0.046$ \\ 
+social & teacher & $0.147$ & $0.056$ & $0.241$ \\ 
+peer & teacher & $-0.115$ & $-0.204$ & $-0.024$ \\ 
+conduct & parent & $-0.107$ & $-0.169$ & $-0.045$ \\ 
+social & parent & $0.087$ & $0.018$ & $0.156$ \\ 
+peer & parent & $-0.135$ & $-0.205$ & $-0.063$ \\ 
 \bottomrule
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results are presented under in Table \ref{tab:table-1}.
-Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (just prior to school enrollment) to age 8 (Year 3).
+Results for school average SES are presented under in Table \ref{tab:table-1} (the full model results are in supplementary materials).
+Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (at or just prior to school enrollment) to age 8 (Year 3).
 School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills.
@@
 \captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{llrrr}
+\begin{longtable}{lllrrr}
 \caption*{
@@
 \toprule
-Outcome & Report Source & Estimate & -95\% CI & +95\% CI \\ 
+Outcome & Report Source & term & Estimate & -95\% CI & +95\% CI \\ 
 \midrule
-conduct & teacher & $-0.160$ & $-0.269$ & $-0.049$ \\ 
-conduct & teacher & $0.151$ & $0.063$ & $0.239$ \\ 
-social & teacher & $0.156$ & $0.060$ & $0.250$ \\ 
-social & teacher & $-0.094$ & $-0.170$ & $-0.018$ \\ 
-peer & teacher & $-0.125$ & $-0.214$ & $-0.036$ \\ 
-peer & teacher & $0.116$ & $0.041$ & $0.189$ \\ 
-conduct & parent & $-0.112$ & $-0.174$ & $-0.049$ \\ 
-conduct & parent & $0.059$ & $0.006$ & $0.112$ \\ 
-social & parent & $0.090$ & $0.020$ & $0.159$ \\ 
-social & parent & $-0.031$ & $-0.088$ & $0.026$ \\ 
-peer & parent & $-0.143$ & $-0.214$ & $-0.071$ \\ 
-peer & parent & $0.118$ & $0.058$ & $0.177$ \\ 
+conduct & teacher & School Average SES & $-0.160$ & $-0.270$ & $-0.053$ \\ 
+conduct & teacher & School Average SES by Student SES & $0.151$ & $0.064$ & $0.240$ \\ 
+social & teacher & School Average SES & $0.156$ & $0.059$ & $0.252$ \\ 
+social & teacher & School Average SES by Student SES & $-0.094$ & $-0.172$ & $-0.017$ \\ 
+peer & teacher & School Average SES & $-0.126$ & $-0.216$ & $-0.036$ \\ 
+peer & teacher & School Average SES by Student SES & $0.117$ & $0.044$ & $0.191$ \\ 
+conduct & parent & School Average SES & $-0.112$ & $-0.175$ & $-0.048$ \\ 
+conduct & parent & School Average SES by Student SES & $0.058$ & $0.006$ & $0.111$ \\ 
+social & parent & School Average SES & $0.090$ & $0.020$ & $0.160$ \\ 
+social & parent & School Average SES by Student SES & $-0.031$ & $-0.087$ & $0.027$ \\ 
+peer & parent & School Average SES & $-0.144$ & $-0.217$ & $-0.074$ \\ 
+peer & parent & School Average SES by Student SES & $0.117$ & $0.058$ & $0.176$ \\ 
 \bottomrule
@@
 
-The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table 2).
+The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \ref{tab:table-2}).
 Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior.
-Interactions are plotted in Figures 2-4.
-Likewise, counterfactual expected score differences on the latent continuous social skill (underlying the censored social skill in the Tobit model) are given in Table 4.
-These results show that low SES children appear to be most influenced by school average SES.
-Middle to high SES children appeared to have largely similar scores regardless of the school they were in (see Figures 2-4).
-For example, both parents and teachers report that low SES children had approximately half a point more peer problems in more SES disadvantaged schools than in more SES advantaged schools (see Table 4).
-High SES children had almost no difference in peer problems regardless of the school they were in.
+Interactions are plotted in Figures \ref{fig:conduct}-\ref{fig:social}.
+Full model results can be found in supplementary materials.
+The interaction plots suggest that school context effects were particularly potent for disadvantaged children but school context had minimal impact on social skills for children from advantaged backgrounds.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/peer-1.pdf}
+\caption{\label{fig:peer}Interaction Plot for Peer Problems}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/conduct-1.pdf}
+\caption{\label{fig:conduct}Interaction Plot for Conduct Problems}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/social-1.pdf}
+\caption{\label{fig:social}Interaction Plot for Prosociality}
+\end{figure}
 
@@
 
-\leavevmode\hypertarget{ref-diamantopoulos2008}{}%
-Diamantopoulos, A., Riefler, P., \& Roth, K.
-(2008).
-Advancing formative measurement models.
-\emph{Journal of Business Research}, \emph{61}(12), 1203--1218.
-\url{https://doi.org/10.1016/j.jbusres.2008.01.009}
-
 \leavevmode\hypertarget{ref-durlak2011}{}%
@@
 \url{https://doi.org/10.1111/cdev.12649}
+
+\leavevmode\hypertarget{ref-gelman2020regression}{}%
+Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.
 
@@
 \url{https://doi.org/10.1016/S0277-9536(00)00346-4}
+
+\leavevmode\hypertarget{ref-youth2016scoring}{}%
+Mind, Y.
+in.
+(2016).
+\emph{Scoring the strengths \& difficulties questionnaire for age 4--17 or 18+}.
+Retrieved from \url{https://www.ehcap.co.uk/content/sites/ehcap/uploads/NewsDocuments/236/SDQEnglishUK4-17scoring-1.PDF}
 
@@
 
-\leavevmode\hypertarget{ref-thongrattana2010}{}%
-Thongrattana, P.
-(2010).
-Assessing reliability and validity of a measurement instrument for studying uncertain factors in thai rice supply chain.
-In.
-Presented at the SBS hdr student conference.
-Retrieved from \url{https://ro.uow.edu.au/sbshdr/2010/papers/4/}
-
 \leavevmode\hypertarget{ref-washbrook2011}{}%

### 2020-10-23 R/functions.R

@@ -236,6 +236,50 @@ consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievem
   
 }
 
+index_fun <- function(data = data, age = c("4","8"), variable = c("social", "peer", "conduct","social_t", "peer_t", "conduct_t")){
+  # Get ids from final data set
+  id = data$cid
+  # Choose id type
+  var = switch (variable,
+                social = 'se03a1',
+                peer = 'se03a5',
+                conduct = 'se03a4',
+                social_t = 'se03t1',
+                peer_t = 'se03t5',
+                conduct_t = 'se03t4'
+  )
+  
+  if(age == "4") {
+    
+    db = readit::readit(here::here("data", "lsacgrb4.sas7bdat")) %>% select(contains(var), cid = hicid)
+    dk = readit::readit(here::here("data", "lsacgrk4.sas7bdat")) %>% select(contains(var), cid = hicid)
+    
+  } else{
+    db = readit::readit(here::here("data", "lsacgrb8.sas7bdat")) %>% select(contains(var), cid = hicid)
+    dk = readit::readit(here::here("data", "lsacgrk8.sas7bdat")) %>% select(contains(var), cid = hicid)
+  }
+  
+  d = bind_rows(db,dk) %>%
+    mutate(total = rowSums(.), .before = 1) %>%
+    filter(cid %in% id) %>%
+    select(-cid)
+  
+  parallel = psych::fa.parallel(d[,-1])
+  glb = psych::glb(d[,-1])
+  
+  out = c(ncom = parallel$ncomp, glb = glb$glb.max)
+  
+  return(out)
+  
+}
+
+index_run <- function(data){
+  d = data
+  args = list(x = c(rep(c("4","8"),3),rep("8",3)), y = c(rep(c("social", "peer", "conduct"),each = 2), "social_t", "peer_t", "conduct_t"))
+  out = purrr::map2(.x = args$x,.y = args$y, .f = ~index_fun(data = d, age = .x, variable = .y)) %>%
+    bind_rows()
+}
+
 
 data_imp <- function(data){
   set.seed(1234)
@@ -247,6 +291,26 @@ data_imp <- function(data){
                            noms = c('geo', 'indig', 'gender', 'lang', 'cohort'),bounds = bound
   )
   return(data_imp)
+}
+
+ses_beta <- function(d = data_imp, outcome, m = 5, delta = 0.99, iterations = 6000){
+  
+  data = list()
+  for (i in 1:m){
[…]

### 2020-10-23 R/make.R

@@ -7,10 +7,3 @@ tar_make()
 # Visualize status of the project
 # tar_glimpse()
 
-
-
-
-
-           
-           
-

### 2020-10-23 _targets.R

@@ -3,7 +3,8 @@ library(targets)
 source(here::here("R","functions.R"))
 options(tidyverse.quiet = TRUE)
 tar_option_set(packages = c("readit", "tidyverse", "dataMaid","here", "glue", "brms",
-                            "DiagrammeR","DiagrammeR", "rsvg", "ggtext", "tarchetypes")
+                            "DiagrammeR","DiagrammeR", "rsvg", "ggtext", "tarchetypes",
+                            "psych")
                )
 
 tar_pipeline(
@@ -99,5 +100,20 @@ tar_pipeline(
     social_teach_plot,
     plots(social_teach)
   ),
-  tarchetypes::tar_render(manuscript,"manuscript.Rmd")
+  tar_target(
+    reliability,
+    index_run(data)
+  ),
+  tar_target(
+    des_conduct,
+    ses_beta(d = data_imp, outcome = 'conduct')
+  ),
+  tar_target(
+    des_peer,
+    ses_beta(d = data_imp, outcome = 'peer')
+  ),
+  tar_target(
+    des_social,
+    ses_beta(d = data_imp, outcome = 'social')
+  )
 )

### 2020-10-26 manuscript.Rmd
added equation

@@
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
-All other variables had 6% missing data or less (see code book in supplementary materials <!--# COnvert to OSF link when ready -->).
+All other variables had 6% missing data or less (see code book in supplementary materials <!--# Convert to OSF link when ready -->).
 To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
@@
 <!--# We could use post-stratification weights to provide clinically relevant proportions of the population for different levels of SES and school SES -->
+
+The models are fit using the following formula:
+
+$$
+y^{*}_{i} \sim N(\alpha_{j[i]} + X_i\beta + \sigma^{2}_{y}), for~i = 1,...,n\\
+\alpha_{j} \sim N(U_j \gamma, \sigma^{2}_{\alpha}), for~j = 1,...k
+$$
+
+Here $X_i$ is a matrix of individual level predictors including individual SES, prior social skills, gender, cohort, geographic location, and an NAPLAN achievement index.
+$U_j$ is a matrix of school level predictors including school section and the critical variable school SES.
+Note that we predict $y^{*}_{i}$ which is the latent continuous variable underlying the observed variables which is defined as:
+
+$$
+y_i =  \begin{cases}      y*_{i} & \text{if $y*_{i} < 0$ or $y*_{i} > 10 $}\\      0 & \text{if $y*_{i} < 0$}\\      10 & \text{if $y*_{i} > 10$}    \end{cases}
+$$
 

### 2020-10-28 manuscript.Rmd
Minor edits to manuscript

@@
 ---
-title             : "School Socioeconomic Status Context and Social Skills in Children"
-shorttitle        : "School SES and Social Skills"
+title: "School Socioeconomic Status Context and Social Skills in Children"
 author:
-  - name          : "Philip Parker"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "33 Berry St, North Sydney, NSW, 2060, Australia"
-    email         : "philip.parker@acu.edu.au"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+
+- name: Philip Parker
+  affiliation: '1'
+  corresponding: yes
+  address: 33 Berry St, North Sydney, NSW, 2060, Australia
+  email: philip.parker@acu.edu.au
+  role:
   - Conceptualization
@@
   - Writing - Review & Editing
-  - name          : "Taren Sanders"
-    affiliation   : "1"
-    corresponding : no    # Define only one corresponding author
-    address       : "33 Berry St, North Sydney, NSW, 2060, Australia"
-    email         : "taren.sanders@acu.edu.au"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+- name: Taren Sanders
+  affiliation: '1'
+  # corresponding: no
+  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
+  # email: taren.sanders@acu.edu.au
+  role:
   - Conceptualization
@@
   - Writing - Review & Editing
-
-affiliation:
-  - id            : "1"
-    institution   : "Institute for Positive Psychology and Education, Australian Catholic University"
-  - id            : "2"
-    institution   : "UCL Institute of Education, UCL"
+- name: Michael Noetel
+  affiliation: '1,3'
+  # corresponding: no
+  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
+  # email: taren.sanders@acu.edu.au
+  role:
+  - Conceptualization
+  - Writing - Original Draft Preparation
+  - Writing - Review & Editing
+shorttitle: School SES and Social Skills
+output:
+  pdf: papaja::apa6_pdf
+  html_document:
+    fig_caption: yes
+    toc: yes
+  pdf_document:
+    toc: yes
+  word: papaja::apa6_doc
+  html_notebook:
+    fig_caption: yes
+    toc: yes
+    highlight: haddock
+    theme: spacelab
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-  Enter author note here.
-abstract: |
-
+abstract: |2
   As researchers and policymakers try to understand the consequences of SES for educational and occupational attainment, they increasingly consider social skills to be critical.
@@
 Taking the evidence together, social stratified schools may harm children from low SES backgrounds while not benefiting children from high SES backgrounds.
-keywords          : "social skills; assimilation effects; socioeconomic status; school context"
-wordcount         : "X"
-bibliography      : ["r-references.bib"]
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-documentclass     : "apa6"
-classoption       : "man"
-output            : 
-  pdf: 
-    papaja::apa6_pdf
-  html_document: 
-    fig_caption: yes
-    toc: yes
-  html_notebook: 
-    fig_caption: yes
-    toc: yes
-    highlight: haddock
-    theme: spacelab
-  word:
-    papaja::apa6_doc
+keywords: social skills; assimilation effects; socioeconomic status; school context
+wordcount: X
+bibliography: r-references.bib
+floatsintext: no
+figurelist: no
+tablelist: no
+footnotelist: no
+linenumbers: yes
+mask: no
+draft: no
+documentclass: apa6
+classoption: man
+affiliation:
+- id: '1'
+  institution: Institute for Positive Psychology and Education, Australian Catholic
+    University
+- id: '2'
+  institution: UCL Institute of Education, UCL
+- id: '3'
+  institution: School of Health and Behavioural Sciences, Australian Catholic University
 ---
@@
 In the vast majority of cases (98%), parent data came from mothers.
-A participant flow diagram is present end in Figure \@ref(fig:consort) including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
+A participant flow diagram is presented in Figure \@ref(fig:consort) including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression.
-Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails ([see @kleiber2008].
+Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails [see @kleiber2008].
 As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
@@
 I general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
-We retain the SDQ scores on the original scale because of the presence of norms [\@youth2016scoring].
+We retain the SDQ scores on the original scale because of the presence of norms [@youth2016scoring].
 Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the 'close to average' group and may thus be a potential concern [@youth2016scoring].
@@
 $$
-y^{*}_{i} \sim N(\alpha_{j[i]} + X_i\beta + \sigma^{2}_{y}), for~i = 1,...,n\\
+\begin{split}
+y^{*}_{i} \sim N(\alpha_{j[i]} + X_i\beta + \sigma^{2}_{y}), for~i = 1,...,n \\
 \alpha_{j} \sim N(U_j \gamma, \sigma^{2}_{\alpha}), for~j = 1,...k
+\end{split}
 $$
@@
 $$
-y_i =  \begin{cases}      y*_{i} & \text{if $y*_{i} < 0$ or $y*_{i} > 10 $}\\      0 & \text{if $y*_{i} < 0$}\\      10 & \text{if $y*_{i} > 10$}    \end{cases}
+\begin{split}
+y_i =  \begin{cases}      
+            y*_{i} & \text{if $y*_{i} < 0$ or $y*_{i} > 10 $}\\      
+            0 & \text{if $y*_{i} < 0$}\\      
+            10 & \text{if $y*_{i} > 10$}    
+        \end{cases}
+\end{split}
 $$
@@
 Parker et al.
-[-@parker2016; -@parker2018 ; -@parker_information_2018; -@parker2019] have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
+[-@parker2016; -@parker2018; -@parker_information_2018; -@parker2019] have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
 They argue that empirical evidence suggests policy should encourage school selection policies that maximize within school heterogeneity.

### 2020-10-29 manuscript.Rmd
full draft 2020-10-29

@@
   affiliation: '1'
+  # corresponding: no
+  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
+  # email: taren.sanders@acu.edu.au
+  role:
+  - Conceptualization
+  - Writing - Original Draft Preparation
+  - Writing - Review & Editing
+- name: Jake Anders
+  affiliation: '2'
   # corresponding: no
@@
 Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development.
-In a representative longitudinal sample of 5361 Australian children, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers.
-All models controlled for prior social skills at age 4, and a wide range of covariates.
+In a representative longitudinal sample of  `r nrow(data)` Australian 8-year olds, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers.
+All models controlled for prior social skills at age 4, and additional school and student covariates.
 We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems.
@@
 In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman.
-[@heckman2006] argues that research and intervention on reducing educational inequality have focused too narrowly on "cognitive" abilities.
+Heckman [-@heckman2006] argues that research and intervention on reducing educational inequality have focused too narrowly on "cognitive" abilities (generally narrowly deinfed to include ability, IQ, and academic achievement).
 Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment.
@@
 In particular, we explore whether school average socioeconomic status predicts social skills in Year 3 (age 8)---using measures of peer problems, conduct problems, and prosocial behavior.
-Using prospective representative data, integrated with government administrative data, we consider if this association is independent of children's incoming social skills at age 4, children's own socioeconomic status, and a set of demographic and academic covariates.
+Using prospective representative data, integrated with government administrative data, we found this is that case but that an interaction between student and school SES means that poorer children appear more susceptible to such effects.
 
@@
 We model social skills using both parent and teacher reports.
-Measures of SES at the student level are taken from in-person interviews with parents <!--# @Taren is this true --> and school SES (as well as student achievement) is taken from government administrative records.
+Measures of SES at the student level are taken from in-person interviews with parents <!--# @taren.sanders@acu.edu.au is this true --> and school SES (as well as student achievement) is taken from government administrative records.
 
@@
 We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills [@delaat2016; @garratt2017; @mcmunn2001; @rajmil2014].
-Given Australian schools are socially stratified [@oecd2015; @parker2018], we also expect low SES children to attend more disadvantaged schools on average.
+Given Australian schools are socially stratified [@oecd2015; @parker2018], we expect low SES children to attend more disadvantaged schools on average.
 
@@
 Together our total sample was `r nrow(data)` (`r boys` boys) children in Year 3 at school (\~age 8).
-In the vast majority of cases (98%), parent data came from mothers.
+In the vast majority of cases (98% <!--# @philip.parker@acu.edu.au should extract this from data rather than previous research?
+-->), parent data came from mothers.
 A participant flow diagram is presented in Figure \@ref(fig:consort) including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
@@
 Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ [@goodman1997].
-We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child's *parent* and the child's *teacher*.
+We explored social skills at age 4-5 (for prior social skills) and age 8-9 (as primary outcomes) as reported by the child's *parent* and the child's *teacher*.
 The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
@@
 Prosocial behaviors were heavily right censored.
-Censoring can be viewed as a special type of missing data where scores on $y*$ below or above the bounds are curtailed to fit within the bounds resulting in the observed scores $y$ [@gelman2020regression].
-Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable $y*$ believed to underlie the censored variable (hence negative predicted scores were possible).
+Censoring can be viewed as a special type of missing data where scores on $y*$ (the hypothesised true latent distribution of the variable) below or above the bounds are curtailed to fit within the bounds resulting in the observed scores $y$ [@gelman2020regression].
+Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable $y*$ believed to underlie the censored variable (hence negative predicted scores or scores over 10 were possible).
 
@@
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
-All other variables had 6% missing data or less (see code book in supplementary materials <!--# Convert to OSF link when ready -->).
+All other variables had 6% missing data or less (see code book in supplementary materials <!--# @philip.parker@acu.edu.au Convert to OSF link when ready -->).
 To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
@@
 All analyses were done using Bayes via the BRMS package in R [@bürkner2017].
-The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before point and uncertainty estimates were extracted.
-Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantages (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students [@gelman2020regression].
+The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before estimates and their uncertainty were extracted.
+Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantaged (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students [@gelman2020regression].
 
@@
 We clearly demarcate the units of analysis in tables and figures.
-I general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
+In general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
 We retain the SDQ scores on the original scale because of the presence of norms [@youth2016scoring].
 Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the 'close to average' group and may thus be a potential concern [@youth2016scoring].
-<!--# We could use post-stratification weights to provide clinically relevant proportions of the population for different levels of SES and school SES -->
+<!--# @everyone We could use post-stratification weights to provide 'clinically' relevant proportions of the population for different levels of SES and school SES -->
 
@@
 
-Here $X_i$ is a matrix of individual level predictors including individual SES, prior social skills, gender, cohort, geographic location, and an NAPLAN achievement index.
-$U_j$ is a matrix of school level predictors including school section and the critical variable school SES.
-Note that we predict $y^{*}_{i}$ which is the latent continuous variable underlying the observed variables which is defined as:
+Here $X$ is a matrix of individual level predictors for student $i$ including individual SES, prior social skills, gender, cohort, geographic location, and an NAPLAN achievement index.
+Prior social skills is particularly critical as this provide stronger conclusions that school context rather than selection effects account for the influence of school average SES on student social skills.
+Put simply the estimate of school average achievement can be interpreted as the influence of school context on the change in children's social skills from age 4 to age 8.
+$U_j$ is a matrix of school level predictors including school section and the aforementioned critical variable *school SES*.
+Note that we predict $y^{*}_{i}$ which is the latent continuous variable underlying the observed variables which are defined as:
 
@@
 In order to provide a context for the main analyses we first aimed to show: a) that children from advantaged backgrounds enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
-Student SES is slightly correlated with prosocial behaviour (`r glue::glue("{format(des_social$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$prior_ses$fixed[3], digits = 2)}, {format(des_social$prior_ses$fixed[4], digits = 2)}]")`) but more strongly and negatively with conduct (`r glue::glue("{format(des_conduct$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_conduct$prior_ses$fixed[3], digits = 2)}, {format(des_conduct$prior_ses$fixed[4], digits = 2)}]")`) and peer (`r glue::glue("{format(des_peer$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_peer$prior_ses$fixed[3], digits = 2)}, {format(des_peer$prior_ses$fixed[4], digits = 2)}]")`) problems.
-Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (`r glue::glue("{format(des_social$sch_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$sch_ses$fixed[3], digits = 2)}, {format(des_social$sch_ses$fixed[4], digits = 2)}]")`)
+Student SES is slightly correlated with prosocial behaviour (`r glue::glue("{format(des_social$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$prior_ses$fixed[3], digits = 2)}, {format(des_social$prior_ses$fixed[4], digits = 2)}]")`) but more strongly and negatively correlated with conduct (`r glue::glue("{format(des_conduct$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_conduct$prior_ses$fixed[3], digits = 2)}, {format(des_conduct$prior_ses$fixed[4], digits = 2)}]")`) and peer (`r glue::glue("{format(des_peer$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_peer$prior_ses$fixed[3], digits = 2)}, {format(des_peer$prior_ses$fixed[4], digits = 2)}]")`) problems.
+Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (`r glue::glue("{format(des_social$sch_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$sch_ses$fixed[3], digits = 2)}, {format(des_social$sch_ses$fixed[4], digits = 2)}]")`).
 
@@
 School average SES negatively predicted conduct problems and positively predicted prosocial behavior for both parent and teacher reported social skills.
-School average SES also negatively predicted peer problems but only for parent reports.
+School average SES also negatively predicted peer problems.
+Results from either teacher or parent provided fairly consistent evidence of the influence of school SES context on social skills (or more specifically change in social skills from age 4 to age 8).
+Interestingly, the association of school average SES with social skills was similar in strength to the association of individual SES with social skills (see supplementary materials for full results).
+The effect sizes were equivalent to approximately a 10th of the median absolute deviation in social skills.
 
@@
 The interaction plots suggest that school context effects were particularly potent for disadvantaged children but school context had minimal impact on social skills for children from advantaged backgrounds.
+The plots suggest that for the very poorest schools in our sample, the average disadvantaged student would have levels of peer problems greater than the threshold for 'close to average' scores and thus potentially of concern [@youth2016scoring].
+This same child would be well within the 'close to average' band in a school with an average levels of SES.
 
@@
 Yet little research in this area has considered the potential influence of school context on social skill development [*cf.* @jerrim2019].
-This is a lacuna that our research sought to fill.
+This is a gap that our research sought to fill.
 Our research considered the association of school socioeconomic context with social skills in early elementary school.
@@
 Importantly, our research used a number of critical controls as well as prior social skills that helped us better identify the effect of school average socioeconomic status.
+Prior scocial skills meant that coefficents for school average SES related to change in social skills from age 4 (at or just before school entry) to age 8.
 
@@
 Our results do suggest that a low SES child who is enrolled in an advantaged school should expect to have similar levels of social skills as their high SES peers.
-The problem is that in countries like Australia low SES children are relatively rarely enrolled in advantaged schools.
+The problem is that in countries like Australia low SES children are considerably less likely to be enrolled in advantaged schools.
 As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA [@oecd2015].
@@
 Consistent with the literature we found that low SES children start school with lower social skills.
-In a country like Australia, whose schools are socially stratified, low SES children will tend to be schooled in more disadvantaged schools (see Figure 1).
-The net effect in such a system is that low SES children will tend to have their already lower social skills depressed by the school climate they are most likely to find themselves in.
+In a country like Australia, whose schools are socially stratified, low SES children will tend to be schooled in more disadvantaged schools.
+The net effect in such a system is that low SES children will tend to have their already lower starting school social skills further depressed by the school climate they are most likely to find themselves in.
 
@@
 This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents' rights to choose.
-However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a disadvantaged school.
+However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a more disadvantaged school.
 
@@
 As educational psychologists our study is framed in relation to assimilation effects in response to children's peer frames-of-reference.
-Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see <!--# @PHIL this citation wasn't in reference list -->Bowls & Gintis, 2001).
+Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures [see \@bowles2001].
 Identifying and comparing these mechanisms is an important future direction for research.

### 2020-10-29 manuscript.tex
full draft 2020-10-29

@@
   pdftitle={School Socioeconomic Status Context and Social Skills in Children},
-  pdfauthor={Philip Parker1 \& Taren Sanders1},
+  pdfauthor={Philip Parker1, Taren Sanders1, Jake Anders2, \& Michael Noetel1,3},
   pdflang={en-EN},
@@
 \title{School Socioeconomic Status Context and Social Skills in Children}
-\author{Philip Parker\textsuperscript{1} \& Taren Sanders\textsuperscript{1}}
+\author{Philip Parker\textsuperscript{1}, Taren Sanders\textsuperscript{1}, Jake Anders\textsuperscript{2}, \& Michael Noetel\textsuperscript{1,3}}
 \date{}
@@
 
-Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-Enter author note here.
+The authors made the following contributions.
+Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Jake Anders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Noetel: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
 
-The authors made the following contributions.
-Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
-
-Correspondence concerning this article should be addressed to Taren Sanders, 33 Berry St, North Sydney, NSW, 2060, Australia.
-E-mail: \href{mailto:taren.sanders@acu.edu.au}{\nolinkurl{taren.sanders@acu.edu.au}}
+Correspondence concerning this article should be addressed to Philip Parker, 33 Berry St, North Sydney, NSW, 2060, Australia.
+E-mail: \href{mailto:philip.parker@acu.edu.au}{\nolinkurl{philip.parker@acu.edu.au}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute for Positive Psychology and Education, Australian Catholic University\\\textsuperscript{2} UCL Institute of Education, UCL}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Institute for Positive Psychology and Education, Australian Catholic University\\\textsuperscript{2} UCL Institute of Education, UCL\\\textsuperscript{3} School of Health and Behavioural Sciences, Australian Catholic University}
 
@@
 Yet little research has considered how the school socio-economic context may differentially promote or hinder social skill development.
-In a representative longitudinal sample of 5361 Australian children, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers.
-All models controlled for prior social skills at age 4, and a wide range of covariates.
+In a representative longitudinal sample of Australian 8-year olds, we tested the association between school average socioeconomic status and social skills as reported by parents and teachers.
+All models controlled for prior social skills at age 4, and additional school and student covariates.
 We found that school context was associated with social skill assimilation: controlling for individual socioeconomic status, children in more advantaged schools had more prosocial behavior and fewer peer and conduct problems.
@@
 In response to this bleak picture, there has been a flurry of interest in social-emotional competencies driven by the research of Nobel prize-winning economist James Heckman.
-(Heckman, 2006) argues that research and intervention on reducing educational inequality have focused too narrowly on ``cognitive'' abilities.
+Heckman (2006) argues that research and intervention on reducing educational inequality have focused too narrowly on ``cognitive'' abilities (generally narrowly deinfed to include ability, IQ, and academic achievement).
 Yet cognitive abilities are not the only pathway through which low socioeconomic status may stifle educational and occupational attainment.
@@
 Australia is a useful country to focus on because Australia's Programme for International Student Assessment (PISA) Index of Social Inclusion score is relatively low (OECD, 2015).
-This means that Australian schools are socially stratified (Parker, Guo, \& Sanders, 2019; Philip D.
-Parker et al., 2018).
+This means that Australian schools are socially stratified (Parker, Guo, \& Sanders, 2019; Parker et al., 2018).
 Our major hypothesis is that school average socioeconomic status will have an assimilation effect on association with social skills at Year 3 (age 8) controlling for incoming social skills (age 4), individual socioeconomic status (age 4) and a range of demographic and achievement covariates\footnote{This assimilation effect will be indicated by a negative coefficient for peer problems and conduct problems regressed on school average socioeconomic status and a positive coefficient for prosocial behavior regressed on school average socioeconomic status.
@@
 We expect the assimilation effects to occur in the context of low SES children entering school with poorer social skills (de Laat et al., 2016; Garratt et al., 2017; McMunn et al., 2001; Rajmil et al., 2014).
-Given Australian schools are socially stratified (OECD, 2015; Parker et al., 2018), we also expect low SES children to attend more disadvantaged schools on average.
+Given Australian schools are socially stratified (OECD, 2015; Parker et al., 2018), we expect low SES children to attend more disadvantaged schools on average.
 
@@
 Together our total sample was 5440 (52.08\% boys) children in Year 3 at school (\textasciitilde age 8).
-In the vast majority of cases (98\%), parent data came from mothers.
-A participant flow diagram is present end in Figure~\ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
+In the vast majority of cases (98\% ), parent data came from mothers.
+A participant flow diagram is presented in Figure \ref{fig:consort} including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 
-Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (\emph{Strengths and difficulties questionnaire}, 1997).
-We explored social skills at age 4 (for prior social skills) and age 8 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}.
+Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (Goodman, 1997).
+We explored social skills at age 4-5 (for prior social skills) and age 8-9 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}.
 The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
@@
 Prosocial behaviors were heavily right censored.
-Censoring can be viewed as a special type of missing data where scores on \(y*\) below or above the bounds are curtailed to fit within the bounds resulting in the observed scores \(y\) (Gelman, Hill, \& Vehtari, 2020).
-Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable \(y*\) believed to underlie the censored variable (hence negative predicted scores were possible).
+Censoring can be viewed as a special type of missing data where scores on \(y*\) (the hypothesised true latent distribution of the variable) below or above the bounds are curtailed to fit within the bounds resulting in the observed scores \(y\) (Gelman, Hill, \& Vehtari, 2020).
+Models accounting for this censoring were used in all cases with results provided on the scale of the latent variable \(y*\) believed to underlie the censored variable (hence negative predicted scores or scores over 10 were possible).
 
@@
 Included in the analysis was the first principal component of the numeracy, grammar, spelling, reading, and writing scores from NAPLAN\footnote{Parallel analysis suggested a single principal component was sufficient to account for the variation in the NAPLAN scores.}.
-For verbal ability (as a proxy for cognitive school readiness) we used the Peabody Picture Vocabulary test (\emph{Peabody picture vocabulary test--fourth edition}, 2007) given to participants at age 4 (i.e., at or just before school commencement).
+For verbal ability (as a proxy for cognitive school readiness) we used the Peabody Picture Vocabulary test (Dunn \& Dunn, 2007) given to participants at age 4 (i.e., at or just before school commencement).
 Cohort was included as a covariate in all models as was measures of rural status and gender---all measured at age 4.
@@
 All analyses were done using Bayes via the BRMS package in R (Bürkner, 2017).
-The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before point and uncertainty estimates were extracted.
-Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantages (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students (Gelman et al., 2020).
+The Bayes models were run once for each imputation and then the resulting posterior samples were pooled before estimates and their uncertainty were extracted.
+Interactions and their associated uncertainties were plotted by taking marginal effect estimates from a random 500 posterior draws for disadvantaged (two standard deviations below the mean of SES), average, and advantaged (two standard deviations above the mean of SES) students (Gelman et al., 2020).
 
 All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression.
-Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails ((see Kleiber \& Zeileis, 2008).
+Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see Kleiber \& Zeileis, 2008).
 As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
@@
 We clearly demarcate the units of analysis in tables and figures.
-I general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
-We retain the SDQ scores on the original scale because of the presence of norms {[}@youth2016scoring{]}.
-Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the `close to average' group and may thus be a potential concern (Mind, 2016).
+In general the SDQ measures are retained on the original scale and all other continuous variables are z-scored.
+We retain the SDQ scores on the original scale because of the presence of norms (Youth in Mind, 2016).
+Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the `close to average' group and may thus be a potential concern (Youth in Mind, 2016).
+
+The models are fit using the following formula:
+
+\[
+\begin{split}
+y^{*}_{i} \sim N(\alpha_{j[i]} + X_i\beta + \sigma^{2}_{y}), for~i = 1,...,n \\
+\alpha_{j} \sim N(U_j \gamma, \sigma^{2}_{\alpha}), for~j = 1,...k
+\end{split}
+\]
+
+Here \(X\) is a matrix of individual level predictors for student \(i\) including individual SES, prior social skills, gender, cohort, geographic location, and an NAPLAN achievement index.
+Prior social skills is particularly critical as this provide stronger conclusions that school context rather than selection effects account for the influence of school average SES on student social skills.
+Put simply the estimate of school average achievement can be interpreted as the influence of school context on the change in children's social skills from age 4 to age 8.
+\(U_j\) is a matrix of school level predictors including school section and the aforementioned critical variable \emph{school SES}.
+Note that we predict \(y^{*}_{i}\) which is the latent continuous variable underlying the observed variables which are defined as:
+
+\[
+\begin{split}
+y_i =  \begin{cases}      
+            y*_{i} & \text{if $y*_{i} < 0$ or $y*_{i} > 10 $}\\      
+            0 & \text{if $y*_{i} < 0$}\\      
+            10 & \text{if $y*_{i} > 10$}    
+        \end{cases}
+\end{split}
+\]
 
@@
 In order to provide a context for the main analyses we first aimed to show: a) that children from advantaged backgrounds enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
-Student SES is slightly correlated with prosocial behaviour (0.05 95\% CI{[}0.023, 0.078{]}) but more strongly and negatively with conduct (-0.18 95\% CI{[}-0.2, -0.15{]}) and peer (-0.16 95\% CI{[}-0.19, -0.13{]}) problems.
-Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (0.48 95\% CI{[}0.46, 0.51{]})
+Student SES is slightly correlated with prosocial behaviour (0.05 95\% CI{[}0.023, 0.078{]}) but more strongly and negatively correlated with conduct (-0.18 95\% CI{[}-0.2, -0.15{]}) and peer (-0.16 95\% CI{[}-0.19, -0.13{]}) problems.
+Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (0.48 95\% CI{[}0.46, 0.51{]}).
 
@@
 The interaction plots suggest that school context effects were particularly potent for disadvantaged children but school context had minimal impact on social skills for children from advantaged backgrounds.
+The plots suggest that for the very poorest schools in our sample, the average disadvantaged student would have levels of peer problems greater than the threshold for `close to average' scores and thus potentially of concern (Youth in Mind, 2016).
+This same child would be well within the `close to average' band in a school with an average levels of SES.
 
@@
 Yet little research in this area has considered the potential influence of school context on social skill development (\emph{cf.} Jerrim \& Sims, 2019).
-This is a lacuna that our research sought to fill.
+This is a gap that our research sought to fill.
 Our research considered the association of school socioeconomic context with social skills in early elementary school.
@@
 Our results do suggest that a low SES child who is enrolled in an advantaged school should expect to have similar levels of social skills as their high SES peers.
-The problem is that in countries like Australia low SES children are relatively rarely enrolled in advantaged schools.
+The problem is that in countries like Australia low SES children are considerably less likely to be enrolled in advantaged schools.
 As we noted in the introduction Australia has relatively low levels of social inclusion as measured by PISA (OECD, 2015).
@@
 
-Parker et al.~(2019, 2018 ; 2016; 2018) have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
+Parker et al.~(2019, 2018; 2016; 2018) have suggested that empirical evidence shows greater school choice at the country level is related to poorer average ability levels, lower aspirations, and paradoxical effects on psychological factors like motivation and self-concept that appears to have negative consequences for all children.
 They argue that empirical evidence suggests policy should encourage school selection policies that maximize within school heterogeneity.
 This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents' rights to choose.
-However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a disadvantaged school.
+However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a more disadvantaged school.
 
@@
 As educational psychologists our study is framed in relation to assimilation effects in response to children's peer frames-of-reference.
-Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures (see Bowls \& Gintis, 2001).
+Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures {[}see @bowles2001{]}.
 Identifying and comparing these mechanisms is an important future direction for research.
@@
 
+\leavevmode\hypertarget{ref-dunn2007}{}%
+Dunn, L., \& Dunn, D. (2007). \emph{Peabody picture vocabulary test--fourth edition}. American Psychological Association. \url{https://doi.org/10.1037/t15144-000}
+
 \leavevmode\hypertarget{ref-durlak2011}{}%
@@
 Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.
+
+\leavevmode\hypertarget{ref-goodman1997}{}%
+Goodman, R. (1997). \emph{Strengths and difficulties questionnaire}. American Psychological Association. \url{https://doi.org/10.1037/t00540-000}
 
@@
 
-\leavevmode\hypertarget{ref-youth2016scoring}{}%
-Mind, Y.
-in.
-(2016).
-\emph{Scoring the strengths \& difficulties questionnaire for age 4--17 or 18+}.
-Retrieved from \url{https://www.ehcap.co.uk/content/sites/ehcap/uploads/NewsDocuments/236/SDQEnglishUK4-17scoring-1.PDF}
-
 \leavevmode\hypertarget{ref-mussweiler2004}{}%
@@
 
-\leavevmode\hypertarget{ref-dunn2007}{}%
-\emph{Peabody picture vocabulary test--fourth edition}. (2007). American Psychological Association. \url{https://doi.org/10.1037/t15144-000}
-
 \leavevmode\hypertarget{ref-rajmil2014}{}%
@@
 
-\leavevmode\hypertarget{ref-goodman1997}{}%
-\emph{Strengths and difficulties questionnaire}. (1997). American Psychological Association. \url{https://doi.org/10.1037/t00540-000}
-
 \leavevmode\hypertarget{ref-washbrook2011}{}%
@@
 Retrieved from \url{https://www.resolutionfoundation.org/app/uploads/2014/08/On-your-marks.pdf}
+
+\leavevmode\hypertarget{ref-youth2016scoring}{}%
+Youth in Mind.
+(2016).
+\emph{Scoring the strengths \& difficulties questionnaire for age 4--17 or 18+}.
+Retrieved from \url{https://www.ehcap.co.uk/content/sites/ehcap/uploads/NewsDocuments/236/SDQEnglishUK4-17scoring-1.PDF}
 \end{cslreferences}

### 2020-10-29 manuscript.Rmd
Addition of all authors

@@
   affiliation: '2'
-  # corresponding: no
-  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
-  # email: taren.sanders@acu.edu.au
+  role:
+  - Conceptualization
+  - Writing - Original Draft Preparation
+  - Writing - Review & Editing
+- name: Nikki Shure
+  affiliation: '2'
+  role:
+  - Conceptualization
+  - Writing - Original Draft Preparation
+  - Writing - Review & Editing
+- name: John Jerrim
+  affiliation: '2'
   role:
@@
   affiliation: '1,3'
-  # corresponding: no
-  # address: 33 Berry St, North Sydney, NSW, 2060, Australia
-  # email: taren.sanders@acu.edu.au
+  role:
+  - Conceptualization
+  - Writing - Original Draft Preparation
+  - Writing - Review & Editing
+- name: Joseph Ciarrochi
+  affiliation: '1'
   role:

### 2020-10-29 manuscript.tex
Addition of all authors

@@
   pdftitle={School Socioeconomic Status Context and Social Skills in Children},
-  pdfauthor={Philip Parker1, Taren Sanders1, Jake Anders2, \& Michael Noetel1,3},
+  pdfauthor={Philip Parker1, Taren Sanders1, Jake Anders2, Nikki Shure2, John Jerrim2, Michael Noetel1,3, \& Joseph Ciarrochi1},
   pdflang={en-EN},
@@
 \title{School Socioeconomic Status Context and Social Skills in Children}
-\author{Philip Parker\textsuperscript{1}, Taren Sanders\textsuperscript{1}, Jake Anders\textsuperscript{2}, \& Michael Noetel\textsuperscript{1,3}}
+\author{Philip Parker\textsuperscript{1}, Taren Sanders\textsuperscript{1}, Jake Anders\textsuperscript{2}, Nikki Shure\textsuperscript{2}, John Jerrim\textsuperscript{2}, Michael Noetel\textsuperscript{1,3}, \& Joseph Ciarrochi\textsuperscript{1}}
 \date{}
@@
 The authors made the following contributions.
-Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Jake Anders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Noetel: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
+Philip Parker: Conceptualization, Formal Analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Taren Sanders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Jake Anders: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Nikki Shure: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; John Jerrim: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Michael Noetel: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Joseph Ciarrochi: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
 

### 2020-10-30 manuscript.Rmd
taren's changes incorporated

@@
 We model social skills using both parent and teacher reports.
-Measures of SES at the student level are taken from in-person interviews with parents <!--# @taren.sanders@acu.edu.au is this true --> and school SES (as well as student achievement) is taken from government administrative records.
+Measures of SES at the student level are taken from in-person interviews with parents and school SES (as well as student achievement) is taken from government administrative records.
 
@@
 We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC).
-LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004.
+LSAC is a government-run study of a representative sample of Australian children who were zero-one (B-cohort) or four-five (K-Cohort) years of age in 2004.
 Both cohorts of children have been followed every two years since [@aifs2015].
@@
 Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ [@goodman1997].
-We explored social skills at age 4-5 (for prior social skills) and age 8-9 (as primary outcomes) as reported by the child's *parent* and the child's *teacher*.
+We explored social skills at age 4-5 (for prior social skills; parent report only) and age 8-9 (as primary outcomes) as reported by the child's *parent* and the child's *teacher*.
 The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
@@
 However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable.
-Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample.
+Attrition was offset by applying combined sample and attrition weights constructed for the age 8 sample.
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
@@
 Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails [see @kleiber2008].
-As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
+As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the left and the right in all models.
 The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
@@
 \begin{split}
-y^{*}_{i} \sim N(\alpha_{j[i]} + X_i\beta + \sigma^{2}_{y}), for~i = 1,...,n \\
+y^{*}_{i} \sim N(\alpha_{j[i]} + X_i\beta, \sigma^{2}_{y}), for~i = 1,...,n \\
 \alpha_{j} \sim N(U_j \gamma, \sigma^{2}_{\alpha}), for~j = 1,...k
@@
 Put simply the estimate of school average achievement can be interpreted as the influence of school context on the change in children's social skills from age 4 to age 8.
-$U_j$ is a matrix of school level predictors including school section and the aforementioned critical variable *school SES*.
+$U$ is a matrix of school level predictors for school $j$ including school section and the aforementioned critical variable *school SES*.
 Note that we predict $y^{*}_{i}$ which is the latent continuous variable underlying the observed variables which are defined as:
@@
 Importantly, our research used a number of critical controls as well as prior social skills that helped us better identify the effect of school average socioeconomic status.
-Prior scocial skills meant that coefficents for school average SES related to change in social skills from age 4 (at or just before school entry) to age 8.
+Prior social skills meant that coefficients for school average SES related to change in social skills from age 4 (at or just before school entry) to age 8.
 
@@
 This would require considerable state intervention to achieve and may thus impose unreasonable restrictions on parents' rights to choose.
-However, it is worth noting that our counterfactual analysis (Table 4) shows that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a more disadvantaged school.
+However, it is worth noting that high SES children appear to not suffer to any notable degree, in relation to social skills, by being enrolled in a more disadvantaged school.
 

### 2020-10-30 manuscript.tex
with taren's comments

@@
 We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC).
-LSAC is a government-run study of a representative sample of Australian children between who were zero-one (B-cohort) and four-five (K-Cohort) years of age in 2004.
+LSAC is a government-run study of a representative sample of Australian children who were zero-one (B-cohort) or four-five (K-Cohort) years of age in 2004.
 Both cohorts of children have been followed every two years since (AIFS, 2015).
@@
 Social skills were estimated using the peer problems, conduct problems, and prosocial behavior component scores from the SDQ (Goodman, 1997).
-We explored social skills at age 4-5 (for prior social skills) and age 8-9 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}.
+We explored social skills at age 4-5 (for prior social skills; parent report only) and age 8-9 (as primary outcomes) as reported by the child's \emph{parent} and the child's \emph{teacher}.
 The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
@@
 However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable.
-Attrition was offset by using applying combined sample and attrition weights constructed for the age 8 sample.
+Attrition was offset by applying combined sample and attrition weights constructed for the age 8 sample.
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14\%.
@@
 Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails (see Kleiber \& Zeileis, 2008).
-As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the right and left in all models.
+As scores on the social skills variables have both a floor of zero and a ceiling of 10, we included censoring on both the left and the right in all models.
 The multilevel nature of the data (children nested within schools) was addressed by the inclusion of a random intercept for schools.
@@
 Put simply the estimate of school average achievement can be interpreted as the influence of school context on the change in children's social skills from age 4 to age 8.
-\(U_j\) is a matrix of school level predictors including school section and the aforementioned critical variable \emph{school SES}.
+\(U\) is a matrix of school level predictors for school \(j\) including school section and the aforementioned critical variable \emph{school SES}.
 Note that we predict \(y^{*}_{i}\) which is the latent continuous variable underlying the observed variables which are defined as:

### 2020-11-02 manuscript.tex
Add supp material
Did the following:

* Added `supplementary_materials.Rmd`
* Modified function to produce full model outputs
* Added function to knit both manuscript and supp materials
* Updated pipeline with new functions
* Removed all old versions of codebook that have names in them

Should close #5

@@
 \centering
-\includegraphics{/Users/philipparker/Dropbox/Projects_Research/CURRENT_WORKING/social_skills_2020/fig/flow.pdf}
+\includegraphics{/Users/tasanders/GitHub/social_skills_2020/fig/flow.pdf}
 \caption{\label{fig:consort}Participant flow diagramme}
@@
 
-Parents and teachers had moderate agreement for children at age 8 with correlations ranging from \emph{r} = 0.41 95\% CI{[}0.38, 0.43{]} for peer and \emph{r} = 0.4 95\% CI{[}0.37, 0.42{]} for conduct problems to \emph{r} = 0.29 95\% CI{[}0.26, 0.31{]} for prosocial behavior.
+Parents and teachers had moderate agreement for children at age 8 with correlations ranging from \emph{r} = 0.41 95\% CI{[}0.38, 0.43{]} for peer and \emph{r} = 0.39 95\% CI{[}0.37, 0.42{]} for conduct problems to \emph{r} = 0.29 95\% CI{[}0.26, 0.31{]} for prosocial behavior.
 
@@
 \midrule
-conduct & teacher & $-0.152$ & $-0.261$ & $-0.046$ \\ 
-social & teacher & $0.147$ & $0.056$ & $0.241$ \\ 
-peer & teacher & $-0.115$ & $-0.204$ & $-0.024$ \\ 
-conduct & parent & $-0.107$ & $-0.169$ & $-0.045$ \\ 
-social & parent & $0.087$ & $0.018$ & $0.156$ \\ 
-peer & parent & $-0.135$ & $-0.205$ & $-0.063$ \\ 
+conduct & teacher & $-0.154$ & $-0.264$ & $-0.043$ \\ 
+social & teacher & $0.148$ & $0.054$ & $0.242$ \\ 
+peer & teacher & $-0.116$ & $-0.204$ & $-0.029$ \\ 
+conduct & parent & $-0.108$ & $-0.170$ & $-0.046$ \\ 
+social & parent & $0.087$ & $0.018$ & $0.155$ \\ 
+peer & parent & $-0.134$ & $-0.203$ & $-0.063$ \\ 
 \bottomrule
@@
 \midrule
-conduct & teacher & School Average SES & $-0.160$ & $-0.270$ & $-0.053$ \\ 
-conduct & teacher & School Average SES by Student SES & $0.151$ & $0.064$ & $0.240$ \\ 
-social & teacher & School Average SES & $0.156$ & $0.059$ & $0.252$ \\ 
-social & teacher & School Average SES by Student SES & $-0.094$ & $-0.172$ & $-0.017$ \\ 
-peer & teacher & School Average SES & $-0.126$ & $-0.216$ & $-0.036$ \\ 
-peer & teacher & School Average SES by Student SES & $0.117$ & $0.044$ & $0.191$ \\ 
-conduct & parent & School Average SES & $-0.112$ & $-0.175$ & $-0.048$ \\ 
-conduct & parent & School Average SES by Student SES & $0.058$ & $0.006$ & $0.111$ \\ 
-social & parent & School Average SES & $0.090$ & $0.020$ & $0.160$ \\ 
-social & parent & School Average SES by Student SES & $-0.031$ & $-0.087$ & $0.027$ \\ 
-peer & parent & School Average SES & $-0.144$ & $-0.217$ & $-0.074$ \\ 
-peer & parent & School Average SES by Student SES & $0.117$ & $0.058$ & $0.176$ \\ 
+conduct & teacher & School Average SES & $-0.161$ & $-0.268$ & $-0.054$ \\ 
+conduct & teacher & School Average SES by Student SES & $0.150$ & $0.063$ & $0.240$ \\ 
+social & teacher & School Average SES & $0.157$ & $0.063$ & $0.250$ \\ 
+social & teacher & School Average SES by Student SES & $-0.095$ & $-0.172$ & $-0.020$ \\ 
+peer & teacher & School Average SES & $-0.126$ & $-0.218$ & $-0.035$ \\ 
+peer & teacher & School Average SES by Student SES & $0.116$ & $0.046$ & $0.189$ \\ 
+conduct & parent & School Average SES & $-0.112$ & $-0.176$ & $-0.047$ \\ 
+conduct & parent & School Average SES by Student SES & $0.058$ & $0.006$ & $0.110$ \\ 
+social & parent & School Average SES & $0.090$ & $0.023$ & $0.157$ \\ 
+social & parent & School Average SES by Student SES & $-0.031$ & $-0.087$ & $0.029$ \\ 
+peer & parent & School Average SES & $-0.143$ & $-0.213$ & $-0.074$ \\ 
+peer & parent & School Average SES by Student SES & $0.117$ & $0.057$ & $0.179$ \\ 
 \bottomrule

### 2020-11-02 supplementary_materials.Rmd
Add supp material
Did the following:

* Added `supplementary_materials.Rmd`
* Modified function to produce full model outputs
* Added function to knit both manuscript and supp materials
* Updated pipeline with new functions
* Removed all old versions of codebook that have names in them

Should close #5

---
title: "Supplementary Materials: School Socioeconomic Status Context and Social Skills in Children"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(targets)
library(scales)
library(tidyverse)
library(DT)

```

# Full Linear Model

```{r}
tar_load(supp_linear)
supp_linear %>% 
  datatable(filter = 'top', caption = "Table S1. Full Linear Model Outputs")
```

# Full Interaction Model

```{r}
tar_load(supp_interaction)
supp_interaction %>%
  datatable(filter = 'top', caption = "Table S2. Full Interaction Model Outputs")
```

# Codebook

```{r, child="data/social_skills_codebook.Rmd"}

```

### 2020-11-02 R/functions.R

@@ -146,12 +146,22 @@ data <- function(data_list = list()){
   return(child_data)
 }
 
-make_codebook <- function(data){
-  dataMaid::makeCodebook(data,
-                         file = here::here("data",
-                                     glue::glue("{Sys.Date()}_codebook.Rmd")),
-                         replace=TRUE)
-  return(here::here("data",glue::glue("{Sys.Date()}_codebook.pdf")))
+make_codebook <- function(data) {
+  dataMaid::makeCodebook(
+    data,
+    file = here::here("data",
+                      glue::glue("social_skills_codebook.Rmd")),
+    replace = TRUE,
+    render = FALSE,
+    standAlone = FALSE
+  )
+  # Remove last lines of codebook to maintain blinding
+  codebook <-
+    readLines(here::here("data", glue::glue("social_skills_codebook.Rmd")))
+  codebook.len <- length(codebook)
+  writeLines(codebook[-((codebook.len - 22):codebook.len)], here::here("data", glue::glue("social_skills_codebook.Rmd")))
+  
+  return(here::here("data", glue::glue("social_skills_codebook.Rmd")))
 }
 
 consort_flow <- function(data_list = list(age_4_data, age_8_data, child_achievement_data, school_achievement_data)) {
@@ -439,39 +449,48 @@ plots <- function(model=conduct_teach){
 }
 
 
-full_model_output <- function(model = conduct_teach) {
-  
-  nm <- deparse(substitute(model)) %>%
-    str_split(., "_",simplify = TRUE) %>% 
-    str_to_sentence()
-  
-  source <- ifelse(nm[2] == "teach", "Teacher", "Parent")
-  outcome <- case_when(
-    nm[1] == 'Conduct' ~ "Conduct Problems",
-    nm[1] == 'Social' ~ "Prosociality",
-    nm[1] == 'Peer' ~ "Peer Problems",
-  )
-  
-  out = model %>%
-    map_dfr(broom.mixed::tidy, .id = "model") %>%
-    mutate(term = case_when(
-      term == '(Intercept)' ~ "Intercept (Units: Censored 0-10)",
-      term == 'ses_sch' ~ "School Average SES (Units: SD)",
-      term == 'ses' ~ "SES (Units: SD)",
-      term == 'prior' ~ glue::glue("Prior {outcome} (Units: 0-10)") %>% as.character(),
-      term == 'cohortK' ~ "Cohort K",
-      term == 'gendergirl' ~ "Girl",
-      term == 'geourban' ~ "Urban",
[…]

### 2020-11-02 _targets.R

@@ -1,4 +1,5 @@
 library(targets)
+library(tarchetypes)
 
 source(here::here("R","functions.R"))
 options(tidyverse.quiet = TRUE)
@@ -115,5 +116,36 @@ tar_pipeline(
   tar_target(
     des_social,
     ses_beta(d = data_imp, outcome = 'social')
-  )
+  ),
+  tar_target(supp_linear,
+             full_model_output(
+               list(
+                 conduct_teach,
+                 social_teach,
+                 peer_teach,
+                 conduct_par,
+                 social_par,
+                 peer_par
+               ),
+               model_type = 'linear'
+             )),
+  tar_target(supp_interaction,
+             full_model_output(
+               list(
+                 conduct_teach,
+                 social_teach,
+                 peer_teach,
+                 conduct_par,
+                 social_par,
+                 peer_par
+               ),
+               model_type = 'interaction'
+             )
+  ),
+  tar_render(manuscript,
+             "manuscript.Rmd",
+             deployment = "master"),
+  tar_render(supplementary_material,
+             "supplementary_materials.Rmd",
+             deployment = "master")
 )

### 2020-11-05 manuscript.Rmd
clean full draft

@@
   - Formal Analysis
+  - Data curation
+  - Visualization
   - Writing - Original Draft Preparation
@@
   - Conceptualization
+  - Data curation
+  - Visualization
   - Writing - Original Draft Preparation
@@
   role:
-  - Conceptualization
+  - Visualization
   - Writing - Original Draft Preparation
@@
   role:
-  - Conceptualization
   - Writing - Original Draft Preparation
@@
 authornote: |
+  Code for this paper is avaliable from [OSF](https://osf.io/q26ej/?view_only=73f60cea7a9c4e6f8affc412b30df118).
+Data is avaliable from the [Australian Data Archive](https://dataverse.ada.edu.au/dataverse/ada?q=LSAC).
 abstract: |2
@@
 keywords: social skills; assimilation effects; socioeconomic status; school context
-wordcount: X
+wordcount: 4627
 bibliography: r-references.bib
@@
 Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
-All other variables had 6% missing data or less (see code book in supplementary materials <!--# @philip.parker@acu.edu.au Convert to OSF link when ready -->).
+All other variables had 6% missing data or less (see code book in [supplementary materials](http://blindedforreview.com/2020_social_skills/supplementary_materials.html)).
 To account for this, we constructed five imputed datasets using a bootstrapped expectation maximization procedure from the Amelia II package [@honaker2011].
@@
 
-All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression.
+All models in the paper that predicted social skills were estimated taking into account the censoring of the social skills means using a Bayes version of Tobit regression (we used default weakly informed priors derived from the data for all models).
 Tobit models can be used when seeking to model outcome variables that are censored on the left, right, or both tails [see @kleiber2008].
@@
 Peer and conduct problems scores of 3 or greater and prosocial scores of 7 or lower place children beyond the 'close to average' group and may thus be a potential concern [@youth2016scoring].
-<!--# @everyone We could use post-stratification weights to provide 'clinically' relevant proportions of the population for different levels of SES and school SES -->
 
@@
 
+All scripts used to produce these results can be found on the [OSF project](https://osf.io/q26ej/?view_only=73f60cea7a9c4e6f8affc412b30df118) associated with this paper.
+Data can be applied for from the Australian Data Archive Dataverse [website](https://dataverse.ada.edu.au/dataverse/ada?q=LSAC).
+
 # Results
@@
 
+(ref:table-linear) Effect of School Average SES on Social Skills.
+
 ```{r table-1, caption='Effect of School Average SES on Social Skills',results='asis'}
 tar_load(linear_table)
-library(gt)
-library(tidyMB)
 
 linear_table %>%
-  gt() %>%
-  fmt_number(columns = 3:5, decimals = 3) %>%
-  apa_style(title = 'Table 1', subtitle = 'Effect of School Average SES on Social Skills') %>%
-  as_latex()
+  set_names(str_to_sentence(names(.))) %>%
+  printnum() %>%
+  apa_table(
+  .,
+  , caption = "(ref:table-linear)"
+  , note = "Results are in unit changes on a 1-10 (Censored) scale for social skills for a standard deviation increase in school SES.
+Full results in supplementary materials.
+Conduct = conduct problems, peer = peer problems, social = prosocial.
+Teacher = teacher reported, parent = parent reported."
+)
+
+#library(gt)
+#library(tidyMB)
+
+#linear_table %>%
+#  gt() %>%
+#  fmt_number(columns = 3:5, decimals = 3) %>%
+#  apa_style(title = 'Table 1', subtitle = 'Effect of School Average SES on Social Skills') %>%
+#  as_latex()
 ```
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results for school average SES are presented under in Table \@ref(tab:table-1) (the full model results are in supplementary materials).
+Results for school average SES are presented under in Table \@ref(table-linear) (the full model results are in [supplementary materials](http://blindedforreview.com/2020_social_skills/supplementary_materials.html)).
 Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (at or just prior to school enrollment) to age 8 (Year 3).
@@
 
+(ref:table-quadratic) Effect of School Average SES by Student SES on Social Skills.
+
 ```{r table-2, caption='Effect of School Average SES by Student SES on Social Skills',results='asis'}
 tar_load(interaction_table)
-library(gt)
-library(tidyMB)
 
 interaction_table %>%
-  gt() %>%
-  fmt_number(columns = 4:6, decimals = 3) %>%
-  apa_style(title = 'Table 2',
-            subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
-  as_latex()
+  set_names(str_to_sentence(names(.))) %>%
+  printnum() %>%
+  apa_table(
+  .,
+  , caption = "(ref:table-quadratic)"
+  , note = "Results are in unit changes on a 1-10 (Censored) scale for social skills for a standard deviation increase in school SES.
+Full results in supplementary materials.
+Conduct = conduct problems, peer = peer problems, social = prosocial.
+Teacher = teacher reported, parent = parent reported."
+)
+
+#library(gt)
+#library(tidyMB)
+#interaction_table %>%
+#  gt() %>%
+#  fmt_number(columns = 4:6, decimals = 3) %>%
+#  apa_style(title = 'Table 2',
+#            subtitle = 'Effect of School Average SES by Student SES on Social Skills') %>%
+#  as_latex()
 ```
@@
 
-The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \@ref(tab:table-2)).
+The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \@ref(table-quadratic)).
 Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior.
@@
 The interaction plots suggest that school context effects were particularly potent for disadvantaged children but school context had minimal impact on social skills for children from advantaged backgrounds.
-The plots suggest that for the very poorest schools in our sample, the average disadvantaged student would have levels of peer problems greater than the threshold for 'close to average' scores and thus potentially of concern [@youth2016scoring].
+The plots suggest that for the very poorest schools in our sample, a disadvantaged student would have levels of peer problems greater than the threshold for 'close to average' scores and thus potentially of concern [@youth2016scoring].
 This same child would be well within the 'close to average' band in a school with an average levels of SES.
 
-```{r peer, fig.cap = 'Interaction Plot for Peer Problems'}
+```{r peer, fig.cap = 'School average SES by Student SES on Peer Problems. Includes 95% CIs.', fig.height = 6}
 library(patchwork)
@@
 
-```{r conduct, fig.cap = 'Interaction Plot for Conduct Problems'}
+```{r conduct, fig.cap = 'School average SES by Student SES on Conduct Problems. Includes 95% CIs.', fig.height = 6}
 library(patchwork)
@@
 
-```{r social, fig.cap = 'Interaction Plot for Prosociality'}
+```{r social, fig.cap = 'School average SES by Student SES on  Prosociality. Includes 95% CIs..', fig.height = 6}
 library(patchwork)

### 2020-11-05 R/functions.R

@@ -415,7 +415,7 @@ plots <- function(model=conduct_teach){
     str_split(., "_",simplify = TRUE) %>% 
     str_to_sentence()
   
-  source <- ifelse(nm[2] == "teach", "Teacher", "Parent")
+  source <- ifelse(nm[2] == "Teach", "Teacher", "Parent")
   outcome <- case_when(
     nm[1] == 'Conduct' ~ "Conduct Problems",
     nm[1] == 'Social' ~ "Prosociality",
@@ -425,12 +425,10 @@ plots <- function(model=conduct_teach){
   conditions <- data.frame(ses = c(-2, 0, 2))
   
   p = plot(brms::conditional_effects(model$interaction,'ses_sch:ses',
-                                 spaghetti = T, nsamples = 500, 
+                                 #spaghetti = T, nsamples = 500, 
                                  int_conditions = conditions))[[1]] +
-    labs(#title = glue::glue('{source} Reported {outcome}'),
-         #subtitle = "<span style='color:red'>High (+2 SD)</span>, <span style='color:green'>Average</span>, and <span style='color:blue'>Low (-2 SD)</span> SES Children",
-         color = "SES (Units: SD)",
-         y = "Outcome (Units: censorded 0-10)",
+    labs(color = "SES (Units: SD)",
+         y = glue::glue('{source} Reported {outcome} (Units: censorded 0-10)'),
          x = "School Average SES (Units: SD)") +
     tidyMB::theme_mb() +
     theme(

### 2020-11-05 R/make.R

@@ -7,3 +7,4 @@ tar_make()
 # Visualize status of the project
 # tar_glimpse()
 
+

### 2020-11-06 manuscript.Rmd
Ed Researcher Version

@@
 wordcount: 4627
+header-includes:
+  - \usepackage{hyperref}
 bibliography: r-references.bib
+urlcolor: blue
 floatsintext: no
@@
 linenumbers: yes
-mask: no
+mask: yes
 draft: no
@@
 
-```{r consort,fig.height=4,fig.width=2, fig.cap = 'Participant flow diagramme'}
-knitr::include_graphics(here::here('fig','flow.pdf'))
-```
-
 We use data for children, their parents, and their teachers from the B-Cohort and K-Cohort of the Longitudinal Study of Australian Children (LSAC).
@@
 Together our total sample was `r nrow(data)` (`r boys` boys) children in Year 3 at school (\~age 8).
-In the vast majority of cases (98% <!--# @philip.parker@acu.edu.au should extract this from data rather than previous research?
--->), parent data came from mothers.
-A participant flow diagram is presented in Figure \@ref(fig:consort) including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
+In the vast majority of cases (98%), parent data came from mothers.
+A participant flow diagram is presented in [supplementary materials](http://blindedforreview.com/2020_social_skills/supplementary_materials.html) including the number of participants excluded because there were a) not in Year 3 during the age 8 LSAC data collection sweep; b) were not eligible to participate NAPLAN (e.g., due to intellectual disability; or c) were not enrolled in school (e.g., were home schooled).
 
@@
 We explored social skills at age 4-5 (for prior social skills; parent report only) and age 8-9 (as primary outcomes) as reported by the child's *parent* and the child's *teacher*.
-The SDQ asks respondents to rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
+The SDQ asks respondents tfo rate a child's behavior in the last six months on a 3-point scale (not true, somewhat true, and certainly true).
 Questions ask about the child's peer problems ("Rather solitary, tends to play alone"), conduct problems ("Often fights with other children or bullies them"), and prosociality ("Is kind to younger children").
@@
 There was a relatively small amount of missing data for most variables.
-However, given the data were longitudinal and included both linked administrative records and data collected from both the child's teacher and parent, missing data was inevitable.
-Attrition was offset by applying combined sample and attrition weights constructed for the age 8 sample.
-Remaining missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
+However, given the data required responses from both parents and teachers and were linked to administrative records and data collected from both the child's teacher and parent, missing data was inevitable.
+Missing data was small with the largest missing data proportion for teacher reported social skills at 14%.
 All other variables had 6% missing data or less (see code book in [supplementary materials](http://blindedforreview.com/2020_social_skills/supplementary_materials.html)).
@@
 Prior social skills is particularly critical as this provide stronger conclusions that school context rather than selection effects account for the influence of school average SES on student social skills.
-Put simply the estimate of school average achievement can be interpreted as the influence of school context on the change in children's social skills from age 4 to age 8.
+Put simply the estimate of school average SES can be interpreted as the influence of school context on the change in childrens' social skills from age 4 to age 8.
 $U$ is a matrix of school level predictors for school $j$ including school section and the aforementioned critical variable *school SES*.
@@
 In order to provide a context for the main analyses we first aimed to show: a) that children from advantaged backgrounds enter school with better social skills (as measured by their parents) and b) that children in Australia tend to be schooled in socially stratified schools.
-Student SES is slightly correlated with prosocial behaviour (`r glue::glue("{format(des_social$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$prior_ses$fixed[3], digits = 2)}, {format(des_social$prior_ses$fixed[4], digits = 2)}]")`) but more strongly and negatively correlated with conduct (`r glue::glue("{format(des_conduct$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_conduct$prior_ses$fixed[3], digits = 2)}, {format(des_conduct$prior_ses$fixed[4], digits = 2)}]")`) and peer (`r glue::glue("{format(des_peer$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_peer$prior_ses$fixed[3], digits = 2)}, {format(des_peer$prior_ses$fixed[4], digits = 2)}]")`) problems.
+Student SES is slightly correlated with prior prosocial behaviour (`r glue::glue("{format(des_social$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$prior_ses$fixed[3], digits = 2)}, {format(des_social$prior_ses$fixed[4], digits = 2)}]")`) but more strongly and negatively correlated with prior conduct (`r glue::glue("{format(des_conduct$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_conduct$prior_ses$fixed[3], digits = 2)}, {format(des_conduct$prior_ses$fixed[4], digits = 2)}]")`) and peer (`r glue::glue("{format(des_peer$prior_ses$fixed[1], digits = 2)} 95% CI[{format(des_peer$prior_ses$fixed[3], digits = 2)}, {format(des_peer$prior_ses$fixed[4], digits = 2)}]")`) problems.
 Children from lower SES backgrounds also tended to enter schools with children from similar backgrounds (`r glue::glue("{format(des_social$sch_ses$fixed[1], digits = 2)} 95% CI[{format(des_social$sch_ses$fixed[3], digits = 2)}, {format(des_social$sch_ses$fixed[4], digits = 2)}]")`).
+Put simply, disadvantaged children tended to enter school with lower social skills and the school they entered tended poorer on average.
 
@@
 Conduct = conduct problems, peer = peer problems, social = prosocial.
-Teacher = teacher reported, parent = parent reported."
+Teacher = teacher reported, parent = parent reported.",
+  booktabs = TRUE
 )
@@
 We next predicted social skills with school average SES controlling for individual SES, a range of demographic covariates and academic performance measures, and social skills at age 4.
-Results for school average SES are presented under in Table \@ref(table-linear) (the full model results are in [supplementary materials](http://blindedforreview.com/2020_social_skills/supplementary_materials.html)).
+Results for school average SES are presented under in Table \@ref(tab:table-1) (the full model results are in [supplementary materials](http://blindedforreview.com/2020_social_skills/supplementary_materials.html)).
 Because we control for prior social skills, the effects of school average SES in these models can be interpreted as predicting change or development in social skills from age 4 (at or just prior to school enrollment) to age 8 (Year 3).
@@
 Interestingly, the association of school average SES with social skills was similar in strength to the association of individual SES with social skills (see supplementary materials for full results).
-The effect sizes were equivalent to approximately a 10th of the median absolute deviation in social skills.
+The effect sizes were equivalent to approximately a $10^{th}$ of the median absolute deviation in social skills.
 
@@
 
-The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \@ref(table-quadratic)).
+The predictive association of school average SES on social skills was not consistent across the SES gradient (see Table \@ref(tab:table-2)).
 Significant school average SES by student SES interactions were significant for peer and conduct problems for both teacher and parent reports and for teacher reported prosocial behavior.
-Interactions are plotted in Figures \@ref(fig:conduct)-\@ref(fig:social).
+Interactions are plotted in Figures \@ref(fig:peer)-\@ref(fig:social).
 Full model results can be found in supplementary materials.
@@
 As educational psychologists our study is framed in relation to assimilation effects in response to children's peer frames-of-reference.
-Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures [see \@bowles2001].
+Yet, economic theory tends to emphasize the socialization influence of teachers and educational structures [@bowles2001].
 Identifying and comparing these mechanisms is an important future direction for research.

### 2020-11-06 supplementary_materials.Rmd
clean full draft
Ed Researcher Version

@@
 library(DT)
+library(knitr)
+```
 
+# Participant Flow Diagram
+
+```{r}
+library(magick)
+pdf = magick::image_read(here::here("fig","flow.pdf"))
+image_convert(pdf, "png")
 ```
@@
 supp_linear %>% 
-  datatable(filter = 'top', caption = "Table S1. Full Linear Model Outputs")
+  datatable(filter = 'top', caption = "Table S1. Full Linear Model Outputs") %>%
+    formatRound(columns=4:6, digits=3)
 ```
@@
 supp_interaction %>%
-  datatable(filter = 'top', caption = "Table S2. Full Interaction Model Outputs")
+  datatable(filter = 'top', caption = "Table S2. Full Interaction Model Outputs") %>%
+    formatRound(columns=4:6, digits=3)
 ```
@@
 ```{r, child="data/social_skills_codebook.Rmd"}
-
 ```
